### 2025-05-04 docs/response/response.qmd
create manuscript and response

---
title: "Author Response to Reviewer Comments"
# format:
#   pdf: default
#bibliography: references.bib
engine: knitr
---

1.  The choice of W (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?

<!-- 2.  While the CAR model shows lower bias and RMSE in the artificial study, does this advantage persist in real-world scenarios, particularly where data sparsity or boundary effects exist? -->

<!-- 3.  The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model). Have the authors examined how sensitive the model results are to changes in $\rho$? -->

<!-- 4.  The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants? Would spatial effect maps or additional diagnostic tools enhance this understanding? -->

<!-- 5.  Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research (page 2). Before reaching this conclusion, did you compare GWR to other spatial regression models? -->

<!-- 6.  The scope of properties analysed includes residential houses and commercial real estate (page 16). How do the authors reconcile this with the abstract, which mentions only house price data? -->

<!-- 7.  After data cleaning, the final dataset included 598 entries representing multiple observations per sub-district (page 16). Why does the distribution of these data not appear on the map (Figure 6)? -->

<!-- 8.  The authors have applied point-level data using three different models (GLMM, GWR, and multilevel CAR) (page 17). How was this approach implemented, given that there is no explanation of how each observation was georeferenced? -->

<!-- 9.  Why is Table 5 not referred to in the text? -->

<!-- 10. The authors have argued that point-level models like GWR are limited when precise location data are unavailable (page 18). However, for the Lombok House Price Data, the authors have used point-level models. How do you reconcile this contradiction? -->

<!-- 11. In the Lombok house price analysis, no sensitivity analysis was conducted to test how different definitions of adjacency (W matrix) influence the results. Would the findings remain consistent if a different spatial weights structure was used? -->

<!-- 12. The study primarily evaluates CAR models within the Lombok context, but how well does it generalize to larger, high-density urban areas (e.g., Jakarta)? Would additional testing in a more complex metropolitan setting strengthen the validity of the conclusions? -->

### 2025-05-04 response/response.qmd
create manuscript and response

---
title: "Response to Journal's 2nd Reviewer Comments"
subtitle: "Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study"
author: 
  - name: Indira Puteri Kinasih
    orcid: 0009-0009-6959-6723
    email: indiraputeri@uinmataram.ac.id
---

1.  **The choice of $W$ (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?**

    Thank you for this insightful comment regarding the specification of the spatial weight matrix, $W$. We fully acknowledge that the choice of $W$—whether binary adjacency, row-standardized, or distance-based—can substantially influence the performance and interpretation of spatial models. In the present study, our primary objective was to evaluate and compare the performance of CAR, SAR, and GWR models using both simulated data and real-world data from Lombok. To maintain consistency across models and to isolate the structural differences in model behavior, we have, at this stage, limited our analysis to a binary adjacency matrix for $W$. We agree that exploring alternative specifications of $W$ (e.g., row-standardized or distance-decay matrices) is a valuable direction for further research. We plan to extend our analysis in future work to assess how varying spatial weight structures affect the robustness and predictive accuracy of these spatial models.

2.  **While the CAR model shows lower bias and RMSE in the artificial study, does this advantage persist in real-world scenarios, particularly where data sparsity or boundary effects exist?**

    Thank you for your thoughtful question. We acknowledge that in small-area data, bias can indeed become a concern due to limited information at the unit level. However, in the case of the multilevel CAR (mlvCAR) model, each area contains multiple observations, allowing us to model area-specific effects—functionally resembling a random effects structure. This hierarchical structure helps mitigate the impact of sparsity within individual areas.

    Regarding sample size, prior studies suggest that having approximately 15–30 observations per area is generally sufficient for stable estimation in multilevel frameworks. Given our data structure aligns with this recommendation, we believe the mlvCAR model offers a robust approach, especially in real-world scenarios where heterogeneity and partial pooling across areas are important. We thus recommend the use of mlvCAR to address both the potential bias and boundary effects effectively.

3.  **The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model). Have the authors examined how sensitive the model results are to changes in $\rho$?**

    In our study, we recognize that the value of $\rho$ plays a crucial role in determining the extent of spatial autocorrelation captured by the model. While we did not conduct a dedicated sensitivity analysis across a range of fixed $\rho$ values, our simulation framework inherently explores variation in $\rho$ through repeated data generation processes, in which $\rho$ is estimated as part of the model fitting. The observed variability in $\rho$ reflects the underlying spatial structure in the data rather than instability in the model.

    That said, we agree that a more systematic sensitivity analysis—such as fitting the model under different fixed values of $\rho$—could offer further insights into model robustness. We consider this a valuable direction for future work, especially in scenarios with varying levels of spatial autocorrelation or data sparsity.

4.  **The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants? Would spatial effect maps or additional diagnostic tools enhance this understanding?**

    Thank you for highlighting the importance of interpreting the spatial effects $\phi$ in the CAR model. We agree that while $\phi$ captures unobserved spatial heterogeneity, its interpretability in the context of house price determinants requires careful consideration. Rather than directly reflecting specific covariates, $\phi$ accounts for spatially structured residual variation—such as omitted neighborhood characteristics, local amenities, or spatial spillovers—that are not captured by observed variables.

    To aid interpretation, we acknowledge that visualizing $\phi$ through spatial effect maps could be a powerful diagnostic tool. Such maps can reveal clusters of under- or over-prediction and provide insights into areas where additional covariates or local effects might be relevant. In the revised version of this study, we will include a Lombok sub-district-level map displaying the estimated $\phi$ values to better communicate the spatial structure captured by the model and to guide further interpretation and refinement.

5.  **Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research (page 2). Before reaching this conclusion, did you compare GWR to other spatial regression models?**

    Thank you for this thoughtful question. In our manuscript, we highlight Geographically Weighted Regression (GWR) as a prominent method in the literature, particularly due to its frequent application in house price studies. This observation is based on a review of previous works rather than a performance-based comparison. While GWR is indeed widely used for its ability to model spatial heterogeneity at the local level, our study goes further by systematically comparing GWR with CAR and SAR models through simulation and real data applications. This comparative analysis aims to evaluate their relative strengths and limitations more objectively.

6.  **The scope of properties analysed includes residential houses and commercial real estate (page 16). How do the authors reconcile this with the abstract, which mentions only house price data?**

     We acknowledge that the description in the manuscript may not have been sufficiently clear. The original dataset obtained through web-scraping did include a broad range of property types—such as residential houses, villas, boarding houses, and vacant land. However, for the purposes of this study, we restricted our analysis to residential houses only. We will revise the manuscript accordingly to ensure that this scope is clearly stated in both the abstract and the data collection section.

7.  **After data cleaning, the final dataset included 598 entries representing multiple observations per sub-district (page 16). Why does the distribution of these data not appear on the map (Figure 6)?**

    For clarity, we note that the Lombok house price dataset used in this study does not contain precise geographic coordinates for each observation. As a result, it is not possible to accurately display the distribution of the data points on a map. While each observation is linked to a sub-district, the absence of point-level location data limits our ability to visualize them spatially within those areas. Therefore, for a more informative visualization, we are preparing a heatmap of house prices at the sub-district level, which will reflect the spatial variation in prices across Lombok using aggregated data.

8.  **The authors have applied point-level data using three different models (GLMM, GWR, and multilevel CAR) (page 17). How was this approach implemented, given that there is no explanation of how each observation was georeferenced?**

    Thank you for this important observation. We clarify that GLMM and multilevel CAR models do not require explicit geographic coordinates for each individual observation. Instead, they rely on an area-level identifier (ID) that links each observation to a corresponding spatial unit (e.g., sub-district).

    In contrast, GWR requires spatial coordinates for each observation. To meet this requirement, we generated random points around the centroid of each sub-district, using `st_sample()`, where the number of points corresponds to the number of observations in that area. Each point was placed within a 0.5 km radius from the centroid, ensuring a reasonable spatial spread while maintaining the area-specific context. This approach enabled us to simulate realistic spatial positions for property transactions in the absence of precise geolocation data.

9.  **Why is Table 5 not referred to in the text?**

    In the revised version of the manuscript, we have made sure to explicitly refer to Table 5 in the main text to improve clarity and consistency. We appreciate your careful reading and helpful suggestion.

10. **The authors have argued that point-level models like GWR are limited when precise location data are unavailable (page 18). However, for the Lombok House Price Data, the authors have used point-level models. How do you reconcile this contradiction?**

    As we noted in our response to Comment 8, the Lombok house price data do not come with precise geographic coordinates. In order to apply the `gwr()` function, which requires point-level location data, we generated synthetic coordinates by simulating random points around the centroid of each sub-district. While this workaround allowed us to run the GWR model, we acknowledge that it is not an ideal or practical solution for real-world applications. This limitation reinforces our argument that point-level models such as GWR can be challenging to apply meaningfully when precise location data are not available.

11. **In the Lombok house price analysis, no sensitivity analysis was conducted to test how different definitions of adjacency (W matrix) influence the results. Would the findings remain consistent if a different spatial weights structure was used?**

    As mentioned in our response to Comment 1, in this study we used a binary adjacency matrix to define spatial relationships, primarily to maintain consistency across the CAR and SAR models and to focus on comparing their general performance. We acknowledge that the choice of spatial weights matrix ($W$) can significantly influence model outcomes. While we have not conducted a sensitivity analysis in this current version, we agree that testing alternative specifications—such as row-standardized or distance-based matrices—would be valuable. We consider this an important direction for future work to assess the robustness of our findings under different definitions of spatial connectivity.

12. **The study primarily evaluates CAR models within the Lombok context, but how well does it generalize to larger, high-density urban areas (e.g., Jakarta)? Would additional testing in a more complex metropolitan setting strengthen the validity of the conclusions?**

    Thank you for this important point. Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater. In contrast, there is relatively limited research focused on low-density or semi-rural contexts such as Lombok. Our study aims to address this gap by evaluating the applicability and performance of spatial models, particularly the CAR model, in a setting that has different data characteristics and spatial structures compared to major metropolitan areas.

    We also acknowledge that in richer datasets—common in urban contexts where more granular, high-frequency, and multidimensional data are available—machine learning approaches may offer competitive or even superior predictive performance. However, our focus here is to assess interpretable spatial models in contexts where data limitations remain a challenge.

### 2025-05-04 docs/manuscript.qmd
create manuscript and response

---
title: 'Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study'
author: 
  - name: Indira Puteri Kinasih
    orcid: 0009-0009-6959-6723
    email: indiraputeri@uinmataram.ac.id
    affiliations: 
      - name: Universiti Brunei Darussalam
        id: 1
        department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410

      - name: Universitas Islam Negeri Mataram
        id: 2
        department: Mathematics Education, Faculty of Education
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    url: 'https://haziqj.ml'
    corresponding: true
    affiliations:
      - ref: 1
  - name: Elvynna Leong
    orcid: 0000-0003-3298-1010
    email: elvynna.leong@ubd.edu.bn
    corresponding: true
    affiliations:
      - ref: 1
date-modified: last-modified
abstract: |
  The modelling of property prices has been extensively studied in econometrics, with widely used approaches including generalised linear regression and geo- graphically weighted regression. These models commonly address local spatial correlations observed in property price data. However, despite its potential to capture spatial effects, the Conditional Autoregressive (CAR) model remains underutilised for this purpose. This study examines the robustness and predictive power of the CAR model, comparing it with established spatial models across three different datasets generation. An illustrative case study on Lombok house price data is also included. Simulation results showed that the CAR model demon- strates a distinct advantage, achieving lower bias and variability compared to other spatial regression models, effectively capturing neighbourhood-based spa- tial relationships, and exhibiting strong predictive power across various scenarios. In the Lombok case study, the CAR model outperformed other models, providing more precise estimates for property-related factors such as land size and built- up area. The results confirm that CAR’s spatial framework enables a nuanced analysis of property values across regions, enhancing accuracy in predictive mod- els. This study also reveals the distinct strengths and limitations of each model, offering insights into their predictive accuracy and applicability across diverse real estate contexts.

keywords: 
- Conditional auto-regressive
- simultaneous auto-regressive
- geograpichally weighted regression
- Property prices
- Neighbourhood structure
# format-options:
#   code-block-font-size: footnotesize
---

```{r setup, include = FALSE}
library(sf)
library(dplyr)
library(knitr)
library(kableExtra)
library(spdep)
```

# Introduction

Real estate valuation represents a complex challenge, requiring a nuanced understanding of spatial dynamics and interdependencies within property markets. Traditional valuation methods often overlook these spatial dimensions, resulting in incomplete predictions and less effective policy interventions (@pagourtzi2003real, @droj2024comprehensive, @mccord2014understanding). In response, spatial regression models have emerged as powerful tools to address these shortcomings by explicitly incorporating spatial relationships into the analysis of property prices (@stewart2018localized, @yang2019does, @soltani2021housing).

Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research. Known for its capacity to capture local variations in property prices, GWR addresses spatial heterogeneity by allowing coefficients to vary across different locations. Significant studies, such as those by @sisman2022modelling, @soltani2021housing, @brunsdon1996geographically, @yu2007modeling, and @lu2011geographically, have highlighted its effectiveness in revealing spatially varying relationships between property values and various explanatory factors, including structural and neighbourhood characteristics, locational attributes, and socio-economic variables.

Spatial Autoregressive (SAR) models account for spatial dependence in the dependent variable by including a spatial lag parameter ($\rho$), which measures the influence of neighboring values. This makes SAR well-suited for data where values are directly impacted by nearby areas or points, such as house prices influenced by surrounding properties. However, SAR relies on a global spatial structure, which can limit its ability to capture localised patterns. Additionally, according to @golgher2016interpret and @LeSage2014 the interdependence created by the spatial lag term introduces feedback effects, where values influence each other in a looping manner, making the interpretation of coefficients more complex.

Conditional autoregressive (CAR) models, in contrast, are designed to handle more localised spatial dependencies effectively, compare to SAR. They assume values at a location are conditionally dependent on neighboring areas, defined through an adjacency matrix $\mathbf{W}$. This approach models spatial dependence via a spatial random effect, $\phi$, which captures the influence of neighboring areas on the value at a specific location (@banerjee2014hierarchical). CAR models are particularly suited for areal data, such as aggregated district-level statistics, where spatial autocorrelation is prominent. Unlike GWR, CAR does not require exact point coordinates and provides better handling of spatial heterogeneity than SAR. Its ability to model finer localised effects through $\phi$ makes CAR a preferred choice for applications like district-level property price analysis, public health studies, or crime mapping (@de2012bayesian).

This study is motivated by the need to address the spatial intricacies that characterise real estate markets. Properties located near each other often share similar price trends, influenced by common amenities, and neighbourhood attributes. Incorporating spatial dependencies into property valuation allows spatial regression models to provide a deeper insight into market dynamics than traditional methods. It explores the theoretical basis, methodological structures, and practical uses of the GWR, SAR, and CAR models within house price modelling. Through a comparative analysis of these models, we aim to clarify their respective advantages, limitations, and appropriateness for enhancing the accuracy and detail of property market assessments, with a particular focus on the distinct context of Lombok, Indonesia.

The paper begins with a literature review on spatial regression techniques for modeling property prices. Section 3 outlines the theoretical foundations of spatial autoregressive models, including GLM, GWR, SAR, CAR, and their multilevel variants. Section 4 provides a comparative analysis of these models across three artificial study regions, evaluating their predictive accuracy and robustness. Section 5 applies these models to Lombok house price data, demonstrating their efficacy in capturing spatial patterns and improving real estate market insights.

# Related works

Recent studies have leveraged GWR to investigate the spatial heterogeneity of housing market determinants. For example, @wang2020geographically applied GWR to examine the impact of local built-environment factors on home prices across different phases of the housing market cycle. Their findings highlighted significant spatial variations, underscoring the critical role of local context in influencing home prices. Similarly, @lu2011geographically employed GWR with a non-Euclidean distance metric to analyze London house price data. By incorporating this alternative metric, they better captured the spatially varying relationships between house prices and their determinants, emphasizing the importance of considering different distance metrics in spatial analyses. In their 2018 study, @stewart2018localized explored spatial and temporal variations in the determinants of apartment prices in Seoul, South Korea, over a decade. Utilizing a hedonic price model with a spatio-temporal lag, calibrated through GWR, they revealed consistent spatial variations and strong spatial lag effects on house prices over time. Additionally, @cellmer2020spatial applied GWR to assess the spatial variations in determinants of the housing market in Poland. The study identified spatial patterns of autocorrelation in average housing prices and market activity, illustrating the value of GWR in examining localized effects of socio-demographic and environmental factors on the housing market. Collectively, these studies demonstrate the efficacy of GWR and its extensions in capturing the nuanced spatial and temporal variations in housing market data, providing valuable insights for urban planning and policy-making.

In parallel, numerous studies have demonstrated the versatility and applicability of SAR models across various fields. For instance, @fix2021simultaneous explored the use of SAR models for analyzing spatial extremes, highlighting their ability to capture spatial dependence structures in extreme values. @sarlas2015localized showcased the effectiveness of SAR models in transportation research by predicting localized speed variations, illustrating their utility in improving traffic flow predictions. @begueria2009comparison emphasized the advantages of SAR models in addressing spatial autocorrelation by comparing them with generalized least squares models, providing more accurate predictions in ecological and biogeographical studies.

In the realm of real estate, SAR models have been particularly valuable. @trojanek2018spatial used SAR models to analyze the spatial and temporal effects of subway availability on property prices in Warsaw. Their study revealed significant price premiums for properties located near subway stations, underscoring the importance of transportation infrastructure on real estate values. @bottero2017buildings applied SAR models to investigate the relationship between buildings' energy performance and real estate market value. Their findings illustrated the model's utility in linking environmental attributes to economic outcomes in the property market. Furthermore, @cellmer2019application demonstrated the practical relevance of SAR models in urban planning and property valuation by employing hierarchical SAR models to develop detailed land value maps in urban areas, underscoring the model's effectiveness in capturing the complexities of urban land valuation.

Despite their limited use in property price modelling, CAR models offer a promising approach for addressing spatial autocorrelation and generating reliable estimates, making them well-suited for future studies in house price analysis. @lee2013carbayes introduced the R package CARBayes, which supports Bayesian spatial modelling using conditional autoregressive priors. By including house price data as a case study, Lee demonstrated the flexibility of CAR models in capturing localised spatial dependencies relevant to real estate valuation. @wall2004close provided a comparative analysis of the spatial structures in CAR and SAR models, highlighting CAR’s strength in handling localised spatial interactions through an adjacency-based framework. While their work focused on educational and environmental applications, the insights are directly applicable to real estate modelling. Similarly, @ver2018spatial examined CAR models in ecological studies, emphasizing their robustness in managing spatial heterogeneity and clarifying their structural differences from SAR models. By addressing the limitations of SAR’s global spatial framework and excelling in the detection of localised patterns, CAR models emerge as a robust and adaptable methodology for accurately modelling spatial dynamics in property prices, advancing the precision of real estate valuation techniques.

# Study region
In the spatial modelling, the study region $\mathcal{S}$ is typically divided into $K$ distinct non-overlapping geographic units, denoted as $\mathcal{S}_k$ for $k = 1, 2, \dots, K$. Each geographic unit $\mathcal{S}_k$ is associated with a target variable $y_k$ and a set of explanatory factors represented as a vector $\mathbf{x}_k$.

In this study, the artificial study region was constructed to facilitate model simulations and analysis. This synthetic region serves as a spatial representation. The dataset consists of a simple feature (`sf`) collection. It spans a bounding box from longitude 116.0355 to 116.5 and latitude -8.500079 to -8.000057, using the WGS 84 coordinate reference system (CRS). The geometry type is polygon, indicating that each feature is a polygonal area, with 216 polygons in total. Each polygon is defined by two attributes: `area`, representing the size in square units, and `geometry`, specifying the polygon's boundaries with longitude and latitude coordinates. The artificial study region is illustrated in @fig-artfreg.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-artfreg
#| fig-cap: "Artificial study region used for simulation analysis, consisting of 216 non-overlapping polygonal areas. Each area represents a distinct spatial unit for modelling and analysis, providing a controlled environment to evaluate spatial models' performance"
#| out-width: "100%"
#| fig-align: "center"

knitr::include_graphics("figures/artfcountry_plot.png", auto_pdf = TRUE)
```

Another specific ingredient in spatial modelling is the existence of $W$ matrix, also known as the spatial weights matrix. It encodes the spatial relationships between areas in a study region. The weight matrix visually shows how different regions are connected to each other, indicating their neighbourhood relationships (@morris2019bayesian). To demonstrate this concept, let's take a simple example of a map with four regions $(R_1, R_2, R_3, R_4)$, as visualised in @fig-neighsimple.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-neighsimple
#| fig-cap: "Illustration of neighbourhood structure. The figure depicts a simplified spatial configuration where each numbered area represents a distinct spatial unit, demonstrating how neighbouring relationships can be defined for spatial modelling purposes."
#| out-width: "70%"
#| fig-align: "center"

knitr::include_graphics("figures/maprec.png", auto_pdf = TRUE)
```

This straightforward neighbourhood structure creates a single connected component with the following relationships: $(R_1 \sim R_2,\ R_2 \sim R_3,\ R_3 \sim R_4)$. Therefore the weight matrix  $\boldsymbol{W}$ is

$$
\begin{array}{ccccc}
    & R_1 & R_2 & R_3 & R_4 \\
R_1 & 0 & 1 & 0 & 0\\ 
R_2 & 1 & 0 & 1 & 0\\ 
R_3 & 0 & 1 & 0 & 1\\
R_4 & 0 & 0 & 1 & 0\\
\end{array}
$$

Later on, in the modelling stage, we will also need to construct a diagonal matrix $\boldsymbol{D}$. It is an $N \times N$ matrix where each diagonal element $d_{ii}$ denotes the number of neighbours of region $R_i$, while all non-diagonal elements are zero. This matrix is essential in spatial econometric models because it captures the local neighbourhood structure of each region. In the context of Geographically Weighted Regression (GWR), Spatial Autoregressive (SAR), and Conditional Autoregressive (CAR) models, $\boldsymbol{D}$ plays a pivotal role in defining spatial relationships and dependencies among different regions.

For instance, in SAR and CAR models, the diagonal matrix $\boldsymbol{D}$ aids in incorporating spatial lag and error components by appropriately weighting the influence of neighbouring regions (@wall2004close, @ver2018spatial). In GWR, this matrix assists in locally calibrating the model by reflecting the density and connectivity of regions (@brunsdon1996geographically, @stewart2018localized). Here, $\boldsymbol{D}$ is represented as:

$$
\begin{array}{ccccc}
   & R_1 & R_2 & R_3 & R_4 \\
R_1 & 1 & 0 & 0 & 0\\ 
R_2 & 0 & 2 & 0 & 0\\ 
R_3 & 0 & 0 & 2 & 0\\
R_4 & 0 & 0 & 0 & 1\\
\end{array}
$$

This matrix format indicates that region $R_1$ has one neighbour, regions $R_2$ and $R_3$ each have two neighbours, and region $R_4$ has one neighbour. By integrating this matrix into our modelling framework, we ensure that the spatial structure and interactions among regions are accurately represented, thereby enhancing the robustness and precision of our spatial analysis.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-neighartfreg
#| fig-cap: "Neighbourhood structure of the artificial region. This figure illustrates the spatial connectivity between polygons in the artificial study region, where each line represents a defined neighbour relationship based on spatial adjacency"
#| out-width: "70%"
#| fig-align: "center"

knitr::include_graphics("figures/nb_artfcountry.png", auto_pdf = TRUE)
```

In relation to our artificial study region, @fig-neighartfreg illustrates the connections between areas, represented by vertices or centroids for each area and nodes connecting them to each other. This visualisation highlights the spatial adjacency structure within the region.

# Spatial regression model
When discussing spatial regression, it's crucial to comprehend the basic notion of linear regression. In classical linear regression, the relationship between the dependent variable $y$ and the $x_1,x_2, \dots, x_p$ independent variables is expressed as

$$
y = \beta_0 + \beta_1x_1 + \dots + \beta_p x_p + \varepsilon
$${#eq-basicreg}

This formulation assumes a global relationship between the variables, where the coefficients $\beta_1,\beta_2, \dots, \beta_p$ are constant across the entire study area. In many spatial datasets, relationships between variables may exhibit spatial variation. For example, in the case of property pricing, a consistent rates of change assumption may not hold true universally. For example, the house price increase for an extra bedroom is often thought to be fixed across all locations. However, local customs or knowledge may actually dictate these rates, rather than a universal utility assigned to each commodity. For instance, in neighbourhoods with families, where extra space is highly valued, the perceived value of an additional bedroom may be greater compared to areas with singles or elderly couples, for whom extra space may not be as desirable.

## Geographically Weighted Regression (GWR)

@brunsdon1996geographically developed GWR which is one such technique that extends classical linear regression by allowing coefficients to vary spatially. It allows for the estimation of local relationships between a response variable and predictor variables. It is particularly useful for exploring spatial non-stationarity and identifying spatially varying relationships across different locations. The GWR model can be expressed as:

$$
y_i(s) = \beta_{0}(s) + \sum_{k=1}^p \beta_k(s)x_{ik}(s) + \varepsilon(s), \\ i = 1,\dots, n
$${#eq-gwr}

@eq-gwr represents a spatial regression model where $\mathbf{y}(s)$ is the dependent variable at location $s$. The term $\beta_{0}(s)$ is the spatially varying intercept, while $\sum_{k=1}^p \beta_k(s)\mathbf{x_k}(s)$ represents the spatially varying coefficients for the independent variables $\mathbf{x_k}(s)$. $\varepsilon(s)$ denotes the error term, capturing unexplained variation at location $s$

Using a weighted least squares approach to calibrate regression models allows different weights to be assigned to different observations, influencing the estimated parameters. In ordinary least squares, the goal is to minimize the sum of squared differences between predicted and actual $y$ values. Weighted least squares, however, apply a weighting factor $w$ to each squared difference, making some prediction inaccuracies more significant. If $w$ is a diagonal matrix of weights, the estimated coefficients are given by @eq-betagwr:

$$
\beta(s) = (\boldsymbol{x}^T \boldsymbol{W}(s) \boldsymbol{x})^{-1} \boldsymbol{x}^T \boldsymbol{W}(s) \boldsymbol{y}
$${#eq-betagwr}

This method allows GWR to address spatial heterogeneity by emphasizing observations near the location of interest, thereby improving the accuracy and relevance of local model estimates.

The estimation of GWR parameters involves fitting a separate regression equation for each location in the study area. Various estimation techniques can be employed, including ordinary least squares (OLS), weighted least squares (WLS), and maximum likelihood estimation (MLE). These techniques aim to optimize the model parameters to minimize the differences between the observed and predicted values of the dependent variable.

## Simultaneously Autoregressive (SAR) Models
While GWR focuses on capturing localised spatial heterogeneity by allowing coefficients to vary across space, SAR take a different approach by explicitly modelling spatial dependencies through a spatial lag. The SAR model is a spatial econometric model used to analyse spatial dependencies and relationships among observations in a geographic space (@anselin1988spatial). It is widely employed in various fields, such as regional economics, environmental studies, and urban planning. SAR is a type of spatial autoregressive model involving a simultaneous equation framework to capture spatial interactions.

The general form of SAR model can be expressed as follows

$$
y(s) = \rho \sum_{s'} w(s,s') y(s') + \sum_{k=1}^p \beta_k x_k(s) + \varepsilon(s)
$${#eq-sar}

where $\mathbf{Y}$ is the vector of observed values for the dependent variable, $W$ is the spatial weights matrix, $\rho$ is the spatial autoregressive coefficient, $X$ is the matrix of observed values for exogenous variables, $\beta$ is the vector of coefficients, and $\boldsymbol{\varepsilon} \sim N(0,\sigma^2)$ is the vector of error terms. Estimation of the SAR model parameters is typically done using statistical techniques such as maximum likelihood estimation (MLE) or generalised method of moments (GMM). The joint distribution of $\mathbf{Y}$ can be written as

$$
\mathbf{Y} \sim \mathcal{N}\left( \left( I - \rho W \right)^{-1} X\beta, \, \sigma^2 \left( I - \rho W \right)^{-1} \left( I - \rho W^T \right)^{-1} \right)
$${#eq-jointysar}

Extensions of the SAR model include the Spatial Lag Model, the Spatial Error Model, and the Spatial Durbin Model (@elhorst2014spatial), each incorporating distinct assumptions regarding the spatial configuration of the errors (@elhorst2012model; @anselin2013spatial). These extensions provide flexibility to account for different types of spatial dependencies and can be chosen based on the specific spatial relationships and hypotheses under investigation.

The SAR model, with its various specifications, provides a flexible framework to account for spatial dependencies and explore the spatial dynamics of observed phenomena. Researchers often choose between these models based on the nature of the spatial relationships in their data and the specific hypotheses they aim to test. The SAR model is a valuable tool for understanding spatial interdependence and making informed policy and planning decisions in diverse spatial contexts

## Conditional Autoregressive (CAR) Models

CAR models are a class of spatial statistical models used to analyze spatially structured data. The general formulation of a CAR model can be expressed as:

$$
y(s) = \sum_{k=1}^p \beta_k x_k(s) + \varepsilon(s) + \phi(s) 
$${#eq-car}

Here, $\boldsymbol{\varepsilon} \sim N(0, \nu^2)$ while $\boldsymbol{\phi}$ is a specific component in CAR model that has a role as spatial effect. It is also common to mention $\boldsymbol{\phi}$ as a CAR priors. It is a type of Gaussian Markov random field (@rue2005gaussian), capture spatial autocorrelation by ensuring that values at nearby locations are more similar than those further apart (@lee2013carbayes). This can be expressed in a general term

$$
\boldsymbol{\phi}\sim N(0,\tau^2\boldsymbol{Q}^{-1}) 
$${#eq-phidist}

where $\boldsymbol{Q}$ is a precision matrix that may be singular (intrinsic model). $\boldsymbol{Q}$ controls the spatial autocorrelation structure of the random effects, and is based on a non-negative symmetric $n \times n$ neighbourhood or weight matrix $\boldsymbol{W}$.

Together with the spatial weights matrix $\boldsymbol{W}$, the prior information are crucial components of CAR models. The choice of $\boldsymbol{W}$ determines the spatial structure of the model, while the priors for the variance parameters and the spatial random effects influence the model's ability to capture spatial dependencies. Commonly used priors for the variance parameters include inverse-gamma distributions, which provide flexibility and can be tuned to reflect prior beliefs about the scale of variability in the data. The prevailing approach typically involves a binary representation using geographical adjacency, where $w_{ki} = 1$ if areal units $(S_k,S_i)$ have a mutual boundary (denoted $k\sim i$, and is zero otherwise). This specification forces $(\phi_k,\phi_i)$ relating to geographically adjacent areas (that is $w_{ki} = 1$) to be correlated. On the other hand, random effects associated with areas that are not adjacent are independent of each other, provided we know the values of the other random effects.

A CAR prior was introduced by @leroux2000estimation and @stern1999inference to model diverse levels of spatial autocorrelation. In this type of prior, a single collection of random effects is utilised and its primary objective is to model spatial data, with a specific focus on dealing with spatial relationships and auto-correlation between data points. This model is especially proficient in the task of smoothing data and detecting spatial patterns within data set. The random effect $\boldsymbol\phi_k$ structured as follow:

$$
\phi_k|\boldsymbol\phi_{-K}, \boldsymbol{W},\rho, \tau^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\tau^2}{A}\right)
$${#eq-phicar}

where $A = \rho\sum_{i=1}^Kw_{ki}+1-\rho$. Note that when $\rho = 1$, the prior forms an intrinsic CAR prior [@besag1991bayesian], indicating full spatial dependency. Conversely, when $\rho = 0$, $A = 1$ it means that there will be no $W$ matrix role in the model, and it will become a comman random effect. In other words, the model reduces to a generalised linear model.

When handling $i$ observations within each area $k$, @eq-ycar closely resembles @eq-gwr in the GWR model, with the exception of the $\phi$ component. It is usually called multilevel CAR models

$$
y_i(s) = \sum_{k=1}^p \beta_k x_{ki}(s) + \varepsilon_i(s) + \phi(s)
$${#eq-ycar}

The Bayesian approach to CAR models entails defining prior distributions for all model parameters, such as the regression coefficients $\beta$, variance parameters $\nu^2$ and $\tau^2$, and spatial random effects $\phi_k$. Subsequently, Bayesian inference techniques like Markov Chain Monte Carlo (MCMC) are employed to derive posterior distributions of these parameters. This method facilitates the integration of prior knowledge and offers a versatile framework for quantifying uncertainty.

In general, @tbl-summarymethod below summarises the key elements utilised in the GWR, SAR, and CAR models:

```{r, echo = FALSE, message = FALSE}
#| label: tbl-summarymethod
#| tbl-cap: "Comparison of Model Elements Across GWR, SAR, and CAR"
#| eval: true

# add estimation method for each model
kbl(
  data.frame(
    Elements = c(
      "Number of Observations ($N$)", 
      "Covariate Matrix ($X$)", 
      "Spatial Weights Matrix ($W$)", 
      "Spatial Coordinates (coords)", 
      "Covariate Coefficients ($\\beta$)", 
      "Error Term variance ($\\sigma^2$)", 
      "Spatial Autocorrelation ($\\rho$)", 
      "Spatial Structure component ($\\phi$)", 
      "Spatial structure variance ($\\tau^2$)", 
      "Additional scale ($\\nu^2$)",
      "Estimation method"
    ),
    GWR = c("$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\times$", "$\\times$", "$\\times$", "MLE"),
    SAR = c("$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\times$", "$\\times$", "MLE/GMM"),
    CAR = c("$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "MCMC")
  ),
  col.names = c("Elements", "$GWR$", "$SAR$", "$CAR$"),
  escape = FALSE, # Needed to render LaTeX math symbols
  booktabs = TRUE,
  longtable = TRUE) %>% 
  kable_styling(font_size = 9)
```

# Comparative Analysis of GLM, SAR, and CAR

This section focuses on generating spatially autocorrelated property price data, exploring two distinct data generation frameworks: one using a CAR specification and another without spatial autocorrelation (non-CAR). The analysis is conducted within an artificially constructed study area using key property covariates: land size, building size, number of bedrooms, and number of bathrooms. The generated datasets are then employed to examine spatial dependencies in property prices and are analysed using a suite of models including GLM, SAR, and CAR.

## The experiment design

The experiment begins by creating a dataset based on the defined study region. The idea is to generate covariates that hypothetically can explain price of a property. It basically consist of property structural characteristics such as land size, building size, number of bedrooms, and number of bathrooms. To represent real-world data, each polygon contains multiple observations.

To generate the covariates, the first step involves creating area-level data with a specific spatial pattern, referred to as the housing density for the study region. The spatial pattern implies that certain areas or polygons will have a higher housing density than others. In this case, the central horizontal region of the study area is set to has a higher housing density, mirroring the housing distribution found on the island of Lombok. Although this density will not be directly utilized in the model simulation, it serves as a critical foundation for determining the spatial distribution of observation points across each polygon. In areas with higher density, a greater number of observation points are generated, which aligns with the notion that densely populated regions typically have more residential developments, thereby increasing the number of properties available for sale. Across the 216 polygons, a total of 4,314 observation points were generated. The number of observations per polygon varies from 5 to 56 points, reflecting the density in each area. @fig-pointsonmap illustrate how each polygon has multiple observations and they are distributed according to a specific pattern.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-pointsonmap
#| fig-cap: "Artificial study region with observation points. The points represent data locations within each polygon, simulating real-world spatial observations for the purpose of model analysis and validation."
#| out-width: "80%"
#| fig-align: "center"

knitr::include_graphics("figures/density_points_plot.png", auto_pdf = TRUE)
```

Following this, covariates were generated using `mvrnorm()` for each observation point, creating a dataset for further analysis. @tbl-covrstr provides a brief slice of the covariates’ structure.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-covrstr
#| tbl-cap: "Slice of generated dataset as covariates for simulations"
#| eval: true

dat <- readRDS("~/Study/R project/article_1/data/gen_dat_v3.rds")
dat <- dat %>% st_drop_geometry(NULL) %>% 
  select(-geometry) %>% 
  rename(density = population_density,
         "#bedrooms" = bedrooms,
         "#bathrooms" = bathrooms)

# Specify the pattern you want for the IDs
desired_pattern <- c(1, 1, 2, 2, 2, 3, 4, 4)

# Count occurrences of each ID in the desired pattern
pattern_count <- table(desired_pattern)

# Create a subset based on `desired_pattern` without excessive rows
dat_pattern <- dat %>%
  filter(ID %in% names(pattern_count)) %>%
  group_by(ID) %>%
  filter(row_number() <= pattern_count[as.character(ID)]) %>%
  ungroup() %>% 
  select(-area, -density) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

# Display the table in the document
knitr::kable(
  dat_pattern,
  escape = FALSE
) %>% kable_styling(font_size = 9)
```

By employing artificially generated covariates and an artificially constructed study region, we developed three distinct property price datasets, each based on a different spatial model: CAR, a combination of CAR and SAR, and GWR. The true parameter values for this purpose are $\boldsymbol{\beta} = \begin{bmatrix} 1 & 0.9 & 0.7 & 0.5 & 0.3 \end{bmatrix}$, $\nu^2 = 0.2$, $\tau^2 = 0.6$ and $\rho = 0.6$. Whereas beta values for GWR dataset is generated using a beta function, which is a function of coordinate points. The matrix $\mathbf{W}$ was derived from the neighbourhood structure of an artificial study region using the functions `poly2nb()` and `nb2mat()`. Data generation for $SAR + CAR$ employs a slightly different strategy. In addition to being generated per area simultaneously using the joint distribution of $\mathbf{Y}$ as in @eq-jointysar, information about the spatial effect $\boldsymbol{\phi}$ is also added at the end of the process. Therefore, we refer to it as $SAR + CAR$ data.

Each data generation resulting two types of datasets: area-level and point-level. These datasets were then modelled using six linear regression approaches, comprising both classical and spatial regression models: GLM, SAR, and CAR for area-level data, and GWR, GLMM, and multilevel CAR for point-level data. Simulations were conducted $sim = 1000$ times. From each simulation, coefficients, parameters, and fitted values were extracted, allowing for the computation of bias and RMSE values for robustness and prediction power analysis.

## Robustness and power prediction analysis

The bias values show that the models exhibit varying levels of accuracy across the different parameters and datasets. The coefficients $\beta_k$ for $k = 0, 1, 2, 3, 4$ represent the values associated with the intercept, land size, building size, number of bedrooms, and number of bathrooms, respectively. For $\beta_0$, the GLM and CAR models display relatively small biases with narrow standard deviations, indicating better consistency, while the SAR model demonstrates larger bias values, suggesting potential issues with accurately estimating the intercept. For higher parameters like $\beta_2$ and $\beta_3$, SAR generally has larger bias and variability, especially in the $SAR + CAR$ dataset, suggesting that the combined spatial structure poses challenges for SAR's estimation. It is important to note that the  @fig-biasarealevel utilises area-level data. In this context, the CAR model consistently demonstrates the smallest bias compared to the other two models.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-biasarealevel
#| fig-cap: "Bias values by method and dataset type for area-level data. This figure compares the bias estimates of different methods (CAR, GLM, SAR) across datasets generated under CAR, GWR, and SAR+CAR frameworks. Each method's performance is visualised for four coefficients ($\\beta_1$, $\\beta_2$, $\\beta_3$, $\\beta_4$), highlighting variations in bias and the effectiveness of each model in addressing spatial dependencies."
#| out-width: "75%"
#| fig-align: "center"

knitr::include_graphics("figures/bias1b_plot.pdf", auto_pdf = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-biaspointlevel
#| fig-cap: "Bias values by method and dataset type for point-level data. This figure compares the bias estimates of different methods (GLMM, GWR, and mlvCAR) across datasets generated under CAR, GWR, and SAR+CAR frameworks. Each method's performance is visualised for four coefficients ($\\beta_1$, $\\beta_2$, $\\beta_3$, $\\beta_4$), highlighting variations in bias and the effectiveness of each model in addressing spatial dependencies."
#| out-width: "75%"
#| fig-align: "center"

knitr::include_graphics("figures/bias2_plot.pdf", auto_pdf = TRUE)
```

As shown in the second plot of @fig-biaspointlevel, this analysis utilises point-level data. Overall, the mlvCAR model demonstrates a consistently lower bias across most parameters and scenarios. In the $SAR + CAR$ scenario, where spatial complexity is most pronounced, mlvCAR achieves notably low bias, particularly for parameters $\beta_1$ through $\beta_4$, indicating its suitability for handling mixed spatial structures. making it a solid choice for a range of spatial modelling tasks.

The GWR model, which provides highly localised estimates, displays slightly higher bias for several parameters compared to mlvCAR, especially for $\beta_1$, $\beta_2$, and $\beta_3$ within the $CAR$ scenario. This suggests GWR’s greater sensitivity to spatial patterns without incorporating neighbourhood-based effects. In contrast, GLMM, which lacks explicit spatial components, shows larger bias variability, particularly in scenarios with pronounced spatial dependencies, highlighting its limitations for purely spatial analyses.

```{r, echo = FALSE, message = FALSE}
#| label: tbl-rmse
#| tbl-cap: "RMSE values across different datasets"
#| eval: true

# Define the data frame
df <- data.frame(
  Dataset = c("$CAR$", "$SAR + CAR$", "$GWR$"),
  GLM     = c(4.09, 11.88, 2.43),
  SAR     = c(3.85, 11.68, 2.42),
  CAR     = c(1.81, 3.88, 2.28),
  GLMM    = c(14.04, 47.67, 18.31),
  GWR     = c(21.13, 48.87, 18.33),
  mlvCAR  = c(9.94, 38.97, 14.87)
)

  # GLM     = c(27.05, 76.62, 28.84),
  # SAR     = c(24.79, 75.72, 28.23),
  # CAR     = c(10.49, 27.71, 27.92),

# Render the table with math symbols in headers and custom header groups
kbl(
  df,
  col.names = c("Dataset", "$GLM$", "$SAR$", "$CAR$", "$GLMM$", "$GWR$", "$mlvCAR$"), # Custom column names with math symbols
  escape = FALSE, # Needed to correctly render LaTeX math symbols
  booktabs = TRUE,
  longtable = TRUE
) %>%
  add_header_above(c(" " = 1, "Area-level data" = 3, "Point-level data" = 3)) %>% # Add grouped headers
  kable_styling(latex_options = c("repeat_header"), font_size = 9) # Repeat headers on new pages if longtable

```

@tbl-rmse provide insights into model accuracy at different spatial levels. For area-level data, the models GLM, SAR, and CAR are compared, with the CAR model consistently achieving the lowest RMSE. This suggests that CAR is more effective at capturing spatial dependencies at the area level. For point-level data, the models GLMM, GWR, and mlvCAR are evaluated, with mlvCAR showing the lowest RMSE across all datasets. This indicates that mlvCAR is particularly well-suited for handling detailed spatial variations at the point level. Overall, these results highlight the strengths of CAR and mlvCAR in reducing error within their respective spatial contexts, underscoring their potential utility in spatial modeling applications.

# Applications in Lombok house prices dataset

In this section, we apply the simulated models to property price data on Lombok Island, Indonesia. Lombok is located in the eastern part of Indonesia, specifically between latitudes -8.9° and -8.1°, and longitudes 115.9° and 116.6°.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-lombokmap
#| fig-cap: "Lombok neighbour structure at the sub-district (Kecamatan) level. This figure illustrates the spatial neighbour structure of sub-districts on Lombok Island. The red lines represent the connections between neighbouring sub-districts based on adjacency relationships, overlaid on a base map."
#| out-width: "75%"
#| fig-align: "center"

knitr::include_graphics("figures/nb_lop2.png", auto_pdf = TRUE)
```

Lombok includes five districts (kota/kabupaten), 53 sub-districts (kecamatan), and 608 villages (desa/kelurahan), which represent three successive levels of administrative division. In this study, the sub-district level is used as the spatial unit of analysis, as illustrated in @fig-lombokmap. The key variables of interest include property price, land size, built-up area, number of bedrooms, and number of bathrooms.

## Data Collection

<!-- The dataset employed in this study comprises property price data from Lombok, Indonesia. The scope of properties analysed includes residential houses and commercial real estate. In addition to data on property prices, the dataset also encompasses various property characteristics including land area, built-up area, and the count of bedrooms and bathrooms. -->

Data were collected from multiple online sources. Property prices and their associated characteristics were harvested using web-scraping techniques from three leading Indonesian property trading platforms, https://www.lamudi.co.id/, https://www.99.co/, and https://www.rumah123.com/. The accuracy of the web-scraped data was ensured by cross-referencing it with official datasets, performing checks for missing or inconsistent entries, and validating key variables through sample comparisons with manually collected data. From this process, the initial dataset comprised 1,188 entries, with 9 variables, including village, prices, land-size, built-up area, number of bedrooms, number of bathrooms, floors, property type/category, and furnishing status.

Further, we conducted several preprocessing steps. First, we filtered the dataset by removing irrelevant variables, including floors, furnishing status, and category. It was then subsequently filtered to retain only properties with plausible characteristics: land area between 90 and 800 square meters, built-up size between 70 and 600 square meters, a maximum of 6 bedrooms, and a maximum 5 bathrooms. These thresholds were applied to ensure the data reflect realistic and context-appropriate housing characteristics in Lombok, based on common residential patterns and local housing norms. This step was essential to enhance the validity of the analysis by excluding outliers or potentially erroneous entries. Following this procedure, the dataset was refined to comprise a total of 576 observations. A summary of the key variables is presented in @tbl-datasumm.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| label: tbl-datasumm
#| tbl-cap: "Summary statistics of variables in the property price dataset"
#| eval: true

# Create the data frame
data <- data.frame(
  Variable = c("Prices (million IDR)", "Land-size (sqm)", "Built-up area (sqm)", 
               "No. of bedroom", "No. of bathroom"),
  Mean   = round(c(1983, 299.6, 184.8, 3.24, 2.55), 2),
  Median = round(c(1400, 255.0, 150.0, 3.00, 2.00), 2),
  Q1     = round(c(800, 147.2, 100.0, 2.00, 2.00), 2),
  Q3     = round(c(2500, 407.2, 210.0, 4.00, 3.00), 2),
  Min    = round(c(170, 90.0, 70.0, 1.00, 1.00), 2),
  Max    = round(c(22500, 800.0, 600.0, 6.00, 5.00), 2),
  SD     = round(c(2056.81, 179.39, 109.20, 1.19, 1.05), 2),
  `NA`   = c(26, 24, 24, 24, 24)
)

# Display the table with knitr::kable
kbl(data, 
         col.names = c("Variable", "Mean", "Median", 
                       "$Q_1$", "$Q_3$", "Min", "Max", "SD", "NA"),
    escape = FALSE,
    booktabs = TRUE,
    longtable = TRUE) %>% 
  kable_styling(font_size = 9)
```

Next, we merged the cleaned dataset with spatial administrative data by matching sub-district and district names. This process resulted in a dataset containing 598 entries. Compared to the previous dataset, which consisted of 576 observations, this indicates that 22 sub-districts did not have any property data available—meaning that property sales in those areas were not recorded on the online property listing platform used as the data source.

To address missing values, we employed the `mice` package (@van2011mice) to perform multiple imputation with parameters set to $maxit = 10$, $m = 10$, and a fixed random seed to ensure reproducibility. The imputed values were then reintegrated into the main dataset, replacing the original missing entries. This process resulted in a final dataset that was complete and ready for subsequent analysis using the selected spatial models.

In addition to the point-level dataset (`ppd`), which contains individual property listings, the imputed dataset was also aggregated to the sub-district level by calculating the mean values of key variables. This aggregation produced an area-level dataset (`ppd_kc`), allowing the analysis to be conducted at both the individual and administrative levels. These two levels of data granularity provide complementary perspectives for evaluating the performance of spatial models in capturing local property market dynamics.

## Model Implementation & Analysis

The cleaned and completed Lombok dataset was then model. The tables below present the results of the data fitting, offering a comparative view of each model’s effectiveness and accuracy in capturing the dataset's characteristics.

$$
prices \sim land_{std} + building_{std} + beds_{std} + baths_{std}
$${#eq-lopreg}

Standardization of the predictors ensures that the resulting coefficients are on a comparable scale, facilitating meaningful interpretation and comparison across different modeling approaches. The results presented in the following tables provide a comprehensive evaluation of each model’s performance in capturing both the structural and spatial characteristics of the property market in Lombok. This comparison allows us to assess the extent to which spatially explicit models (e.g., GWR, CAR, SAR) offer improvements over conventional methods in explaining price variability across regions. --- Let me know if you’d like to include metrics such as R², AIC, or Moran’s I in the paragraph as well.

We conducted model fitting on the area-level data, with detailed results provided in the Appendix (@tbl-modelcomp2). However, this method does not provide optimal parameter estimates because data aggregation can reduce variability and obscure spatial details essential for capturing local effects. To address this, we used point-level data to apply three different models: GLMM, GWR, and multilevel CAR.

```{r, echo=FALSE, message=FALSE}
#| label: tbl-modelcomp
#| tbl-cap: "Point-level Model Comparison"
#| eval: true

# Define the dataset using tribble
table_data <- tribble(
  ~"",                             ~GLMM,                    ~GWR,                     ~mlvCAR,
  "Intercept",                   "7.09 [6.93, 7.24]*",     "7.25 [7.25, 7.26]*",     "7.08 [6.99, 7.17]*",
  "Land-size",                   "0.26 [0.21, 0.32]*",     "0.31 [0.30, 0.32]*",     "0.26 [0.21, 0.32]*",
  "Built-up area",               "0.16 [0.10, 0.22]*",     "0.19 [0.19, 0.20]*",     "0.16 [0.10, 0.21]*",
  "No. of bedroom",              "-0.03 [-0.09, 0.01]",    "-0.15 [-0.17, -0.15]*",  "-0.03 [-0.09, 0.02]",
  "No. of bathroom",             "0.19 [0.13, 0.25]*",     "0.24 [0.23, 0.26]*",     "0.19 [0.13, 0.25]*",
  "$\\rho$",                     "-",                      "-",                      "0.48 [0.07, 0.90]*",
  "$\\nu^2$",                    "0.22 [0.19, 0.25]*",     "0.33 [0.29, 0.37]*",     "0.22 [0.20, 0.25]*",
  "$\\tau^2$",                   "0.20 [0.11, 0.34]*",     "-",                      "0.31 [0.14, 0.58]*",
  "AIC",                         "803.38",                 "-645.104",               "786.95",
  "DIC",                         "-",                      "-",                      "850.39",
  "WAIC",                        "838.3",                  "-",                      "851.14",
  "LMPL",                        "-421.31",               "-",                      "-427.21",
  "Log-likelihood",              "-400.69",                "-520.97",                "-385.47"
)

#table_data

kbl(table_data, booktabs = T, escape = FALSE) %>%
  kable_styling(font_size = 9) %>%
  pack_rows("Coefficients", 1, 5) %>%
  pack_rows("Spatial parameter", 6,6) %>%
  pack_rows("Variances", 7, 8) %>%
  pack_rows("Model fit criterion", 9, 13)
```

Here, @tbl-modelcomp compares parameter estimates and fit criteria for three models: GLMM, GWR, and multilevel CAR. Significant coefficients are marked with an asterisk (*). Key determinants like land-size and built-up area have positive, significant effects across models, while the number of bedrooms shows only minor, varying effects, indicating model-specific differences in capturing this variable's influence. Only the multilevel CAR model includes the spatial parameter $\rho$, with a significant estimate of 0.48, showing the model's ability to account for spatial dependence. Variance components $\nu^2$ (residual) and $\tau^2$ (spatial) are also detailed, with both variance terms significant. Fit criteria (AIC, WAIC, and Log-likelihood) suggest mlvCAR and GWR may better capture local spatial variation, with mlvCAR balancing spatial structure and precision.

# Results and Discussion

The findings from simulation consistently indicate that both the CAR and mlvCAR model outperform the other models in capturing spatial dependencies. This is evidenced by lower RMSE values and reduced bias across various dataset scenarios, highlighting the CAR model’s robustness. In datasets with strong spatial structures, such as those generated under $CAR$ and $SAR + CAR$ conditions, the CAR model achieves substantially lower RMSE values compared to GLM and SAR, suggesting its effectiveness in handling localized spatial variations.

In the context of house price modelling, it is common to encounter multiple sales observations within a single area, yet precise coordinates for each sale are often unavailable. Obtaining precise geographic coordinates in real-world datasets is often challenging due to limitations in data collection, privacy concerns, or the aggregation of data into broader administrative units.  This limitation renders point-level approaches, such as GWR, less effective. With multiple observations available for each area, models like mlvCAR offer a more practical and accurate solution, effectively capturing spatial dependencies at an area level without requiring precise point-level locations.

The results underscore the importance of using models that explicitly account for spatial dependencies in fields like property valuation, where spatial heterogeneity plays a significant role. The superior performance of the CAR model, particularly in datasets with strong spatial structure, suggests that it may be a more reliable choice for applications where local spatial variations are critical. The findings align with previous studies indicating that CAR models are better suited for managing spatial autocorrelation and heterogeneity, especially when spatial dependencies are strong.

Additionally, the limitations of point-level models such as GWR in the absence of precise location data highlight a practical challenge in spatial modeling. When only area-level data is available, the use of models like mlvCAR becomes essential, as it allows for spatial analysis without the need for detailed coordinates, thus expanding the applicability of spatial models in real-world settings. These insights suggest potential for further exploration of the CAR and mlvCAR models in various domains, especially in urban planning, real estate, and other fields where spatial relationships impact outcomes. Future research could examine the use of CAR and mlvCAR models with different spatial resolutions or apply these models to other datasets with varying levels of spatial dependency to further validate their effectiveness.

# Supplementary information {.unnumbered}

Not applicable

# Acknowledgments {.unnumbered}

Not applicable

# Declarations {.unnumbered}

* Funding
* Competing interests

No, I declare that the authors have no competing interests as defined by Springer, or other interests that might be perceived to influence the results and/or discussion reported in this paper.

* Ethics approval
* Consent to participate
* Consent for publication
* Availability of data and materials
* Code availability
* Authors' contributions

::: {.appendices latex=true}

# Simple Feature Data Slice {#secA1}

```{r, echo=FALSE, eval=TRUE, message=FALSE}
#| label: tbl-geoartfex
#| tbl-cap: "a slice of a simple feature collection with 6 polygon features and 1 attribute field, area(sqm), representing polygon sizes in square units. The dataset, using the WGS 84 CRS, spans longitudes 116.0355–116.5 and latitudes -8.500079–-8.000057. The geometry field defines each polygon's spatial boundaries, representing artificial study regions for spatial modeling in this research"
#| eval: true

artfcountry <- sf::read_sf("data/", quiet = TRUE)

# Create the table with a caption
artfcountry %>%
  head(n = 15) %>%
  kable() %>%
  kable_styling()
```

# Bias values for Area-level Data {#secA2}

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| label: tbl-bias1
#| tbl-cap: "Mean ± SD values for bias across different area-level datasets and parameters. This table presents the mean and standard deviation (SD) of bias values across area-level datasets (CAR, SAR + CAR, GWR) and spatial models (GLM, SAR, CAR), with $nsim = 1000$. This is also act as numerical representation of fig. 5 (top panel)"
#| eval: true

biasarea <- readRDS("data/biasarea1.rds")
avg <- round(biasarea$avg_bias, 3)
sd  <- round(biasarea$sd_bias, 2)

# Define the data frame
df <- data.frame(
  Dataset = c(
    "CAR", "SAR + CAR", "GWR",
    "CAR", "SAR + CAR", "GWR",
    "CAR", "SAR + CAR", "GWR",
    "CAR", "SAR + CAR", "GWR"
  ),
  GLM = c(
    paste0("$", sprintf("%.2e", avg[13]), " \\pm ", sprintf("%.4f", sd[13]), "$"),
    paste0("$", sprintf("%.2e", avg[14]), " \\pm ", sprintf("%.4f", sd[14]), "$"),
    paste0("$", sprintf("%.2e", avg[15]), " \\pm ", sprintf("%.4f", sd[15]), "$"),
    paste0("$", sprintf("%.2e", avg[16]), " \\pm ", sprintf("%.4f", sd[16]), "$"),
    paste0("$", sprintf("%.2e", avg[17]), " \\pm ", sprintf("%.4f", sd[17]), "$"),
    paste0("$", sprintf("%.2e", avg[18]), " \\pm ", sprintf("%.4f", sd[18]), "$"),
    paste0("$", sprintf("%.2e", avg[19]), " \\pm ", sprintf("%.4f", sd[19]), "$"),
    paste0("$", sprintf("%.2e", avg[20]), " \\pm ", sprintf("%.4f", sd[20]), "$"),
    paste0("$", sprintf("%.2e", avg[21]), " \\pm ", sprintf("%.4f", sd[21]), "$"),
    paste0("$", sprintf("%.2e", avg[22]), " \\pm ", sprintf("%.4f", sd[22]), "$"),
    paste0("$", sprintf("%.2e", avg[23]), " \\pm ", sprintf("%.4f", sd[23]), "$"),
    paste0("$", sprintf("%.2e", avg[24]), " \\pm ", sprintf("%.4f", sd[24]), "$")
  ),

  SAR = c(
    paste0("$", sprintf("%.2e", avg[25]), " \\pm ", sprintf("%.4f", sd[25]), "$"),
    paste0("$", sprintf("%.2e", avg[26]), " \\pm ", sprintf("%.4f", sd[26]), "$"),
    paste0("$", sprintf("%.2e", avg[27]), " \\pm ", sprintf("%.4f", sd[27]), "$"),
    paste0("$", sprintf("%.2e", avg[28]), " \\pm ", sprintf("%.4f", sd[28]), "$"),
    paste0("$", sprintf("%.2e", avg[29]), " \\pm ", sprintf("%.4f", sd[29]), "$"),
    paste0("$", sprintf("%.2e", avg[30]), " \\pm ", sprintf("%.4f", sd[30]), "$"),
    paste0("$", sprintf("%.2e", avg[31]), " \\pm ", sprintf("%.4f", sd[31]), "$"),
    paste0("$", sprintf("%.2e", avg[32]), " \\pm ", sprintf("%.4f", sd[32]), "$"),
    paste0("$", sprintf("%.2e", avg[33]), " \\pm ", sprintf("%.4f", sd[33]), "$"),
    paste0("$", sprintf("%.2e", avg[34]), " \\pm ", sprintf("%.4f", sd[34]), "$"),
    paste0("$", sprintf("%.2e", avg[35]), " \\pm ", sprintf("%.4f", sd[35]), "$"),
    paste0("$", sprintf("%.2e", avg[36]), " \\pm ", sprintf("%.4f", sd[36]), "$")
  ),

  CAR = c(
    paste0("$", sprintf("%.2e", avg[1]), " \\pm ", sprintf("%.4f", sd[1]), "$"),
    paste0("$", sprintf("%.2e", avg[2]), " \\pm ", sprintf("%.4f", sd[2]), "$"),
    paste0("$", sprintf("%.2e", avg[3]), " \\pm ", sprintf("%.4f", sd[3]), "$"),
    paste0("$", sprintf("%.2e", avg[4]), " \\pm ", sprintf("%.4f", sd[4]), "$"),
    paste0("$", sprintf("%.2e", avg[5]), " \\pm ", sprintf("%.4f", sd[5]), "$"),
    paste0("$", sprintf("%.2e", avg[6]), " \\pm ", sprintf("%.4f", sd[6]), "$"),
    paste0("$", sprintf("%.2e", avg[7]), " \\pm ", sprintf("%.4f", sd[7]), "$"),
    paste0("$", sprintf("%.2e", avg[8]), " \\pm ", sprintf("%.4f", sd[8]), "$"),
    paste0("$", sprintf("%.2e", avg[9]), " \\pm ", sprintf("%.4f", sd[9]), "$"),
    paste0("$", sprintf("%.2e", avg[10]), " \\pm ", sprintf("%.4f", sd[10]), "$"),
    paste0("$", sprintf("%.2e", avg[11]), " \\pm ", sprintf("%.4f", sd[11]), "$"),
    paste0("$", sprintf("%.2e", avg[12]), " \\pm ", sprintf("%.4f", sd[12]), "$")
  )
)

# Render the table with kableExtra
kbl(
  df, 
  booktabs = TRUE, 
  escape = FALSE,
  align = c("l", "c", "c", "c")
) %>%
  kable_styling(font_size = 9, latex_options = "hold_position") %>%
  add_header_above(c(" " = 1, "Mean ± SD" = 3)) %>%
  pack_rows("$\\beta_{1}$", 1, 3, escape = FALSE, bold = TRUE) %>%
  pack_rows("$\\beta_{2}$", 4, 6, escape = FALSE, bold = TRUE) %>%
  pack_rows("$\\beta_{3}$", 7, 9, escape = FALSE, bold = TRUE) %>%
  pack_rows("$\\beta_{4}$", 10, 12, escape = FALSE, bold = TRUE)

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| label: tbl-bias2
#| tbl-cap: "Mean ± SD values for bias across different point-level datasets and parameters. This table presents the mean and standard deviation (SD) of bias values across point-level datasets (CAR, SAR + CAR, GWR) and spatial models (GLM, SAR, CAR), with $nsim = 1000$. This is also act as numerical representation of fig. 5 (bottom panel)"
#| eval: true

biaspoint <- readRDS("~/Study/R project/article_1/data/biaspoint.rds")
avg <- round(biaspoint$avg_bias, 5)
sd  <- round(biaspoint$sd_bias, 4)

# Define the data frame
df <- data.frame(
  Dataset = c(
    "CAR", "SAR + CAR", "GWR",
    "CAR", "SAR + CAR", "GWR",
    "CAR", "SAR + CAR", "GWR",
    "CAR", "SAR + CAR", "GWR"
  ),
  GLMM = c(
    paste0("$", sprintf("%.2e", avg[1]), " \\pm ", sprintf("%.4f", sd[1]), "$"),
    paste0("$", sprintf("%.2e", avg[2]), " \\pm ", sprintf("%.4f", sd[2]), "$"),
    paste0("$", sprintf("%.2e", avg[3]), " \\pm ", sprintf("%.4f", sd[3]), "$"),
    paste0("$", sprintf("%.2e", avg[4]), " \\pm ", sprintf("%.4f", sd[4]), "$"),
    paste0("$", sprintf("%.2e", avg[5]), " \\pm ", sprintf("%.4f", sd[5]), "$"),
    paste0("$", sprintf("%.2e", avg[6]), " \\pm ", sprintf("%.4f", sd[6]), "$"),
    paste0("$", sprintf("%.2e", avg[7]), " \\pm ", sprintf("%.4f", sd[7]), "$"),
    paste0("$", sprintf("%.2e", avg[8]), " \\pm ", sprintf("%.4f", sd[8]), "$"),
    paste0("$", sprintf("%.2e", avg[9]), " \\pm ", sprintf("%.4f", sd[9]), "$"),
    paste0("$", sprintf("%.2e", avg[10]), " \\pm ", sprintf("%.4f", sd[10]), "$"),
    paste0("$", sprintf("%.2e", avg[11]), " \\pm ", sprintf("%.4f", sd[11]), "$"),
    paste0("$", sprintf("%.2e", avg[12]), " \\pm ", sprintf("%.4f", sd[12]), "$")
  ),
  GWR = c(
    paste0("$", sprintf("%.2e", avg[13]), " \\pm ", sprintf("%.4f", sd[13]), "$"),
    paste0("$", sprintf("%.2e", avg[14]), " \\pm ", sprintf("%.4f", sd[14]), "$"),
    paste0("$", sprintf("%.2e", avg[15]), " \\pm ", sprintf("%.4f", sd[15]), "$"),
    paste0("$", sprintf("%.2e", avg[16]), " \\pm ", sprintf("%.4f", sd[16]), "$"),
    paste0("$", sprintf("%.2e", avg[17]), " \\pm ", sprintf("%.4f", sd[17]), "$"),
    paste0("$", sprintf("%.2e", avg[18]), " \\pm ", sprintf("%.4f", sd[18]), "$"),
    paste0("$", sprintf("%.2e", avg[19]), " \\pm ", sprintf("%.4f", sd[19]), "$"),
    paste0("$", sprintf("%.2e", avg[20]), " \\pm ", sprintf("%.4f", sd[20]), "$"),
    paste0("$", sprintf("%.2e", avg[21]), " \\pm ", sprintf("%.4f", sd[21]), "$"),
    paste0("$", sprintf("%.2e", avg[22]), " \\pm ", sprintf("%.4f", sd[22]), "$"),
    paste0("$", sprintf("%.2e", avg[23]), " \\pm ", sprintf("%.4f", sd[23]), "$"),
    paste0("$", sprintf("%.2e", avg[24]), " \\pm ", sprintf("%.4f", sd[24]), "$")
  ),
  mlvCAR = c(
    paste0("$", sprintf("%.2e", avg[25]), " \\pm ", sprintf("%.4f", sd[25]), "$"),
    paste0("$", sprintf("%.2e", avg[26]), " \\pm ", sprintf("%.4f", sd[26]), "$"),
    paste0("$", sprintf("%.2e", avg[27]), " \\pm ", sprintf("%.4f", sd[27]), "$"),
    paste0("$", sprintf("%.2e", avg[28]), " \\pm ", sprintf("%.4f", sd[28]), "$"),
    paste0("$", sprintf("%.2e", avg[29]), " \\pm ", sprintf("%.4f", sd[29]), "$"),
    paste0("$", sprintf("%.2e", avg[30]), " \\pm ", sprintf("%.4f", sd[30]), "$"),
    paste0("$", sprintf("%.2e", avg[31]), " \\pm ", sprintf("%.4f", sd[31]), "$"),
    paste0("$", sprintf("%.2e", avg[32]), " \\pm ", sprintf("%.4f", sd[32]), "$"),
    paste0("$", sprintf("%.2e", avg[33]), " \\pm ", sprintf("%.4f", sd[33]), "$"),
    paste0("$", sprintf("%.2e", avg[34]), " \\pm ", sprintf("%.4f", sd[34]), "$"),
    paste0("$", sprintf("%.2e", avg[35]), " \\pm ", sprintf("%.4f", sd[35]), "$"),
    paste0("$", sprintf("%.2e", avg[36]), " \\pm ", sprintf("%.4f", sd[36]), "$")
  )
)

# Render the table with kableExtra
kbl(
  df, 
  booktabs = TRUE, 
  escape = FALSE,
  align = c("l", "c", "c", "c")
) %>%
  kable_styling(font_size = 9, latex_options = "hold_position") %>%
  add_header_above(c(" " = 1, "Mean ± SD" = 3)) %>%
  pack_rows("$\\beta_{1}$", 1, 3, escape = FALSE, bold = TRUE) %>%
  pack_rows("$\\beta_{2}$", 4, 6, escape = FALSE, bold = TRUE) %>%
  pack_rows("$\\beta_{3}$", 7, 9, escape = FALSE, bold = TRUE) %>%
  pack_rows("$\\beta_{4}$", 10, 12, escape = FALSE, bold = TRUE)

```

# Model Comparison for area-level data {#secA3}

```{r, echo=FALSE, message=FALSE}
#| label: tbl-modelcomp2
#| tbl-cap: "This table compares the performance of GLM, SAR, and CAR models on area-level data. It reports parameter estimates (with 95% confidence intervals), spatial parameters ($\\rho$), variances ($\\nu^2$, $\\tau^2$), and model fit criteria (AIC, DIC, WAIC, LMPL, log-likelihood). The results highlight each model's capability to capture spatial dependencies and variability"
#| eval: true

# Define the dataset using tribble
table_data <- tribble(
  ~"",                             ~GLM,                    ~SAR,                     ~CAR,
  "Intercept",                   "7.19 [7.02, 7.35]*",     "6.92 [6.48, 7.37]*",     "7.19 [7.03, 7.35]*",
  "Land-size",                   "0.45 [0.24, 0.66]*",     "0.45 [0.25, 0.64]*",     "0.45 [0.23, 0.65]*",
  "Built-up area",               "0.01 [-0.20, 0.23]",     "0.01 [-0.18, 0.22]",     "0.02 [-0.20, 0.23]",
  "No. of bedroom",              "-0.20 [-0.43, 0.02]",    "-0.23 [-0.45, -0.01]*",  "-0.21 [-0.44, 0.02]",
  "No. of bathroom",             "0.54 [0.33, 0.76]*",     "0.56 [0.36, 0.77]*",     "0.55 [0.34, 0.76]*",
  "$\\rho$",                     "-",                      "0.009[-0.005, 0.023]",    "0.39 [0.01, 0.91]*",
  "$\\nu^2$",                    "0.31 [0.21, 0.45]*",     "0.27 [0.19, 0.41]*",     "0.28 [0.01, 0.44]*",
  "$\\tau^2$",                   "0.20 [0.11, 0.34]*",     "-",                      "0.04 [0.01, 0.91]*",
  "AIC",                         "-",                       "-",                          "-",
  "DIC",                         "94.73",                  "94.73",                   "82.52",
  "WAIC",                        "95.64",                  "-",                      "87.43",
  "LMPL",                        "-48.04",               "-",                         "-46.71",
  "Log-likelihood",              "-41.49",                "-41.36",                  "-37.29"
)

kbl(
  table_data,
    booktabs = T, escape = FALSE) %>%
  kable_styling(font_size = 9, latex_options = "hold_position") %>%
  pack_rows("Coefficients", 1, 5) %>%
  pack_rows("Spatial parameter", 6,6) %>%
  pack_rows("Variances", 7, 8) %>% 
  pack_rows("Model fit criterion", 9, 13)
```

# Example of beta-values in GWR model {#secA4}

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#| label: fig-betaplot
#| fig-cap: "Spatial variation of $\\beta$ values generated using a custom beta function. The color gradient represents the magnitude of $\\beta$ values across the artificial study region, highlighting localised patterns and heterogeneity. The visualisation demonstrates the spatially varying coefficient structure modeled in the study"
#| out-width: "90%"
#| fig-align: "center"

knitr::include_graphics("figures/betavalues_plot.png", auto_pdf = TRUE)
```

# References

### 2025-05-04 docs/_tex/manuscript.tex
add doc folder

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  default,
]{sn-jnl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\captionsetup[figure]{font=small}
\captionsetup[table]{font = small}
\usepackage{xcolor}
\definecolor{mycodecolor}{RGB}{0,0,255}
\usepackage{fancyvrb}
\BeforeBeginEnvironment{verbatim}{\small}
\fvset{fontsize=\small}
\usepackage{setspace}\singlespacing
\usepackage{float}
\usepackage{booktabs}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study},
  pdfauthor={Indira Puteri Kinasih; Haziq Jamil; Elvynna Leong},
  pdfkeywords={Conditional auto-regressive, simultaneous
auto-regressive, geograpichally weighted regression, Property
prices, Neighbourhood structure},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Conditionally Autoregressive Models for House Price Data:
Insights From a Comparative Simulation Study]{Conditionally
Autoregressive Models for House Price Data: Insights From a Comparative
Simulation Study}

% author setup
\author[1,2]{\fnm{Indira Puteri} \sur{Kinasih}}\email{indiraputeri@uinmataram.ac.id}\author*[1]{\fnm{Haziq} \sur{Jamil}}\email{haziq.jamil@ubd.edu.bn}\author*[1]{\fnm{Elvynna} \sur{Leong}}\email{elvynna.leong@ubd.edu.bn}
% affil setup
\affil[1]{\orgdiv{Mathematical Sciences, Faculty of
Science}, \orgname{Universiti Brunei
Darussalam}, \orgaddress{\street{Universiti Brunei Darussalam, Jalan
Tungku Link}, \city{Bandar Seri Begawan}, \postcode{BE
1410}, \country{Brunei}}}
\affil[2]{\orgdiv{Mathematics Education, Faculty of
Education}, \orgname{Universitas Islam Negeri Mataram}}

% abstract

\abstract{The modelling of property prices has been extensively studied
in econometrics, with widely used approaches including generalised
linear regression and geo- graphically weighted regression. These models
commonly address local spatial correlations observed in property price
data. However, despite its potential to capture spatial effects, the
Conditional Autoregressive (CAR) model remains underutilised for this
purpose. This study examines the robustness and predictive power of the
CAR model, comparing it with established spatial models across three
different datasets generation. An illustrative case study on Lombok
house price data is also included. Simulation results showed that the
CAR model demon- strates a distinct advantage, achieving lower bias and
variability compared to other spatial regression models, effectively
capturing neighbourhood-based spa- tial relationships, and exhibiting
strong predictive power across various scenarios. In the Lombok case
study, the CAR model outperformed other models, providing more precise
estimates for property-related factors such as land size and built- up
area. The results confirm that CAR's spatial framework enables a nuanced
analysis of property values across regions, enhancing accuracy in
predictive mod- els. This study also reveals the distinct strengths and
limitations of each model, offering insights into their predictive
accuracy and applicability across diverse real estate contexts.}

% keywords
\keywords{Conditional auto-regressive,  simultaneous
auto-regressive,  geograpichally weighted regression,  Property
prices,  Neighbourhood structure}

\begin{document}
\maketitle

\section{Introduction}\label{introduction}

Real estate valuation represents a complex challenge, requiring a
nuanced understanding of spatial dynamics and interdependencies within
property markets. Traditional valuation methods often overlook these
spatial dimensions, resulting in incomplete predictions and less
effective policy interventions (Pagourtzi et al. (2003), Droj,
Kwartnik-Pruc, and Droj (2024), McCord et al. (2014)). In response,
spatial regression models have emerged as powerful tools to address
these shortcomings by explicitly incorporating spatial relationships
into the analysis of property prices (Stewart Fotheringham and Park
(2018), Yang et al. (2019), Soltani et al. (2021)).

Among the various spatial regression models, Geographically Weighted
Regression (GWR) is particularly prominent in property price research.
Known for its capacity to capture local variations in property prices,
GWR addresses spatial heterogeneity by allowing coefficients to vary
across different locations. Significant studies, such as those by Sisman
and Aydinoglu (2022), Soltani et al. (2021), Brunsdon, Fotheringham, and
Charlton (1996), Yu (2007), and Lu, Charlton, and Fotheringham (2011),
have highlighted its effectiveness in revealing spatially varying
relationships between property values and various explanatory factors,
including structural and neighbourhood characteristics, locational
attributes, and socio-economic variables.

Spatial Autoregressive (SAR) models account for spatial dependence in
the dependent variable by including a spatial lag parameter (\(\rho\)),
which measures the influence of neighboring values. This makes SAR
well-suited for data where values are directly impacted by nearby areas
or points, such as house prices influenced by surrounding properties.
However, SAR relies on a global spatial structure, which can limit its
ability to capture localised patterns. Additionally, according to
Golgher and Voss (2016) and LeSage and Pace (2014) the interdependence
created by the spatial lag term introduces feedback effects, where
values influence each other in a looping manner, making the
interpretation of coefficients more complex.

Conditional autoregressive (CAR) models, in contrast, are designed to
handle more localised spatial dependencies effectively, compare to SAR.
They assume values at a location are conditionally dependent on
neighboring areas, defined through an adjacency matrix \(\mathbf{W}\).
This approach models spatial dependence via a spatial random effect,
\(\phi\), which captures the influence of neighboring areas on the value
at a specific location (Banerjee, Carlin, and Gelfand (2014)). CAR
models are particularly suited for areal data, such as aggregated
district-level statistics, where spatial autocorrelation is prominent.
Unlike GWR, CAR does not require exact point coordinates and provides
better handling of spatial heterogeneity than SAR. Its ability to model
finer localised effects through \(\phi\) makes CAR a preferred choice
for applications like district-level property price analysis, public
health studies, or crime mapping (De Oliveira (2012)).

This study is motivated by the need to address the spatial intricacies
that characterise real estate markets. Properties located near each
other often share similar price trends, influenced by common amenities,
and neighbourhood attributes. Incorporating spatial dependencies into
property valuation allows spatial regression models to provide a deeper
insight into market dynamics than traditional methods. It explores the
theoretical basis, methodological structures, and practical uses of the
GWR, SAR, and CAR models within house price modelling. Through a
comparative analysis of these models, we aim to clarify their respective
advantages, limitations, and appropriateness for enhancing the accuracy
and detail of property market assessments, with a particular focus on
the distinct context of Lombok, Indonesia.

The paper begins with a literature review on spatial regression
techniques for modeling property prices. Section 3 outlines the
theoretical foundations of spatial autoregressive models, including GLM,
GWR, SAR, CAR, and their multilevel variants. Section 4 provides a
comparative analysis of these models across three artificial study
regions, evaluating their predictive accuracy and robustness. Section 5
applies these models to Lombok house price data, demonstrating their
efficacy in capturing spatial patterns and improving real estate market
insights.

\section{Related works}\label{related-works}

Recent studies have leveraged GWR to investigate the spatial
heterogeneity of housing market determinants. For example, Wang and Chen
(2020) applied GWR to examine the impact of local built-environment
factors on home prices across different phases of the housing market
cycle. Their findings highlighted significant spatial variations,
underscoring the critical role of local context in influencing home
prices. Similarly, Lu, Charlton, and Fotheringham (2011) employed GWR
with a non-Euclidean distance metric to analyze London house price data.
By incorporating this alternative metric, they better captured the
spatially varying relationships between house prices and their
determinants, emphasizing the importance of considering different
distance metrics in spatial analyses. In their 2018 study, Stewart
Fotheringham and Park (2018) explored spatial and temporal variations in
the determinants of apartment prices in Seoul, South Korea, over a
decade. Utilizing a hedonic price model with a spatio-temporal lag,
calibrated through GWR, they revealed consistent spatial variations and
strong spatial lag effects on house prices over time. Additionally,
Cellmer, Cichulska, and Bełej (2020) applied GWR to assess the spatial
variations in determinants of the housing market in Poland. The study
identified spatial patterns of autocorrelation in average housing prices
and market activity, illustrating the value of GWR in examining
localized effects of socio-demographic and environmental factors on the
housing market. Collectively, these studies demonstrate the efficacy of
GWR and its extensions in capturing the nuanced spatial and temporal
variations in housing market data, providing valuable insights for urban
planning and policy-making.

In parallel, numerous studies have demonstrated the versatility and
applicability of SAR models across various fields. For instance, Fix,
Cooley, and Thibaud (2021) explored the use of SAR models for analyzing
spatial extremes, highlighting their ability to capture spatial
dependence structures in extreme values. Sarlas and Axhausen (2015)
showcased the effectiveness of SAR models in transportation research by
predicting localized speed variations, illustrating their utility in
improving traffic flow predictions. Beguería and Pueyo (2009) emphasized
the advantages of SAR models in addressing spatial autocorrelation by
comparing them with generalized least squares models, providing more
accurate predictions in ecological and biogeographical studies.

In the realm of real estate, SAR models have been particularly valuable.
Trojanek and Gluszak (2018) used SAR models to analyze the spatial and
temporal effects of subway availability on property prices in Warsaw.
Their study revealed significant price premiums for properties located
near subway stations, underscoring the importance of transportation
infrastructure on real estate values. Bottero et al. (2017) applied SAR
models to investigate the relationship between buildings' energy
performance and real estate market value. Their findings illustrated the
model's utility in linking environmental attributes to economic outcomes
in the property market. Furthermore, Cellmer, Kobylińska, and Bełej
(2019) demonstrated the practical relevance of SAR models in urban
planning and property valuation by employing hierarchical SAR models to
develop detailed land value maps in urban areas, underscoring the
model's effectiveness in capturing the complexities of urban land
valuation.

Despite their limited use in property price modelling, CAR models offer
a promising approach for addressing spatial autocorrelation and
generating reliable estimates, making them well-suited for future
studies in house price analysis. Lee (2013) introduced the R package
CARBayes, which supports Bayesian spatial modelling using conditional
autoregressive priors. By including house price data as a case study,
Lee demonstrated the flexibility of CAR models in capturing localised
spatial dependencies relevant to real estate valuation. Wall (2004)
provided a comparative analysis of the spatial structures in CAR and SAR
models, highlighting CAR's strength in handling localised spatial
interactions through an adjacency-based framework. While their work
focused on educational and environmental applications, the insights are
directly applicable to real estate modelling. Similarly, Ver Hoef et al.
(2018) examined CAR models in ecological studies, emphasizing their
robustness in managing spatial heterogeneity and clarifying their
structural differences from SAR models. By addressing the limitations of
SAR's global spatial framework and excelling in the detection of
localised patterns, CAR models emerge as a robust and adaptable
methodology for accurately modelling spatial dynamics in property
prices, advancing the precision of real estate valuation techniques.

\section{Study region}\label{study-region}

In the spatial modelling, the study region \(\mathcal{S}\) is typically
divided into \(K\) distinct non-overlapping geographic units, denoted as
\(\mathcal{S}_k\) for \(k = 1, 2, \dots, K\). Each geographic unit
\(\mathcal{S}_k\) is associated with a target variable \(y_k\) and a set
of explanatory factors represented as a vector \(\mathbf{x}_k\).

In this study, the artificial study region was constructed to facilitate
model simulations and analysis. This synthetic region serves as a
spatial representation. The dataset consists of a simple feature
(\texttt{sf}) collection. It spans a bounding box from longitude
116.0355 to 116.5 and latitude -8.500079 to -8.000057, using the WGS 84
coordinate reference system (CRS). The geometry type is polygon,
indicating that each feature is a polygonal area, with 216 polygons in
total. Each polygon is defined by two attributes: \texttt{area},
representing the size in square units, and \texttt{geometry}, specifying
the polygon's boundaries with longitude and latitude coordinates. The
artificial study region is illustrated in Figure~\ref{fig-artfreg}.

\phantomsection\label{cell-fig-artfreg}
\begin{figure}[H]

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{figures/artfcountry_plot.png}

}

\caption{\label{fig-artfreg}Artificial study region used for simulation
analysis, consisting of 216 non-overlapping polygonal areas. Each area
represents a distinct spatial unit for modelling and analysis, providing
a controlled environment to evaluate spatial models' performance}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

Another specific ingredient in spatial modelling is the existence of
\(W\) matrix, also known as the spatial weights matrix. It encodes the
spatial relationships between areas in a study region. The weight matrix
visually shows how different regions are connected to each other,
indicating their neighbourhood relationships (Morris et al. (2019)). To
demonstrate this concept, let's take a simple example of a map with four
regions \((R_1, R_2, R_3, R_4)\), as visualised in
Figure~\ref{fig-neighsimple}.

\phantomsection\label{cell-fig-neighsimple}
\begin{figure}[H]

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{figures/maprec.png}

}

\caption{\label{fig-neighsimple}Illustration of neighbourhood structure.
The figure depicts a simplified spatial configuration where each
numbered area represents a distinct spatial unit, demonstrating how
neighbouring relationships can be defined for spatial modelling
purposes.}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

This straightforward neighbourhood structure creates a single connected
component with the following relationships:
\((R_1 \sim R_2,\ R_2 \sim R_3,\ R_3 \sim R_4)\). Therefore the weight
matrix \(\boldsymbol{W}\) is

\[
\begin{array}{ccccc}
    & R_1 & R_2 & R_3 & R_4 \\
R_1 & 0 & 1 & 0 & 0\\ 
R_2 & 1 & 0 & 1 & 0\\ 
R_3 & 0 & 1 & 0 & 1\\
R_4 & 0 & 0 & 1 & 0\\
\end{array}
\]

Later on, in the modelling stage, we will also need to construct a
diagonal matrix \(\boldsymbol{D}\). It is an \(N \times N\) matrix where
each diagonal element \(d_{ii}\) denotes the number of neighbours of
region \(R_i\), while all non-diagonal elements are zero. This matrix is
essential in spatial econometric models because it captures the local
neighbourhood structure of each region. In the context of Geographically
Weighted Regression (GWR), Spatial Autoregressive (SAR), and Conditional
Autoregressive (CAR) models, \(\boldsymbol{D}\) plays a pivotal role in
defining spatial relationships and dependencies among different regions.

For instance, in SAR and CAR models, the diagonal matrix
\(\boldsymbol{D}\) aids in incorporating spatial lag and error
components by appropriately weighting the influence of neighbouring
regions (Wall (2004), Ver Hoef et al. (2018)). In GWR, this matrix
assists in locally calibrating the model by reflecting the density and
connectivity of regions (Brunsdon, Fotheringham, and Charlton (1996),
Stewart Fotheringham and Park (2018)). Here, \(\boldsymbol{D}\) is
represented as:

\[
\begin{array}{ccccc}
   & R_1 & R_2 & R_3 & R_4 \\
R_1 & 1 & 0 & 0 & 0\\ 
R_2 & 0 & 2 & 0 & 0\\ 
R_3 & 0 & 0 & 2 & 0\\
R_4 & 0 & 0 & 0 & 1\\
\end{array}
\]

This matrix format indicates that region \(R_1\) has one neighbour,
regions \(R_2\) and \(R_3\) each have two neighbours, and region \(R_4\)
has one neighbour. By integrating this matrix into our modelling
framework, we ensure that the spatial structure and interactions among
regions are accurately represented, thereby enhancing the robustness and
precision of our spatial analysis.

\phantomsection\label{cell-fig-neighartfreg}
\begin{figure}[H]

\centering{

\includegraphics[width=0.7\textwidth,height=\textheight]{figures/nb_artfcountry.png}

}

\caption{\label{fig-neighartfreg}Neighbourhood structure of the
artificial region. This figure illustrates the spatial connectivity
between polygons in the artificial study region, where each line
represents a defined neighbour relationship based on spatial adjacency}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

In relation to our artificial study region,
Figure~\ref{fig-neighartfreg} illustrates the connections between areas,
represented by vertices or centroids for each area and nodes connecting
them to each other. This visualisation highlights the spatial adjacency
structure within the region.

\section{Spatial regression model}\label{spatial-regression-model}

When discussing spatial regression, it's crucial to comprehend the basic
notion of linear regression. In classical linear regression, the
relationship between the dependent variable \(y\) and the
\(x_1,x_2, \dots, x_p\) independent variables is expressed as

\begin{equation}\phantomsection\label{eq-basicreg}{
y = \beta_0 + \beta_1x_1 + \dots + \beta_p x_p + \varepsilon
}\end{equation}

This formulation assumes a global relationship between the variables,
where the coefficients \(\beta_1,\beta_2, \dots, \beta_p\) are constant
across the entire study area. In many spatial datasets, relationships
between variables may exhibit spatial variation. For example, in the
case of property pricing, a consistent rates of change assumption may
not hold true universally. For example, the house price increase for an
extra bedroom is often thought to be fixed across all locations.
However, local customs or knowledge may actually dictate these rates,
rather than a universal utility assigned to each commodity. For
instance, in neighbourhoods with families, where extra space is highly
valued, the perceived value of an additional bedroom may be greater
compared to areas with singles or elderly couples, for whom extra space
may not be as desirable.

\subsection{Geographically Weighted Regression
(GWR)}\label{geographically-weighted-regression-gwr}

Brunsdon, Fotheringham, and Charlton (1996) developed GWR which is one
such technique that extends classical linear regression by allowing
coefficients to vary spatially. It allows for the estimation of local
relationships between a response variable and predictor variables. It is
particularly useful for exploring spatial non-stationarity and
identifying spatially varying relationships across different locations.
The GWR model can be expressed as:

\begin{equation}\phantomsection\label{eq-gwr}{
y_i(s) = \beta_{0}(s) + \sum_{k=1}^p \beta_k(s)x_{ik}(s) + \varepsilon(s), \\ i = 1,\dots, n
}\end{equation}

Equation~\ref{eq-gwr} represents a spatial regression model where
\(\mathbf{y}(s)\) is the dependent variable at location \(s\). The term
\(\beta_{0}(s)\) is the spatially varying intercept, while
\(\sum_{k=1}^p \beta_k(s)\mathbf{x_k}(s)\) represents the spatially
varying coefficients for the independent variables \(\mathbf{x_k}(s)\).
\(\varepsilon(s)\) denotes the error term, capturing unexplained
variation at location \(s\)

Using a weighted least squares approach to calibrate regression models
allows different weights to be assigned to different observations,
influencing the estimated parameters. In ordinary least squares, the
goal is to minimize the sum of squared differences between predicted and
actual \(y\) values. Weighted least squares, however, apply a weighting
factor \(w\) to each squared difference, making some prediction
inaccuracies more significant. If \(w\) is a diagonal matrix of weights,
the estimated coefficients are given by Equation~\ref{eq-betagwr}:

\begin{equation}\phantomsection\label{eq-betagwr}{
\beta(s) = (\boldsymbol{x}^T \boldsymbol{W}(s) \boldsymbol{x})^{-1} \boldsymbol{x}^T \boldsymbol{W}(s) \boldsymbol{y}
}\end{equation}

This method allows GWR to address spatial heterogeneity by emphasizing
observations near the location of interest, thereby improving the
accuracy and relevance of local model estimates.

The estimation of GWR parameters involves fitting a separate regression
equation for each location in the study area. Various estimation
techniques can be employed, including ordinary least squares (OLS),
weighted least squares (WLS), and maximum likelihood estimation (MLE).
These techniques aim to optimize the model parameters to minimize the
differences between the observed and predicted values of the dependent
variable.

\subsection{Simultaneously Autoregressive (SAR)
Models}\label{simultaneously-autoregressive-sar-models}

While GWR focuses on capturing localised spatial heterogeneity by
allowing coefficients to vary across space, SAR take a different
approach by explicitly modelling spatial dependencies through a spatial
lag. The SAR model is a spatial econometric model used to analyse
spatial dependencies and relationships among observations in a
geographic space (Anselin and Griffith (1988)). It is widely employed in
various fields, such as regional economics, environmental studies, and
urban planning. SAR is a type of spatial autoregressive model involving
a simultaneous equation framework to capture spatial interactions.

The general form of SAR model can be expressed as follows

\begin{equation}\phantomsection\label{eq-sar}{
y(s) = \rho \sum_{s'} w(s,s') y(s') + \sum_{k=1}^p \beta_k x_k(s) + \varepsilon(s)
}\end{equation}

where \(\mathbf{Y}\) is the vector of observed values for the dependent
variable, \(W\) is the spatial weights matrix, \(\rho\) is the spatial
autoregressive coefficient, \(X\) is the matrix of observed values for
exogenous variables, \(\beta\) is the vector of coefficients, and
\(\boldsymbol{\varepsilon} \sim N(0,\sigma^2)\) is the vector of error
terms. Estimation of the SAR model parameters is typically done using
statistical techniques such as maximum likelihood estimation (MLE) or
generalised method of moments (GMM). The joint distribution of
\(\mathbf{Y}\) can be written as

\begin{equation}\phantomsection\label{eq-jointysar}{
\mathbf{Y} \sim \mathcal{N}\left( \left( I - \rho W \right)^{-1} X\beta, \, \sigma^2 \left( I - \rho W \right)^{-1} \left( I - \rho W^T \right)^{-1} \right)
}\end{equation}

Extensions of the SAR model include the Spatial Lag Model, the Spatial
Error Model, and the Spatial Durbin Model (Elhorst et al. (2014)), each
incorporating distinct assumptions regarding the spatial configuration
of the errors (Elhorst, Lacombe, and Piras (2012); Anselin (2013)).
These extensions provide flexibility to account for different types of
spatial dependencies and can be chosen based on the specific spatial
relationships and hypotheses under investigation.

The SAR model, with its various specifications, provides a flexible
framework to account for spatial dependencies and explore the spatial
dynamics of observed phenomena. Researchers often choose between these
models based on the nature of the spatial relationships in their data
and the specific hypotheses they aim to test. The SAR model is a
valuable tool for understanding spatial interdependence and making
informed policy and planning decisions in diverse spatial contexts

\subsection{Conditional Autoregressive (CAR)
Models}\label{conditional-autoregressive-car-models}

CAR models are a class of spatial statistical models used to analyze
spatially structured data. The general formulation of a CAR model can be
expressed as:

\begin{equation}\phantomsection\label{eq-car}{
y(s) = \sum_{k=1}^p \beta_k x_k(s) + \varepsilon(s) + \phi(s) 
}\end{equation}

Here, \(\boldsymbol{\varepsilon} \sim N(0, \nu^2)\) while
\(\boldsymbol{\phi}\) is a specific component in CAR model that has a
role as spatial effect. It is also common to mention
\(\boldsymbol{\phi}\) as a CAR priors. It is a type of Gaussian Markov
random field (Rue and Held (2005)), capture spatial autocorrelation by
ensuring that values at nearby locations are more similar than those
further apart (Lee (2013)). This can be expressed in a general term

\begin{equation}\phantomsection\label{eq-phidist}{
\boldsymbol{\phi}\sim N(0,\tau^2\boldsymbol{Q}^{-1}) 
}\end{equation}

where \(\boldsymbol{Q}\) is a precision matrix that may be singular
(intrinsic model). \(\boldsymbol{Q}\) controls the spatial
autocorrelation structure of the random effects, and is based on a
non-negative symmetric \(n \times n\) neighbourhood or weight matrix
\(\boldsymbol{W}\).

Together with the spatial weights matrix \(\boldsymbol{W}\), the prior
information are crucial components of CAR models. The choice of
\(\boldsymbol{W}\) determines the spatial structure of the model, while
the priors for the variance parameters and the spatial random effects
influence the model's ability to capture spatial dependencies. Commonly
used priors for the variance parameters include inverse-gamma
distributions, which provide flexibility and can be tuned to reflect
prior beliefs about the scale of variability in the data. The prevailing
approach typically involves a binary representation using geographical
adjacency, where \(w_{ki} = 1\) if areal units \((S_k,S_i)\) have a
mutual boundary (denoted \(k\sim i\), and is zero otherwise). This
specification forces \((\phi_k,\phi_i)\) relating to geographically
adjacent areas (that is \(w_{ki} = 1\)) to be correlated. On the other
hand, random effects associated with areas that are not adjacent are
independent of each other, provided we know the values of the other
random effects.

A CAR prior was introduced by Leroux, Lei, and Breslow (2000) and Stern
and Cressie (1999) to model diverse levels of spatial autocorrelation.
In this type of prior, a single collection of random effects is utilised
and its primary objective is to model spatial data, with a specific
focus on dealing with spatial relationships and auto-correlation between
data points. This model is especially proficient in the task of
smoothing data and detecting spatial patterns within data set. The
random effect \(\boldsymbol\phi_k\) structured as follow:

\begin{equation}\phantomsection\label{eq-phicar}{
\phi_k|\boldsymbol\phi_{-K}, \boldsymbol{W},\rho, \tau^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\tau^2}{A}\right)
}\end{equation}

where \(A = \rho\sum_{i=1}^Kw_{ki}+1-\rho\). Note that when
\(\rho = 1\), the prior forms an intrinsic CAR prior (Besag, York, and
Mollié 1991), indicating full spatial dependency. Conversely, when
\(\rho = 0\), \(A = 1\) it means that there will be no \(W\) matrix role
in the model, and it will become a comman random effect. In other words,
the model reduces to a generalised linear model.

When handling \(i\) observations within each area \(k\),
Equation~\ref{eq-ycar} closely resembles Equation~\ref{eq-gwr} in the
GWR model, with the exception of the \(\phi\) component. It is usually
called multilevel CAR models

\begin{equation}\phantomsection\label{eq-ycar}{
y_i(s) = \sum_{k=1}^p \beta_k x_{ki}(s) + \varepsilon_i(s) + \phi(s)
}\end{equation}

The Bayesian approach to CAR models entails defining prior distributions
for all model parameters, such as the regression coefficients \(\beta\),
variance parameters \(\nu^2\) and \(\tau^2\), and spatial random effects
\(\phi_k\). Subsequently, Bayesian inference techniques like Markov
Chain Monte Carlo (MCMC) are employed to derive posterior distributions
of these parameters. This method facilitates the integration of prior
knowledge and offers a versatile framework for quantifying uncertainty.

In general, Table~\ref{tbl-summarymethod} below summarises the key
elements utilised in the GWR, SAR, and CAR models:

\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}[t]{llll}

\caption{\label{tbl-summarymethod}Comparison of Model Elements Across
GWR, SAR, and CAR}

\tabularnewline

\toprule
Elements & $GWR$ & $SAR$ & $CAR$\\
\midrule
Number of Observations ($N$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Covariate Matrix ($X$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Spatial Weights Matrix ($W$) & $\times$ & $\checkmark$ & $\checkmark$\\
Spatial Coordinates (coords) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Covariate Coefficients ($\beta$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Error Term variance ($\sigma^2$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Spatial Autocorrelation ($\rho$) & $\times$ & $\checkmark$ & $\checkmark$\\
Spatial Structure component ($\phi$) & $\times$ & $\times$ & $\checkmark$\\
Spatial structure variance ($\tau^2$) & $\times$ & $\times$ & $\checkmark$\\
Additional scale ($\nu^2$) & $\times$ & $\times$ & $\checkmark$\\
\addlinespace
Estimation method & MLE & MLE/GMM & MCMC\\
\bottomrule

\end{longtable}

\endgroup{}

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

\section{Comparative Analysis of GLM, SAR, and
CAR}\label{comparative-analysis-of-glm-sar-and-car}

This section focuses on generating spatially autocorrelated property
price data, exploring two distinct data generation frameworks: one using
a CAR specification and another without spatial autocorrelation
(non-CAR). The analysis is conducted within an artificially constructed
study area using key property covariates: land size, building size,
number of bedrooms, and number of bathrooms. The generated datasets are
then employed to examine spatial dependencies in property prices and are
analysed using a suite of models including GLM, SAR, and CAR.

\subsection{The experiment design}\label{the-experiment-design}

The experiment begins by creating a dataset based on the defined study
region. The idea is to generate covariates that hypothetically can
explain price of a property. It basically consist of property structural
characteristics such as land size, building size, number of bedrooms,
and number of bathrooms. To represent real-world data, each polygon
contains multiple observations.

To generate the covariates, the first step involves creating area-level
data with a specific spatial pattern, referred to as the housing density
for the study region. The spatial pattern implies that certain areas or
polygons will have a higher housing density than others. In this case,
the central horizontal region of the study area is set to has a higher
housing density, mirroring the housing distribution found on the island
of Lombok. Although this density will not be directly utilized in the
model simulation, it serves as a critical foundation for determining the
spatial distribution of observation points across each polygon. In areas
with higher density, a greater number of observation points are
generated, which aligns with the notion that densely populated regions
typically have more residential developments, thereby increasing the
number of properties available for sale. Across the 216 polygons, a
total of 4,314 observation points were generated. The number of
observations per polygon varies from 5 to 56 points, reflecting the
density in each area. Figure~\ref{fig-pointsonmap} illustrate how each
polygon has multiple observations and they are distributed according to
a specific pattern.

\phantomsection\label{cell-fig-pointsonmap}
\begin{figure}[H]

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{figures/density_points_plot.png}

}

\caption{\label{fig-pointsonmap}Artificial study region with observation
points. The points represent data locations within each polygon,
simulating real-world spatial observations for the purpose of model
analysis and validation.}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

Following this, covariates were generated using \texttt{mvrnorm()} for
each observation point, creating a dataset for further analysis.
Table~\ref{tbl-covrstr} provides a brief slice of the covariates'
structure.

\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}[t]{rrrrr}

\caption{\label{tbl-covrstr}Slice of generated dataset as covariates for
simulations}

\tabularnewline

\toprule
ID & land\_size & building\_size & \#bedrooms & \#bathrooms\\
\midrule
1 & 270.48 & 108.55 & 4 & 2\\
1 & 295.99 & 100.13 & 4 & 2\\
2 & 319.52 & 126.31 & 3 & 1\\
2 & 325.01 & 120.84 & 4 & 3\\
2 & 294.15 & 144.05 & 4 & 3\\
\addlinespace
3 & 222.18 & 81.40 & 3 & 2\\
4 & 414.23 & 150.15 & 2 & 2\\
4 & 278.75 & 101.41 & 2 & 1\\
\bottomrule

\end{longtable}

\endgroup{}

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

By employing artificially generated covariates and an artificially
constructed study region, we developed three distinct property price
datasets, each based on a different spatial model: CAR, a combination of
CAR and SAR, and GWR. The true parameter values for this purpose are
\(\boldsymbol{\beta} = \begin{bmatrix} 1 & 0.9 & 0.7 & 0.5 & 0.3 \end{bmatrix}\),
\(\nu^2 = 0.2\), \(\tau^2 = 0.6\) and \(\rho = 0.6\). Whereas beta
values for GWR dataset is generated using a beta function, which is a
function of coordinate points. The matrix \(\mathbf{W}\) was derived
from the neighbourhood structure of an artificial study region using the
functions \texttt{poly2nb()} and \texttt{nb2mat()}. Data generation for
\(SAR + CAR\) employs a slightly different strategy. In addition to
being generated per area simultaneously using the joint distribution of
\(\mathbf{Y}\) as in Equation~\ref{eq-jointysar}, information about the
spatial effect \(\boldsymbol{\phi}\) is also added at the end of the
process. Therefore, we refer to it as \(SAR + CAR\) data.

Each data generation resulting two types of datasets: area-level and
point-level. These datasets were then modelled using six linear
regression approaches, comprising both classical and spatial regression
models: GLM, SAR, and CAR for area-level data, and GWR, GLMM, and
multilevel CAR for point-level data. Simulations were conducted
\(sim = 1000\) times. From each simulation, coefficients, parameters,
and fitted values were extracted, allowing for the computation of bias
and RMSE values for robustness and prediction power analysis.

\subsection{Robustness and power prediction
analysis}\label{robustness-and-power-prediction-analysis}

The bias values show that the models exhibit varying levels of accuracy
across the different parameters and datasets. The coefficients
\(\beta_k\) for \(k = 0, 1, 2, 3, 4\) represent the values associated
with the intercept, land size, building size, number of bedrooms, and
number of bathrooms, respectively. For \(\beta_0\), the GLM and CAR
models display relatively small biases with narrow standard deviations,
indicating better consistency, while the SAR model demonstrates larger
bias values, suggesting potential issues with accurately estimating the
intercept. For higher parameters like \(\beta_2\) and \(\beta_3\), SAR
generally has larger bias and variability, especially in the
\(SAR + CAR\) dataset, suggesting that the combined spatial structure
poses challenges for SAR's estimation. It is important to note that the
Figure~\ref{fig-biasarealevel} utilises area-level data. In this
context, the CAR model consistently demonstrates the smallest bias
compared to the other two models.

\phantomsection\label{cell-fig-biasarealevel}
\begin{figure}[H]

\centering{

\includegraphics[width=0.75\textwidth,height=\textheight]{figures/bias1b_plot.pdf}

}

\caption{\label{fig-biasarealevel}Bias values by method and dataset type
for area-level data. This figure compares the bias estimates of
different methods (CAR, GLM, SAR) across datasets generated under CAR,
GWR, and SAR+CAR frameworks. Each method's performance is visualised for
four coefficients (\(\beta_1\), \(\beta_2\), \(\beta_3\), \(\beta_4\)),
highlighting variations in bias and the effectiveness of each model in
addressing spatial dependencies.}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

\phantomsection\label{cell-fig-biaspointlevel}
\begin{figure}[H]

\centering{

\includegraphics[width=0.75\textwidth,height=\textheight]{figures/bias2_plot.pdf}

}

\caption{\label{fig-biaspointlevel}Bias values by method and dataset
type for point-level data. This figure compares the bias estimates of
different methods (GLMM, GWR, and mlvCAR) across datasets generated
under CAR, GWR, and SAR+CAR frameworks. Each method's performance is
visualised for four coefficients (\(\beta_1\), \(\beta_2\), \(\beta_3\),
\(\beta_4\)), highlighting variations in bias and the effectiveness of
each model in addressing spatial dependencies.}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

As shown in the second plot of Figure~\ref{fig-biaspointlevel}, this
analysis utilises point-level data. Overall, the mlvCAR model
demonstrates a consistently lower bias across most parameters and
scenarios. In the \(SAR + CAR\) scenario, where spatial complexity is
most pronounced, mlvCAR achieves notably low bias, particularly for
parameters \(\beta_1\) through \(\beta_4\), indicating its suitability
for handling mixed spatial structures. making it a solid choice for a
range of spatial modelling tasks.

The GWR model, which provides highly localised estimates, displays
slightly higher bias for several parameters compared to mlvCAR,
especially for \(\beta_1\), \(\beta_2\), and \(\beta_3\) within the
\(CAR\) scenario. This suggests GWR's greater sensitivity to spatial
patterns without incorporating neighbourhood-based effects. In contrast,
GLMM, which lacks explicit spatial components, shows larger bias
variability, particularly in scenarios with pronounced spatial
dependencies, highlighting its limitations for purely spatial analyses.

\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}[t]{lrrrrrr}

\caption{\label{tbl-rmse}RMSE values across different datasets}

\tabularnewline

\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Area-level data} & \multicolumn{3}{c}{Point-level data} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Dataset & $GLM$ & $SAR$ & $CAR$ & $GLMM$ & $GWR$ & $mlvCAR$\\
\midrule
\endfirsthead
\multicolumn{7}{@{}l}{\textit{(continued)}}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Area-level data} & \multicolumn{3}{c}{Point-level data} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Dataset & $GLM$ & $SAR$ & $CAR$ & $GLMM$ & $GWR$ & $mlvCAR$\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
$CAR$ & 4.09 & 3.85 & 1.81 & 14.04 & 21.13 & 9.94\\
$SAR + CAR$ & 11.88 & 11.68 & 3.88 & 47.67 & 48.87 & 38.97\\
$GWR$ & 2.43 & 2.42 & 2.28 & 18.31 & 18.33 & 14.87\\*

\end{longtable}

\endgroup{}

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

Table~\ref{tbl-rmse} provide insights into model accuracy at different
spatial levels. For area-level data, the models GLM, SAR, and CAR are
compared, with the CAR model consistently achieving the lowest RMSE.
This suggests that CAR is more effective at capturing spatial
dependencies at the area level. For point-level data, the models GLMM,
GWR, and mlvCAR are evaluated, with mlvCAR showing the lowest RMSE
across all datasets. This indicates that mlvCAR is particularly
well-suited for handling detailed spatial variations at the point level.
Overall, these results highlight the strengths of CAR and mlvCAR in
reducing error within their respective spatial contexts, underscoring
their potential utility in spatial modeling applications.

\section{Applications in Lombok house prices
dataset}\label{applications-in-lombok-house-prices-dataset}

In this section, we apply the simulated models to property price data on
Lombok Island, Indonesia. Lombok is located in the eastern part of
Indonesia, specifically between latitudes -8.9° and -8.1°, and
longitudes 115.9° and 116.6°.

\phantomsection\label{cell-fig-lombokmap}
\begin{figure}[H]

\centering{

\includegraphics[width=0.75\textwidth,height=\textheight]{figures/nb_lop2.png}

}

\caption{\label{fig-lombokmap}Lombok neighbour structure at the
sub-district (Kecamatan) level. This figure illustrates the spatial
neighbour structure of sub-districts on Lombok Island. The red lines
represent the connections between neighbouring sub-districts based on
adjacency relationships, overlaid on a base map.}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

Lombok includes five districts (kota/kabupaten), 53 sub-districts
(kecamatan), and 608 villages (desa/kelurahan), which represent three
successive levels of administrative division. In this study, the
sub-district level is used as the spatial unit of analysis, as
illustrated in Figure~\ref{fig-lombokmap}. The key variables of interest
include property price, land size, built-up area, number of bedrooms,
and number of bathrooms.

\subsection{Data Collection}\label{data-collection}

Data were collected from multiple online sources. Property prices and
their associated characteristics were harvested using web-scraping
techniques from three leading Indonesian property trading platforms,
https://www.lamudi.co.id/, https://www.99.co/, and
https://www.rumah123.com/. The accuracy of the web-scraped data was
ensured by cross-referencing it with official datasets, performing
checks for missing or inconsistent entries, and validating key variables
through sample comparisons with manually collected data. From this
process, the initial dataset comprised 1,188 entries, with 9 variables,
including village, prices, land-size, built-up area, number of bedrooms,
number of bathrooms, floors, property type/category, and furnishing
status.

Further, we conducted several preprocessing steps. First, we filtered
the dataset by removing irrelevant variables, including floors,
furnishing status, and category. It was then subsequently filtered to
retain only properties with plausible characteristics: land area between
90 and 800 square meters, built-up size between 70 and 600 square
meters, a maximum of 6 bedrooms, and a maximum 5 bathrooms. These
thresholds were applied to ensure the data reflect realistic and
context-appropriate housing characteristics in Lombok, based on common
residential patterns and local housing norms. This step was essential to
enhance the validity of the analysis by excluding outliers or
potentially erroneous entries. Following this procedure, the dataset was
refined to comprise a total of 576 observations. A summary of the key
variables is presented in Table~\ref{tbl-datasumm}.

\begingroup\fontsize{9}{11}\selectfont

\begin{longtable}[t]{lrrrrrrrr}

\caption{\label{tbl-datasumm}Summary statistics of variables in the
property price dataset}

\tabularnewline

\toprule
Variable & Mean & Median & $Q_1$ & $Q_3$ & Min & Max & SD & NA\\
\midrule
Prices (million IDR) & 1983.00 & 1400 & 800.0 & 2500.0 & 170 & 22500 & 2056.81 & 26\\
Land-size (sqm) & 299.60 & 255 & 147.2 & 407.2 & 90 & 800 & 179.39 & 24\\
Built-up area (sqm) & 184.80 & 150 & 100.0 & 210.0 & 70 & 600 & 109.20 & 24\\
No. of bedroom & 3.24 & 3 & 2.0 & 4.0 & 1 & 6 & 1.19 & 24\\
No. of bathroom & 2.55 & 2 & 2.0 & 3.0 & 1 & 5 & 1.05 & 24\\
\bottomrule

\end{longtable}

\endgroup{}

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

Next, we merged the cleaned dataset with spatial administrative data by
matching sub-district and district names. This process resulted in a
dataset containing 598 entries. Compared to the previous dataset, which
consisted of 576 observations, this indicates that 22 sub-districts did
not have any property data available---meaning that property sales in
those areas were not recorded on the online property listing platform
used as the data source.

To address missing values, we employed the \texttt{mice} package (Van
Buuren and Groothuis-Oudshoorn (2011)) to perform multiple imputation
with parameters set to \(maxit = 10\), \(m = 10\), and a fixed random
seed to ensure reproducibility. The imputed values were then
reintegrated into the main dataset, replacing the original missing
entries. This process resulted in a final dataset that was complete and
ready for subsequent analysis using the selected spatial models.

In addition to the point-level dataset (\texttt{ppd}), which contains
individual property listings, the imputed dataset was also aggregated to
the sub-district level by calculating the mean values of key variables.
This aggregation produced an area-level dataset (\texttt{ppd\_kc}),
allowing the analysis to be conducted at both the individual and
administrative levels. These two levels of data granularity provide
complementary perspectives for evaluating the performance of spatial
models in capturing local property market dynamics.

\subsection{Model Implementation \&
Analysis}\label{model-implementation-analysis}

The cleaned and completed Lombok dataset was then model. The tables
below present the results of the data fitting, offering a comparative
view of each model's effectiveness and accuracy in capturing the
dataset's characteristics.

\begin{equation}\phantomsection\label{eq-lopreg}{
prices \sim land_{std} + building_{std} + beds_{std} + baths_{std}
}\end{equation}

Standardization of the predictors ensures that the resulting
coefficients are on a comparable scale, facilitating meaningful
interpretation and comparison across different modeling approaches. The
results presented in the following tables provide a comprehensive
evaluation of each model's performance in capturing both the structural
and spatial characteristics of the property market in Lombok. This
comparison allows us to assess the extent to which spatially explicit
models (e.g., GWR, CAR, SAR) offer improvements over conventional
methods in explaining price variability across regions. --- Let me know
if you'd like to include metrics such as R², AIC, or Moran's I in the
paragraph as well.

We conducted model fitting on the area-level data, with detailed results
provided in the Appendix (Table~\ref{tbl-modelcomp2}). However, this
method does not provide optimal parameter estimates because data
aggregation can reduce variability and obscure spatial details essential
for capturing local effects. To address this, we used point-level data
to apply three different models: GLMM, GWR, and multilevel CAR.

\begin{table}

\caption{\label{tbl-modelcomp}Point-level Model Comparison}

\centering{

\centering\begingroup\fontsize{9}{11}\selectfont

\begin{tabular}[t]{llll}
\toprule
 & GLMM & GWR & mlvCAR\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Coefficients}}\\
\hspace{1em}Intercept & 7.09 [6.93, 7.24]* & 7.25 [7.25, 7.26]* & 7.08 [6.99, 7.17]*\\
\hspace{1em}Land-size & 0.26 [0.21, 0.32]* & 0.31 [0.30, 0.32]* & 0.26 [0.21, 0.32]*\\
\hspace{1em}Built-up area & 0.16 [0.10, 0.22]* & 0.19 [0.19, 0.20]* & 0.16 [0.10, 0.21]*\\
\hspace{1em}No. of bedroom & -0.03 [-0.09, 0.01] & -0.15 [-0.17, -0.15]* & -0.03 [-0.09, 0.02]\\
\hspace{1em}No. of bathroom & 0.19 [0.13, 0.25]* & 0.24 [0.23, 0.26]* & 0.19 [0.13, 0.25]*\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Spatial parameter}}\\
\hspace{1em}$\rho$ & - & - & 0.48 [0.07, 0.90]*\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Variances}}\\
\hspace{1em}$\nu^2$ & 0.22 [0.19, 0.25]* & 0.33 [0.29, 0.37]* & 0.22 [0.20, 0.25]*\\
\hspace{1em}$\tau^2$ & 0.20 [0.11, 0.34]* & - & 0.31 [0.14, 0.58]*\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Model fit criterion}}\\
\hspace{1em}AIC & 803.38 & -645.104 & 786.95\\
\hspace{1em}DIC & - & - & 850.39\\
\hspace{1em}WAIC & 838.3 & - & 851.14\\
\hspace{1em}LMPL & -421.31 & - & -427.21\\
\hspace{1em}Log-likelihood & -400.69 & -520.97 & -385.47\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

Here, Table~\ref{tbl-modelcomp} compares parameter estimates and fit
criteria for three models: GLMM, GWR, and multilevel CAR. Significant
coefficients are marked with an asterisk (*). Key determinants like
land-size and built-up area have positive, significant effects across
models, while the number of bedrooms shows only minor, varying effects,
indicating model-specific differences in capturing this variable's
influence. Only the multilevel CAR model includes the spatial parameter
\(\rho\), with a significant estimate of 0.48, showing the model's
ability to account for spatial dependence. Variance components \(\nu^2\)
(residual) and \(\tau^2\) (spatial) are also detailed, with both
variance terms significant. Fit criteria (AIC, WAIC, and Log-likelihood)
suggest mlvCAR and GWR may better capture local spatial variation, with
mlvCAR balancing spatial structure and precision.

\section{Results and Discussion}\label{results-and-discussion}

The findings from simulation consistently indicate that both the CAR and
mlvCAR model outperform the other models in capturing spatial
dependencies. This is evidenced by lower RMSE values and reduced bias
across various dataset scenarios, highlighting the CAR model's
robustness. In datasets with strong spatial structures, such as those
generated under \(CAR\) and \(SAR + CAR\) conditions, the CAR model
achieves substantially lower RMSE values compared to GLM and SAR,
suggesting its effectiveness in handling localized spatial variations.

In the context of house price modelling, it is common to encounter
multiple sales observations within a single area, yet precise
coordinates for each sale are often unavailable. Obtaining precise
geographic coordinates in real-world datasets is often challenging due
to limitations in data collection, privacy concerns, or the aggregation
of data into broader administrative units. This limitation renders
point-level approaches, such as GWR, less effective. With multiple
observations available for each area, models like mlvCAR offer a more
practical and accurate solution, effectively capturing spatial
dependencies at an area level without requiring precise point-level
locations.

The results underscore the importance of using models that explicitly
account for spatial dependencies in fields like property valuation,
where spatial heterogeneity plays a significant role. The superior
performance of the CAR model, particularly in datasets with strong
spatial structure, suggests that it may be a more reliable choice for
applications where local spatial variations are critical. The findings
align with previous studies indicating that CAR models are better suited
for managing spatial autocorrelation and heterogeneity, especially when
spatial dependencies are strong.

Additionally, the limitations of point-level models such as GWR in the
absence of precise location data highlight a practical challenge in
spatial modeling. When only area-level data is available, the use of
models like mlvCAR becomes essential, as it allows for spatial analysis
without the need for detailed coordinates, thus expanding the
applicability of spatial models in real-world settings. These insights
suggest potential for further exploration of the CAR and mlvCAR models
in various domains, especially in urban planning, real estate, and other
fields where spatial relationships impact outcomes. Future research
could examine the use of CAR and mlvCAR models with different spatial
resolutions or apply these models to other datasets with varying levels
of spatial dependency to further validate their effectiveness.

\section*{Supplementary information}\label{supplementary-information}
\addcontentsline{toc}{section}{Supplementary information}

Not applicable

\section*{Acknowledgments}\label{acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

Not applicable

\section*{Declarations}\label{declarations}
\addcontentsline{toc}{section}{Declarations}

\begin{itemize}
\tightlist
\item
  Funding
\item
  Competing interests
\end{itemize}

No, I declare that the authors have no competing interests as defined by
Springer, or other interests that might be perceived to influence the
results and/or discussion reported in this paper.

\begin{itemize}
\tightlist
\item
  Ethics approval
\item
  Consent to participate
\item
  Consent for publication
\item
  Availability of data and materials
\item
  Code availability
\item
  Authors' contributions
\end{itemize}

::: \{.appendices latex=true\}

\section{Simple Feature Data Slice}\label{secA1}

\begin{longtable}[t]{rl}

\caption{\label{tbl-geoartfex}a slice of a simple feature collection
with 6 polygon features and 1 attribute field, area(sqm), representing
polygon sizes in square units. The dataset, using the WGS 84 CRS, spans
longitudes 116.0355--116.5 and latitudes -8.500079---8.000057. The
geometry field defines each polygon's spatial boundaries, representing
artificial study regions for spatial modeling in this research}

\tabularnewline

\toprule
area & geometry\\
\midrule
17616519 & POLYGON ((116.2296 -8.50007...\\
11485356 & POLYGON ((116.0778 -8.32479...\\
4755629 & POLYGON ((116.399 -8.380894...\\
15959450 & POLYGON ((116.3567 -8.00006...\\
12727929 & POLYGON ((116.14 -8.292773,...\\
\addlinespace
8848979 & POLYGON ((116.5 -8.210459, ...\\
7248031 & POLYGON ((116.4308 -8.45091...\\
16906811 & POLYGON ((116.1352 -8.26696...\\
6346890 & POLYGON ((116.0723 -8.30335...\\
7620652 & POLYGON ((116.4541 -8.43327...\\
\addlinespace
10987363 & POLYGON ((116.4982 -8.06795...\\
10411472 & POLYGON ((116.1071 -8.07289...\\
20452118 & POLYGON ((116.4233 -8.12744...\\
19613075 & POLYGON ((116 -8.5, 116 -8....\\
9252319 & POLYGON ((116.4541 -8.43327...\\
\bottomrule

\end{longtable}

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

\section{Bias values for Area-level Data}\label{secA2}

\begin{table}

\caption{\label{tbl-bias1}Mean ± SD values for bias across different
area-level datasets and parameters. This table presents the mean and
standard deviation (SD) of bias values across area-level datasets (CAR,
SAR + CAR, GWR) and spatial models (GLM, SAR, CAR), with
\(nsim = 1000\). This is also act as numerical representation of fig.~5
(top panel)}

\centering{

[!h]
\centering\begingroup\fontsize{9}{11}\selectfont

\begin{tabular}[t]{lccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Mean ± SD} \\
\cmidrule(l{3pt}r{3pt}){2-4}
Dataset & GLM & SAR & CAR\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{1}$}}\\
\hspace{1em}CAR & $7.80e-02 \pm 0.1900$ & $9.00e-03 \pm 0.1900$ & $-8.70e-02 \pm 0.1800$\\
\hspace{1em}SAR + CAR & $3.30e-01 \pm 0.3800$ & $2.77e-01 \pm 0.3700$ & $2.10e-01 \pm 0.3500$\\
\hspace{1em}GWR & $-6.00e-02 \pm 0.1100$ & $-6.50e-02 \pm 0.1100$ & $-6.50e-02 \pm 0.1100$\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{2}$}}\\
\hspace{1em}CAR & $2.19e-01 \pm 0.1500$ & $2.37e-01 \pm 0.1400$ & $2.46e-01 \pm 0.1400$\\
\hspace{1em}SAR + CAR & $-2.47e-01 \pm 0.3900$ & $-2.29e-01 \pm 0.3600$ & $-2.12e-01 \pm 0.3700$\\
\hspace{1em}GWR & $3.05e-01 \pm 0.0900$ & $3.02e-01 \pm 0.0900$ & $3.04e-01 \pm 0.0900$\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{3}$}}\\
\hspace{1em}CAR & $-1.24e-01 \pm 0.2000$ & $0.00e+00 \pm 0.1900$ & $-3.80e-02 \pm 0.1800$\\
\hspace{1em}SAR + CAR & $-5.80e-02 \pm 0.3700$ & $6.30e-02 \pm 0.3500$ & $-5.00e-03 \pm 0.3500$\\
\hspace{1em}GWR & $-6.20e-02 \pm 0.1100$ & $-3.50e-02 \pm 0.1100$ & $-5.90e-02 \pm 0.1100$\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{4}$}}\\
\hspace{1em}CAR & $1.83e-01 \pm 0.1600$ & $1.27e-01 \pm 0.1500$ & $1.64e-01 \pm 0.1500$\\
\hspace{1em}SAR + CAR & $1.22e-01 \pm 0.3000$ & $7.60e-02 \pm 0.2900$ & $1.10e-01 \pm 0.2800$\\
\hspace{1em}GWR & $1.82e-01 \pm 0.0900$ & $1.73e-01 \pm 0.0900$ & $1.83e-01 \pm 0.0900$\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

\begin{table}

\caption{\label{tbl-bias2}Mean ± SD values for bias across different
point-level datasets and parameters. This table presents the mean and
standard deviation (SD) of bias values across point-level datasets (CAR,
SAR + CAR, GWR) and spatial models (GLM, SAR, CAR), with
\(nsim = 1000\). This is also act as numerical representation of fig.~5
(bottom panel)}

\centering{

[!h]
\centering\begingroup\fontsize{9}{11}\selectfont

\begin{tabular}[t]{lccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Mean ± SD} \\
\cmidrule(l{3pt}r{3pt}){2-4}
Dataset & GLMM & GWR & mlvCAR\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{1}$}}\\
\hspace{1em}CAR & $2.80e-04 \pm 0.0115$ & $8.77e-03 \pm 0.0160$ & $-2.00e-05 \pm 0.0114$\\
\hspace{1em}SAR + CAR & $4.06e-03 \pm 0.0525$ & $9.00e-03 \pm 0.0538$ & $9.80e-04 \pm 0.0524$\\
\hspace{1em}GWR & $-5.00e-04 \pm 0.0109$ & $-5.20e-04 \pm 0.0108$ & $-5.40e-04 \pm 0.0108$\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{2}$}}\\
\hspace{1em}CAR & $-3.10e-04 \pm 0.0125$ & $-5.50e-03 \pm 0.0172$ & $-1.70e-04 \pm 0.0124$\\
\hspace{1em}SAR + CAR & $-2.81e-03 \pm 0.0588$ & $-5.72e-03 \pm 0.0602$ & $-1.02e-03 \pm 0.0585$\\
\hspace{1em}GWR & $9.54e-03 \pm 0.0113$ & $9.55e-03 \pm 0.0113$ & $9.56e-03 \pm 0.0113$\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{3}$}}\\
\hspace{1em}CAR & $3.80e-04 \pm 0.0096$ & $-2.86e-03 \pm 0.0142$ & $4.50e-04 \pm 0.0095$\\
\hspace{1em}SAR + CAR & $9.00e-04 \pm 0.0391$ & $-8.00e-04 \pm 0.0405$ & $1.67e-03 \pm 0.0389$\\
\hspace{1em}GWR & $-4.90e-04 \pm 0.0092$ & $-5.00e-04 \pm 0.0091$ & $-5.20e-04 \pm 0.0091$\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{$\beta_{4}$}}\\
\hspace{1em}CAR & $-4.30e-04 \pm 0.0089$ & $1.18e-03 \pm 0.0122$ & $-5.10e-04 \pm 0.0089$\\
\hspace{1em}SAR + CAR & $3.10e-04 \pm 0.0393$ & $1.03e-03 \pm 0.0394$ & $-8.00e-05 \pm 0.0391$\\
\hspace{1em}GWR & $1.09e-03 \pm 0.0089$ & $1.05e-03 \pm 0.0088$ & $1.07e-03 \pm 0.0087$\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

\section{Model Comparison for area-level data}\label{secA3}

\begin{table}

\caption{\label{tbl-modelcomp2}This table compares the performance of
GLM, SAR, and CAR models on area-level data. It reports parameter
estimates (with 95\% confidence intervals), spatial parameters
(\(\rho\)), variances (\(\nu^2\), \(\tau^2\)), and model fit criteria
(AIC, DIC, WAIC, LMPL, log-likelihood). The results highlight each
model's capability to capture spatial dependencies and variability}

\centering{

[!h]
\centering\begingroup\fontsize{9}{11}\selectfont

\begin{tabular}[t]{llll}
\toprule
 & GLM & SAR & CAR\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Coefficients}}\\
\hspace{1em}Intercept & 7.19 [7.02, 7.35]* & 6.92 [6.48, 7.37]* & 7.19 [7.03, 7.35]*\\
\hspace{1em}Land-size & 0.45 [0.24, 0.66]* & 0.45 [0.25, 0.64]* & 0.45 [0.23, 0.65]*\\
\hspace{1em}Built-up area & 0.01 [-0.20, 0.23] & 0.01 [-0.18, 0.22] & 0.02 [-0.20, 0.23]\\
\hspace{1em}No. of bedroom & -0.20 [-0.43, 0.02] & -0.23 [-0.45, -0.01]* & -0.21 [-0.44, 0.02]\\
\hspace{1em}No. of bathroom & 0.54 [0.33, 0.76]* & 0.56 [0.36, 0.77]* & 0.55 [0.34, 0.76]*\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Spatial parameter}}\\
\hspace{1em}$\rho$ & - & 0.009[-0.005, 0.023] & 0.39 [0.01, 0.91]*\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Variances}}\\
\hspace{1em}$\nu^2$ & 0.31 [0.21, 0.45]* & 0.27 [0.19, 0.41]* & 0.28 [0.01, 0.44]*\\
\hspace{1em}$\tau^2$ & 0.20 [0.11, 0.34]* & - & 0.04 [0.01, 0.91]*\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Model fit criterion}}\\
\hspace{1em}AIC & - & - & -\\
\hspace{1em}DIC & 94.73 & 94.73 & 82.52\\
\hspace{1em}WAIC & 95.64 & - & 87.43\\
\hspace{1em}LMPL & -48.04 & - & -46.71\\
\hspace{1em}Log-likelihood & -41.49 & -41.36 & -37.29\\
\bottomrule
\end{tabular}
\endgroup{}

}

\end{table}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

\section{Example of beta-values in GWR model}\label{secA4}

\phantomsection\label{cell-fig-betaplot}
\begin{figure}[H]

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{figures/betavalues_plot.png}

}

\caption{\label{fig-betaplot}Spatial variation of \(\beta\) values
generated using a custom beta function. The color gradient represents
the magnitude of \(\beta\) values across the artificial study region,
highlighting localised patterns and heterogeneity. The visualisation
demonstrates the spatially varying coefficient structure modeled in the
study}

\end{figure}%

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-anselin2013spatial}
Anselin, Luc. 2013. \emph{Spatial Econometrics: Methods and Models}.
Vol. 4. Springer Science \& Business Media.

\bibitem[\citeproctext]{ref-anselin1988spatial}
Anselin, Luc, and Daniel A Griffith. 1988. {``Do Spatial Effecfs Really
Matter in Regression Analysis?''} \emph{Papers in Regional Science} 65
(1): 11--34.

\bibitem[\citeproctext]{ref-banerjee2014hierarchical}
Banerjee, Sudipto, Bradley P Carlin, and Alan E Gelfand. 2014.
\emph{Hierarchical Modeling and Analysis for Spatial Data}. CRC press.

\bibitem[\citeproctext]{ref-begueria2009comparison}
Beguería, Santiago, and Yolanda Pueyo. 2009. {``A Comparison of
Simultaneous Autoregressive and Generalized Least Squares Models for
Dealing with Spatial Autocorrelation.''} \emph{Global Ecology and
Biogeography} 18 (3): 273--79.

\bibitem[\citeproctext]{ref-besag1991bayesian}
Besag, Julian, Jeremy York, and Annie Mollié. 1991. {``Bayesian Image
Restoration, with Two Applications in Spatial Statistics.''}
\emph{Annals of the Institute of Statistical Mathematics} 43 (March):
1--20. \url{https://doi.org/10.1007/BF00116466}.

\bibitem[\citeproctext]{ref-bottero2017buildings}
Bottero, Marta, Marina Bravi, Giulio Mondini, and Antonio Talarico.
2017. {``Buildings Energy Performance and Real Estate Market Value: An
Application of the Spatial Auto Regressive (SAR) Model.''}
\emph{Appraisal: From Theory to Practice: Results of SIEV 2015},
221--30.

\bibitem[\citeproctext]{ref-brunsdon1996geographically}
Brunsdon, Chris, A Stewart Fotheringham, and Martin E Charlton. 1996.
{``Geographically Weighted Regression: A Method for Exploring Spatial
Nonstationarity.''} \emph{Geographical Analysis} 28 (4): 281--98.

\bibitem[\citeproctext]{ref-cellmer2020spatial}
Cellmer, Radosław, Aneta Cichulska, and Mirosław Bełej. 2020. {``Spatial
Analysis of Housing Prices and Market Activity with the Geographically
Weighted Regression.''} \emph{ISPRS International Journal of
Geo-Information} 9 (6): 380. \url{https://doi.org/10.3390/ijgi9060380}.

\bibitem[\citeproctext]{ref-cellmer2019application}
Cellmer, Radosław, Katarzyna Kobylińska, and Mirosław Bełej. 2019.
{``Application of Hierarchical Spatial Autoregressive Models to Develop
Land Value Maps in Urbanized Areas.''} \emph{ISPRS International Journal
of Geo-Information} 8 (4): 195.

\bibitem[\citeproctext]{ref-de2012bayesian}
De Oliveira, Victor. 2012. {``Bayesian Analysis of Conditional
Autoregressive Models.''} \emph{Annals of the Institute of Statistical
Mathematics} 64: 107--33.

\bibitem[\citeproctext]{ref-droj2024comprehensive}
Droj, Gabriela, Anita Kwartnik-Pruc, and Laurențiu Droj. 2024. {``A
Comprehensive Overview Regarding the Impact of GIS on Property
Valuation.''} \emph{ISPRS International Journal of Geo-Information} 13
(6): 175.

\bibitem[\citeproctext]{ref-elhorst2014spatial}
Elhorst, J Paul et al. 2014. \emph{Spatial Econometrics: From
Cross-Sectional Data to Spatial Panels}. Vol. 479. Springer.

\bibitem[\citeproctext]{ref-elhorst2012model}
Elhorst, J Paul, Donald J Lacombe, and Gianfranco Piras. 2012. {``On
Model Specification and Parameter Space Definitions in Higher Order
Spatial Econometric Models.''} \emph{Regional Science and Urban
Economics} 42 (1-2): 211--20.

\bibitem[\citeproctext]{ref-fix2021simultaneous}
Fix, Miranda J, Daniel S Cooley, and Emeric Thibaud. 2021.
{``Simultaneous Autoregressive Models for Spatial Extremes.''}
\emph{Environmetrics} 32 (2): e2656.

\bibitem[\citeproctext]{ref-golgher2016interpret}
Golgher, André Braz, and Paul R Voss. 2016. {``How to Interpret the
Coefficients of Spatial Models: Spillovers, Direct and Indirect
Effects.''} \emph{Spatial Demography} 4: 175--205.

\bibitem[\citeproctext]{ref-lee2013carbayes}
Lee, Duncan. 2013. {``CARBayes: An r Package for Bayesian Spatial
Modeling with Conditional Autoregressive Priors.''} \emph{Journal of
Statistical Software} 55 (13): 1--24.
\url{https://doi.org/10.18637/jss.v055.i13}.

\bibitem[\citeproctext]{ref-leroux2000estimation}
Leroux, Brian G, Xingye Lei, and Norman Breslow. 2000. {``Estimation of
Disease Rates in Small Areas: A New Mixed Model for Spatial
Dependence.''} In \emph{Statistical Models in Epidemiology, the
Environment, and Clinical Trials}, 179--91.

\bibitem[\citeproctext]{ref-LeSage2014}
LeSage, James P., and R. Kelley Pace. 2014. {``Interpreting Spatial
Econometric Models.''} In \emph{Handbook of Regional Science}, edited by
Manfred M. Fischer and Peter Nijkamp, 1535--52. Berlin, Heidelberg:
Springer Berlin Heidelberg.
\url{https://doi.org/10.1007/978-3-642-23430-9_91}.

\bibitem[\citeproctext]{ref-lu2011geographically}
Lu, Binbin, Martin Charlton, and A Stewart Fotheringham. 2011.
{``Geographically Weighted Regression Using a Non-Euclidean Distance
Metric with a Study on London House Price Data.''} \emph{Procedia
Environmental Sciences} 7: 92--97.

\bibitem[\citeproctext]{ref-mccord2014understanding}
McCord, Michael, Peadar T Davis, Martin Haran, David McIlhatton, and
John McCord. 2014. {``Understanding Rental Prices in the UK: A
Comparative Application of Spatial Modelling Approaches.''}
\emph{International Journal of Housing Markets and Analysis} 7 (1):
98--128.

\bibitem[\citeproctext]{ref-morris2019bayesian}
Morris, Mitzi, Katherine Wheeler-Martin, Dan Simpson, Stephen J Mooney,
Andrew Gelman, and Charles DiMaggio. 2019. {``Bayesian Hierarchical
Spatial Models: Implementing the Besag York Molli{é} Model in Stan.''}
\emph{Spatial and Spatio-Temporal Epidemiology} 31 (November): 100301.
\url{https://doi.org/10.1016/j.sste.2019.100301}.

\bibitem[\citeproctext]{ref-pagourtzi2003real}
Pagourtzi, Elli, Vassilis Assimakopoulos, Thomas Hatzichristos, and Nick
French. 2003. {``Real Estate Appraisal: A Review of Valuation
Methods.''} \emph{Journal of Property Investment \& Finance} 21 (4):
383--401.

\bibitem[\citeproctext]{ref-rue2005gaussian}
Rue, Havard, and Leonhard Held. 2005. \emph{Gaussian Markov Random
Fields: Theory and Applications}. Chapman; Hall/CRC.

\bibitem[\citeproctext]{ref-sarlas2015localized}
Sarlas, Georgios, and Kay W Axhausen. 2015. {``Localized Speed
Prediction with the Use of Spatial Simultaneous Autoregressive
Models.''} In \emph{TRB 94th Annual Meeting Compendium of Papers}.
National Academy of Sciences.

\bibitem[\citeproctext]{ref-sisman2022modelling}
Sisman, Suleyman, and Arif Cagdas Aydinoglu. 2022. {``A Modelling
Approach with Geographically Weighted Regression Methods for Determining
Geographic Variation and Influencing Factors in Housing Price: A Case in
Istanbul.''} \emph{Land Use Policy} 119 (August): 106183.
\url{https://doi.org/10.1016/j.landusepol.2022.106183}.

\bibitem[\citeproctext]{ref-soltani2021housing}
Soltani, Ali, Christopher James Pettit, Mohammad Heydari, and Fatemeh
Aghaei. 2021. {``Housing Price Variations Using Spatio-Temporal Data
Mining Techniques.''} \emph{Journal of Housing and the Built
Environment} 36 (3): 1--29.
\url{https://doi.org/10.1007/s10901-020-09811-y}.

\bibitem[\citeproctext]{ref-stern1999inference}
Stern, Hal, and Noel A Cressie. 1999. {``Inference for Extremes in
Disease Mapping.''} \emph{John Wiley and Sons} 11: 63--84.

\bibitem[\citeproctext]{ref-stewart2018localized}
Stewart Fotheringham, A, and Bumsub Park. 2018. {``Localized
Spatiotemporal Effects in the Determinants of Property Prices: A Case
Study of Seoul.''} \emph{Applied Spatial Analysis and Policy} 11
(August): 581--98. \url{https://doi.org/10.1007/s12061-017-9232-8}.

\bibitem[\citeproctext]{ref-trojanek2018spatial}
Trojanek, Radoslaw, and Michal Gluszak. 2018. {``Spatial and Time Effect
of Subway on Property Prices.''} \emph{Journal of Housing and the Built
Environment} 33 (October): 359--84.
\url{https://doi.org/10.1007/s10901-017-9569-y}.

\bibitem[\citeproctext]{ref-van2011mice}
Van Buuren, Stef, and Karin Groothuis-Oudshoorn. 2011. {``Mice:
Multivariate Imputation by Chained Equations in r.''} \emph{Journal of
Statistical Software} 45: 1--67.

\bibitem[\citeproctext]{ref-ver2018spatial}
Ver Hoef, Jay M, Erin E Peterson, Mevin B Hooten, Ephraim M Hanks, and
Marie-Josèe Fortin. 2018. {``Spatial Autoregressive Models for
Statistical Inference from Ecological Data.''} \emph{Ecological
Monographs} 88 (1): 36--59.

\bibitem[\citeproctext]{ref-wall2004close}
Wall, Melanie M. 2004. {``A Close Look at the Spatial Structure Implied
by the CAR and SAR Models.''} \emph{Journal of Statistical Planning and
Inference} 121 (2): 311--24.

\bibitem[\citeproctext]{ref-wang2020geographically}
Wang, Chih-Hao, and Na Chen. 2020. {``A Geographically Weighted
Regression Approach to Investigating Local Built-Environment Effects on
Home Prices in the Housing Downturn, Recovery, and Subsequent
Increases.''} \emph{Journal of Housing and the Built Environment} 35:
1283--1302.

\bibitem[\citeproctext]{ref-yang2019does}
Yang, Linchuan, Jiangping Zhou, Oliver F Shyr, et al. 2019. {``Does Bus
Accessibility Affect Property Prices?''} \emph{Cities} 84 (January):
56--65. \url{https://doi.org/10.1016/j.cities.2018.07.005}.

\bibitem[\citeproctext]{ref-yu2007modeling}
Yu, Danlin. 2007. {``Modeling Owner-Occupied Single-Family House Values
in the City of Milwaukee: A Geographically Weighted Regression
Approach.''} \emph{GIScience \& Remote Sensing} 44 (3): 267--82.

\end{CSLReferences}

\end{document}

### 2025-05-05 docs/_tex/manuscript.tex
update

@@
 % author setup
-\author[1,2]{\fnm{Indira Puteri} \sur{Kinasih}}\email{indiraputeri@uinmataram.ac.id}\author*[1]{\fnm{Haziq} \sur{Jamil}}\email{haziq.jamil@ubd.edu.bn}\author*[1]{\fnm{Elvynna} \sur{Leong}}\email{elvynna.leong@ubd.edu.bn}
+\author*[1,2]{\fnm{Indira Puteri} \sur{Kinasih}}\email{indiraputeri@uinmataram.ac.id}\author[1]{\fnm{Haziq} \sur{Jamil}}\email{haziq.jamil@ubd.edu.bn}\author[1]{\fnm{Elvynna} \sur{Leong}}\email{elvynna.leong@ubd.edu.bn}
 % affil setup
@@
 indicating their neighbourhood relationships (Morris et al. (2019)). To
-demonstrate this concept, let's take a simple example of a map with four
-regions \((R_1, R_2, R_3, R_4)\), as visualised in
+demonstrate this concept, let's take a simple example of a map with 5
+regions \((A, B, C, D, E)\), as visualised in
 Figure~\ref{fig-neighsimple}.
@@
 
-\includegraphics[width=0.7\textwidth,height=\textheight]{figures/maprec.png}
+\includegraphics[width=0.5\textwidth,height=\textheight]{figures/sartf_plot.png}
 
@@
 
-This straightforward neighbourhood structure creates a single connected
-component with the following relationships:
-\((R_1 \sim R_2,\ R_2 \sim R_3,\ R_3 \sim R_4)\). Therefore the weight
-matrix \(\boldsymbol{W}\) is
+From the Figure~\ref{fig-neighsimple} we can derive an adjacency matrix
+which represents the neighborhood structure of five spatial units
+labeled \(A\) to \(E\). Each entry 1 indicates a direct adjacency (areas
+share a common border), while 0 denotes no direct connection. For
+example, Area \(A\) is adjacent to all other areas (\(B\), \(C\), \(D\),
+and \(E\)), whereas Area \(B\) is only adjacent to \(A\) and \(C\). This
+neighbourhood configuration captures a clear spatial interaction pattern
+among the areas, which forms the basis for constructing a spatial weight
+matrix, commonly denoted as the \(\mathbf{W}\) matrix
 
 \[
-\begin{array}{ccccc}
-    & R_1 & R_2 & R_3 & R_4 \\
-R_1 & 0 & 1 & 0 & 0\\ 
-R_2 & 1 & 0 & 1 & 0\\ 
-R_3 & 0 & 1 & 0 & 1\\
-R_4 & 0 & 0 & 1 & 0\\
+\begin{array}{cccccc}
+    & A & B & C & D & E \\
+A & 0 & 1 & 1 & 1 & 1\\ 
+B & 1 & 0 & 1 & 0 & 0\\ 
+C & 1 & 1 & 0 & 1 & 0\\
+D & 1 & 0 & 1 & 0 & 1\\
+E & 1 & 0 & 0 & 1 & 0\\
 \end{array}
@@
 each diagonal element \(d_{ii}\) denotes the number of neighbours of
-region \(R_i\), while all non-diagonal elements are zero. This matrix is
+region. while all non-diagonal elements are zero. This matrix is
 essential in spatial econometric models because it captures the local
@@
 \[
-\begin{array}{ccccc}
-   & R_1 & R_2 & R_3 & R_4 \\
-R_1 & 1 & 0 & 0 & 0\\ 
-R_2 & 0 & 2 & 0 & 0\\ 
-R_3 & 0 & 0 & 2 & 0\\
-R_4 & 0 & 0 & 0 & 1\\
+\begin{array}{cccccc}
+   & A & B & C & D & E \\
+A & 4 & 0 & 0 & 0 & 0\\ 
+B & 0 & 2 & 0 & 0 & 0\\ 
+C & 0 & 0 & 3 & 0 & 0\\
+D & 0 & 0 & 0 & 3 & 0\\
+E & 0 & 0 & 0 & 0 & 2\\
 \end{array}
@@
 
-This matrix format indicates that region \(R_1\) has one neighbour,
-regions \(R_2\) and \(R_3\) each have two neighbours, and region \(R_4\)
-has one neighbour. By integrating this matrix into our modelling
-framework, we ensure that the spatial structure and interactions among
-regions are accurately represented, thereby enhancing the robustness and
-precision of our spatial analysis.
+This matrix format indicates that region \(A\) has 4 neighbour, regions
+\(C\) and \(D\) each have 3 neighbours, and region \(B\) and \(E\) has 2
+neighbour.
 
@@
 
-The bias values show that the models exhibit varying levels of accuracy
-across the different parameters and datasets. The coefficients
-\(\beta_k\) for \(k = 0, 1, 2, 3, 4\) represent the values associated
-with the intercept, land size, building size, number of bedrooms, and
-number of bathrooms, respectively. For \(\beta_0\), the GLM and CAR
-models display relatively small biases with narrow standard deviations,
-indicating better consistency, while the SAR model demonstrates larger
-bias values, suggesting potential issues with accurately estimating the
-intercept. For higher parameters like \(\beta_2\) and \(\beta_3\), SAR
-generally has larger bias and variability, especially in the
-\(SAR + CAR\) dataset, suggesting that the combined spatial structure
-poses challenges for SAR's estimation. It is important to note that the
-Figure~\ref{fig-biasarealevel} utilises area-level data. In this
-context, the CAR model consistently demonstrates the smallest bias
-compared to the other two models.
-
-\phantomsection\label{cell-fig-biasarealevel}
-\begin{figure}[H]
+The bias values are summarised in Table~\ref{tbl-bias1} and
+Table~\ref{tbl-bias2}. It shows that the models exhibit varying levels
+of accuracy across the different parameters and datasets.
+Figure~\ref{fig-biasplot} presents the bias estimates of regression
+coefficients for different modelling approaches, separated by data type:
+(a) area-level data and (b) point-level data. Each panel corresponds to
+a data-generating mechanism (CAR, GWR, or SAR+CAR), and bias values are
+shown with 95\% confidence intervals.
+
+\begin{figure}
+
+\begin{minipage}{0.50\linewidth}
 
@@
 
-\includegraphics[width=0.75\textwidth,height=\textheight]{figures/bias1b_plot.pdf}
+\includegraphics{figures/bias1_plot.png}
 
@@
 
-\caption{\label{fig-biasarealevel}Bias values by method and dataset type
-for area-level data. This figure compares the bias estimates of
-different methods (CAR, GLM, SAR) across datasets generated under CAR,
-GWR, and SAR+CAR frameworks. Each method's performance is visualised for
-four coefficients (\(\beta_1\), \(\beta_2\), \(\beta_3\), \(\beta_4\)),
-highlighting variations in bias and the effectiveness of each model in
-addressing spatial dependencies.}
+\subcaption{\label{fig-biasplot1}Bias values by method and dataset type
+for area-level data}
+
+\end{minipage}%
+%
+\begin{minipage}{0.50\linewidth}
+
+\centering{
+
+\includegraphics{figures/bias2_plot.png}
+
+}
+
+\subcaption{\label{fig-biasplot2}Bias values by method and dataset type
+for point-level data}
+
+\end{minipage}%
+
+\caption{\label{fig-biasplot}Bias estimates comparison of six different
+methods across datasets generated under CAR, GWR, and SAR+CAR
+frameworks. Each method's performance is visualised for \(\beta_1\),
+\(\beta_2\), \(\beta_3\), and \(\beta_4\), highlighting variations in
+bias and the effectiveness of each model in addressing spatial
+dependencies.}
 
@@
 
-\phantomsection\label{cell-fig-biaspointlevel}
-\begin{figure}[H]
-
-\centering{
-
-\includegraphics[width=0.75\textwidth,height=\textheight]{figures/bias2_plot.pdf}
-
-}
-
-\caption{\label{fig-biaspointlevel}Bias values by method and dataset
-type for point-level data. This figure compares the bias estimates of
-different methods (GLMM, GWR, and mlvCAR) across datasets generated
-under CAR, GWR, and SAR+CAR frameworks. Each method's performance is
-visualised for four coefficients (\(\beta_1\), \(\beta_2\), \(\beta_3\),
-\(\beta_4\)), highlighting variations in bias and the effectiveness of
-each model in addressing spatial dependencies.}
-
-\end{figure}%
-
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
-As shown in the second plot of Figure~\ref{fig-biaspointlevel}, this
-analysis utilises point-level data. Overall, the mlvCAR model
-demonstrates a consistently lower bias across most parameters and
-scenarios. In the \(SAR + CAR\) scenario, where spatial complexity is
-most pronounced, mlvCAR achieves notably low bias, particularly for
-parameters \(\beta_1\) through \(\beta_4\), indicating its suitability
-for handling mixed spatial structures. making it a solid choice for a
-range of spatial modelling tasks.
-
-The GWR model, which provides highly localised estimates, displays
-slightly higher bias for several parameters compared to mlvCAR,
-especially for \(\beta_1\), \(\beta_2\), and \(\beta_3\) within the
-\(CAR\) scenario. This suggests GWR's greater sensitivity to spatial
-patterns without incorporating neighbourhood-based effects. In contrast,
-GLMM, which lacks explicit spatial components, shows larger bias
-variability, particularly in scenarios with pronounced spatial
-dependencies, highlighting its limitations for purely spatial analyses.
+For area-level data (Figure~\ref{fig-biasplot1}), CAR model consistently
+produces the lowest bias across all coefficients and data-generating
+scenarios. This confirms its ability to recover true parameter values
+under spatial dependency, even when data is aggregated. The SAR model
+displays higher variability and often substantial bias, especially in
+CAR-generated datasets, suggesting model misspecification when the true
+structure is localised (as in CAR). GLM, which ignores spatial effects,
+performs reasonably in some settings but often underperforms compared to
+CAR.
+
+In the point-level setting (Figure~\ref{fig-biasplot2}), the mlvCAR
+model shows superior performance, achieving minimal bias across all
+coefficients and data-generating mechanisms. This reflects its strength
+in modeling spatial dependencies while accounting for multilevel data
+structure (i.e., multiple observations within areas). GLMM performs
+moderately well but tends to show slightly higher bias than mlvCAR. In
+contrast, GWR, despite being designed for point-level spatial variation,
+yields higher bias, especially under SAR+CAR structures. This suggests
+GWR's limitation when spatial heterogeneity is not smooth or when model
+assumptions are violated.
 
@@
 
-Table~\ref{tbl-rmse} provide insights into model accuracy at different
-spatial levels. For area-level data, the models GLM, SAR, and CAR are
-compared, with the CAR model consistently achieving the lowest RMSE.
-This suggests that CAR is more effective at capturing spatial
+Moreover, Table~\ref{tbl-rmse} provide insights into model accuracy at
+different spatial levels. For area-level data, the models GLM, SAR, and
+CAR are compared, with the CAR model consistently achieving the lowest
+RMSE. This suggests that CAR is more effective at capturing spatial
 dependencies at the area level. For point-level data, the models GLMM,
@@
 
-In addition to the point-level dataset (\texttt{ppd}), which contains
-individual property listings, the imputed dataset was also aggregated to
-the sub-district level by calculating the mean values of key variables.
-This aggregation produced an area-level dataset (\texttt{ppd\_kc}),
-allowing the analysis to be conducted at both the individual and
-administrative levels. These two levels of data granularity provide
-complementary perspectives for evaluating the performance of spatial
-models in capturing local property market dynamics.
+In addition to the point-level dataset, which contains individual
+property listings, the imputed dataset was also aggregated to the
+sub-district level by calculating the mean values of key variables. This
+aggregation produced an area-level dataset, allowing the analysis to be
+conducted at both the individual and administrative levels. These two
+levels of data granularity provide complementary perspectives for
+evaluating the performance of spatial models in capturing local property
+market dynamics.
 
@@
 interpretation and comparison across different modeling approaches. The
-results presented in the following tables provide a comprehensive
+results presented in Table~\ref{tbl-modelcomp} provide a comprehensive
 evaluation of each model's performance in capturing both the structural
-and spatial characteristics of the property market in Lombok. This
-comparison allows us to assess the extent to which spatially explicit
-models (e.g., GWR, CAR, SAR) offer improvements over conventional
-methods in explaining price variability across regions. --- Let me know
-if you'd like to include metrics such as R², AIC, or Moran's I in the
-paragraph as well.
+and spatial characteristics of the property market in Lombok.
 
 We conducted model fitting on the area-level data, with detailed results
-provided in the Appendix (Table~\ref{tbl-modelcomp2}). However, this
+provided in the Appendix (Table~\ref{tbl-modelcomp1}). However, this
 method does not provide optimal parameter estimates because data
@@
 
+\begin{figure}
+
+\begin{minipage}{0.50\linewidth}
+
+\centering{
+
+\includegraphics{figures/phi_map.png}
+
+}
+
+\subcaption{\label{fig-pricenphi1}phi values}
+
+\end{minipage}%
+%
+\begin{minipage}{0.50\linewidth}
+
+\centering{
+
+\includegraphics{figures/logprice_map.png}
+
+}
+
+\subcaption{\label{fig-pricenphi2}Aggregate price data (log scale)}
+
+\end{minipage}%
+
+\caption{\label{fig-pricenphi}Heatmap of aggregated price data and phi
+values in each Lombok sub-district}
+
+\end{figure}%
+
+\textsubscript{Source:
+\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
+Notebook}}
+
+Figure~\ref{fig-pricenphi1} and Figure~\ref{fig-pricenphi2} respectively
+present spatial heatmaps illustrating the distribution of estimated
+spatial random effects (\(\phi\)) from the CAR model and the average
+log-transformed house prices per sub-district accross Lombok island.
+
+Here, the \(\phi\) values represent the spatially structured residuals
+not explained by the observed covariates in the CAR model. Specifically,
+it captures latent spatial influences such as neighbourhood
+desirability, access to infrastructure, or even local policy, that
+affect house prices beyond what structural features like land and
+building size can account for.
+
+Because the model uses log-transformed prices, \(\phi\) has
+multiplicative interpretation on the original price scale. For instance,
+if \(\phi = 0.57\) for a given sub-district, then after accounting for
+covariates, properties in that area are estimated to be approximately
+\(exp(0.57) \approx 1.76\) or 76\% more expensive than average.
+Similarly, a sub-district with \(\phi = -0.2\), this suggests that
+properties in that area are approximately 18\% less expensive than
+average, since \(exp(-0.2) \approx 0.82\).
+
 \section{Results and Discussion}\label{results-and-discussion}
@@
 achieves substantially lower RMSE values compared to GLM and SAR,
-suggesting its effectiveness in handling localized spatial variations.
+suggesting its effectiveness in handling localised spatial variations.
 
@@
 spatial dependencies are strong.
+
+Spatial maps of the estimated random effects \(\phi\) from the CAR model
+offer further insight into the spatial structure of housing prices
+across Lombok Island. These effects represent residual spatial patterns
+after accounting for structural covariates like land size, building
+size, and number of rooms. Such interpretability supports the use of CAR
+models not only for improved prediction but also for spatial
+diagnostics. The visual comparison of \(\phi\) values and aggregated
+log-transformed prices (as shown in Figure~\ref{fig-pricenphi})
+highlights spatial patterns that may warrant further
+investigation---such as clusters of residual over- or under-prediction
+in West and East Lombok respectively. These patterns underscore the
+added value of CAR-type models in revealing latent geographic effects.
+
+These findings echo the limitations of GWR when precise coordinates are
+lacking, affirming the relevance of area-level approaches. Moreover,
+while we use a binary adjacency matrix for consistency, we recognise
+that the choice of spatial weight \(\mathbf{W}\) may influences model
+performance. Future studies should explore alternative specifications
+such as row-standardized or distance-decay matrices to assess their
+impact on robustness and interpretability.
+
+Additionally, while the spatial dependence parameter \(\rho\) is
+estimated as part of the model fitting, its variability across
+simulations suggests that further sensitivity analyses---especially
+under different fixed \(\rho\) values---could shed light on model
+stability in different spatial contexts.
 
@@
 of spatial dependency to further validate their effectiveness.
+
+In summary, this study highlights the importance of explicitly modeling
+spatial dependence in house price analysis. CAR and mlvCAR models emerge
+as practical and theoretically sound options, especially when
+high-resolution spatial data are unavailable. Future research may extend
+these models to finer spatial scales, incorporate additional spatial
+diagnostics, or apply them in other domains where spatial heterogeneity
+significantly influences outcomes.
 
@@
 
-::: \{.appendices latex=true\}
-
-\section{Simple Feature Data Slice}\label{secA1}
-
-\begin{longtable}[t]{rl}
-
-\caption{\label{tbl-geoartfex}a slice of a simple feature collection
-with 6 polygon features and 1 attribute field, area(sqm), representing
-polygon sizes in square units. The dataset, using the WGS 84 CRS, spans
-longitudes 116.0355--116.5 and latitudes -8.500079---8.000057. The
-geometry field defines each polygon's spatial boundaries, representing
-artificial study regions for spatial modeling in this research}
-
-\tabularnewline
-
-\toprule
-area & geometry\\
-\midrule
-17616519 & POLYGON ((116.2296 -8.50007...\\
-11485356 & POLYGON ((116.0778 -8.32479...\\
-4755629 & POLYGON ((116.399 -8.380894...\\
-15959450 & POLYGON ((116.3567 -8.00006...\\
-12727929 & POLYGON ((116.14 -8.292773,...\\
-\addlinespace
-8848979 & POLYGON ((116.5 -8.210459, ...\\
-7248031 & POLYGON ((116.4308 -8.45091...\\
-16906811 & POLYGON ((116.1352 -8.26696...\\
-6346890 & POLYGON ((116.0723 -8.30335...\\
-7620652 & POLYGON ((116.4541 -8.43327...\\
-\addlinespace
-10987363 & POLYGON ((116.4982 -8.06795...\\
-10411472 & POLYGON ((116.1071 -8.07289...\\
-20452118 & POLYGON ((116.4233 -8.12744...\\
-19613075 & POLYGON ((116 -8.5, 116 -8....\\
-9252319 & POLYGON ((116.4541 -8.43327...\\
-\bottomrule
-
-\end{longtable}
-
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
-\section{Bias values for Area-level Data}\label{secA2}
+\section*{Appendix}\label{appendix}
+\addcontentsline{toc}{section}{Appendix}
+
+\subsection*{Bias values for Area-level
+Data}\label{bias-values-for-area-level-data}
+\addcontentsline{toc}{subsection}{Bias values for Area-level Data}
 
@@
 
-[!h]
 \centering\begingroup\fontsize{9}{11}\selectfont
@@
 
-\section{Model Comparison for area-level data}\label{secA3}
+\subsection*{Model Comparison for area-level
+data}\label{model-comparison-for-area-level-data}
+\addcontentsline{toc}{subsection}{Model Comparison for area-level data}
 
@@
 
-\caption{\label{tbl-modelcomp2}This table compares the performance of
+\caption{\label{tbl-modelcomp1}This table compares the performance of
 GLM, SAR, and CAR models on area-level data. It reports parameter
@@
 
-\section{Example of beta-values in GWR model}\label{secA4}
+\subsection*{Example of beta-values in GWR
+model}\label{example-of-beta-values-in-gwr-model}
+\addcontentsline{toc}{subsection}{Example of beta-values in GWR model}
 

### 2025-05-05 docs/manuscript.qmd
update

@@
     email: indiraputeri@uinmataram.ac.id
+    corresponding: true
     affiliations: 
@@
     url: 'https://haziqj.ml'
-    corresponding: true
+    corresponding: false
     affiliations:
@@
     email: elvynna.leong@ubd.edu.bn
-    corresponding: true
+    corresponding: false
     affiliations:
@@
 The weight matrix visually shows how different regions are connected to each other, indicating their neighbourhood relationships (@morris2019bayesian).
-To demonstrate this concept, let's take a simple example of a map with four regions $(R_1, R_2, R_3, R_4)$, as visualised in @fig-neighsimple.
+To demonstrate this concept, let's take a simple example of a map with 5 regions $(A, B, C, D, E)$, as visualised in @fig-neighsimple.
 
@@
 The figure depicts a simplified spatial configuration where each numbered area represents a distinct spatial unit, demonstrating how neighbouring relationships can be defined for spatial modelling purposes."
-#| out-width: "70%"
+#| out-width: "50%"
 #| fig-align: "center"
 
-knitr::include_graphics("figures/maprec.png", auto_pdf = TRUE)
+knitr::include_graphics("figures/sartf_plot.png", auto_pdf = TRUE)
 ```
 
-This straightforward neighbourhood structure creates a single connected component with the following relationships: $(R_1 \sim R_2,\ R_2 \sim R_3,\ R_3 \sim R_4)$.
-Therefore the weight matrix  $\boldsymbol{W}$ is
+From the @fig-neighsimple we can derive an adjacency matrix which represents the neighborhood structure of five spatial units labeled $A$ to $E$.
+Each entry 1 indicates a direct adjacency (areas share a common border), while 0 denotes no direct connection.
+For example, Area $A$ is adjacent to all other areas ($B$, $C$, $D$, and $E$), whereas Area $B$ is only adjacent to $A$ and $C$.
+This neighbourhood configuration captures a clear spatial interaction pattern among the areas, which forms the basis for constructing a spatial weight matrix, commonly denoted as the $\mathbf{W}$ matrix
 
 $$
-\begin{array}{ccccc}
-    & R_1 & R_2 & R_3 & R_4 \\
-R_1 & 0 & 1 & 0 & 0\\ 
-R_2 & 1 & 0 & 1 & 0\\ 
-R_3 & 0 & 1 & 0 & 1\\
-R_4 & 0 & 0 & 1 & 0\\
+\begin{array}{cccccc}
+    & A & B & C & D & E \\
+A & 0 & 1 & 1 & 1 & 1\\ 
+B & 1 & 0 & 1 & 0 & 0\\ 
+C & 1 & 1 & 0 & 1 & 0\\
+D & 1 & 0 & 1 & 0 & 1\\
+E & 1 & 0 & 0 & 1 & 0\\
 \end{array}
@@
 Later on, in the modelling stage, we will also need to construct a diagonal matrix $\boldsymbol{D}$.
-It is an $N \times N$ matrix where each diagonal element $d_{ii}$ denotes the number of neighbours of region $R_i$, while all non-diagonal elements are zero.
+It is an $N \times N$ matrix where each diagonal element $d_{ii}$ denotes the number of neighbours of region.
+while all non-diagonal elements are zero.
 This matrix is essential in spatial econometric models because it captures the local neighbourhood structure of each region.
@@
 $$
-\begin{array}{ccccc}
-   & R_1 & R_2 & R_3 & R_4 \\
-R_1 & 1 & 0 & 0 & 0\\ 
-R_2 & 0 & 2 & 0 & 0\\ 
-R_3 & 0 & 0 & 2 & 0\\
-R_4 & 0 & 0 & 0 & 1\\
+\begin{array}{cccccc}
+   & A & B & C & D & E \\
+A & 4 & 0 & 0 & 0 & 0\\ 
+B & 0 & 2 & 0 & 0 & 0\\ 
+C & 0 & 0 & 3 & 0 & 0\\
+D & 0 & 0 & 0 & 3 & 0\\
+E & 0 & 0 & 0 & 0 & 2\\
 \end{array}
@@
 
-This matrix format indicates that region $R_1$ has one neighbour, regions $R_2$ and $R_3$ each have two neighbours, and region $R_4$ has one neighbour.
-By integrating this matrix into our modelling framework, we ensure that the spatial structure and interactions among regions are accurately represented, thereby enhancing the robustness and precision of our spatial analysis.
+This matrix format indicates that region $A$ has 4 neighbour, regions $C$ and $D$ each have 3 neighbours, and region $B$ and $E$ has 2 neighbour.
 
@@
 
-The bias values show that the models exhibit varying levels of accuracy across the different parameters and datasets.
-The coefficients $\beta_k$ for $k = 0, 1, 2, 3, 4$ represent the values associated with the intercept, land size, building size, number of bedrooms, and number of bathrooms, respectively.
-For $\beta_0$, the GLM and CAR models display relatively small biases with narrow standard deviations, indicating better consistency, while the SAR model demonstrates larger bias values, suggesting potential issues with accurately estimating the intercept.
-For higher parameters like $\beta_2$ and $\beta_3$, SAR generally has larger bias and variability, especially in the $SAR + CAR$ dataset, suggesting that the combined spatial structure poses challenges for SAR's estimation.
-It is important to note that the  @fig-biasarealevel utilises area-level data.
-In this context, the CAR model consistently demonstrates the smallest bias compared to the other two models.
+The bias values are summarised in @tbl-bias1 and @tbl-bias2.
+It shows that the models exhibit varying levels of accuracy across the different parameters and datasets.
+@fig-biasplot presents the bias estimates of regression coefficients for different modelling approaches, separated by data type: (a) area-level data and (b) point-level data.
+Each panel corresponds to a data-generating mechanism (CAR, GWR, or SAR+CAR), and bias values are shown with 95% confidence intervals.
+
+::: {#fig-biasplot layout-ncol=2}
+
+![Bias values by method and dataset type for area-level data](figures/bias1_plot.png){#fig-biasplot1}
+
+![Bias values by method and dataset type for point-level data](figures/bias2_plot.png){#fig-biasplot2}
+
+Bias estimates comparison of six different methods across datasets generated under CAR, GWR, and SAR+CAR frameworks.
+Each method's performance is visualised for $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$, highlighting variations in bias and the effectiveness of each model in addressing spatial dependencies.
+:::
 
 ```{r, echo = FALSE, message = FALSE, warning = FALSE}
-#| label: fig-biasarealevel
-#| fig-cap: "Bias values by method and dataset type for area-level data.
-This figure compares the bias estimates of different methods (CAR, GLM, SAR) across datasets generated under CAR, GWR, and SAR+CAR frameworks.
+# #| label: fig-biasplot
+# #| fig-cap: "Bias values by method and dataset type for area-level data and also point-level data.
+This figure compares the bias estimates of different methods (CAR, GLM, SAR, GLMM, GWR, and mlvCAR) across datasets generated under CAR, GWR, and SAR+CAR frameworks.
 Each method's performance is visualised for four coefficients ($\\beta_1$, $\\beta_2$, $\\beta_3$, $\\beta_4$), highlighting variations in bias and the effectiveness of each model in addressing spatial dependencies."
-#| out-width: "75%"
-#| fig-align: "center"
-
-knitr::include_graphics("figures/bias1b_plot.pdf", auto_pdf = TRUE)
+# #| out-width: "100%"
+# #| fig-align: "center"
+# #| layout-ncol: 2
+# 
+# knitr::include_graphics("figures/bias1_plot.png", auto_pdf = TRUE)
+# knitr::include_graphics("figures/bias2_plot.png", auto_pdf = TRUE)
 ```
 
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
-#| label: fig-biaspointlevel
-#| fig-cap: "Bias values by method and dataset type for point-level data.
-This figure compares the bias estimates of different methods (GLMM, GWR, and mlvCAR) across datasets generated under CAR, GWR, and SAR+CAR frameworks.
-Each method's performance is visualised for four coefficients ($\\beta_1$, $\\beta_2$, $\\beta_3$, $\\beta_4$), highlighting variations in bias and the effectiveness of each model in addressing spatial dependencies."
-#| out-width: "75%"
-#| fig-align: "center"
+For area-level data (@fig-biasplot1), CAR model consistently produces the lowest bias across all coefficients and data-generating scenarios.
+This confirms its ability to recover true parameter values under spatial dependency, even when data is aggregated.
+The SAR model displays higher variability and often substantial bias, especially in CAR-generated datasets, suggesting model misspecification when the true structure is localised (as in CAR).
+GLM, which ignores spatial effects, performs reasonably in some settings but often underperforms compared to CAR.
 
-knitr::include_graphics("figures/bias2_plot.pdf", auto_pdf = TRUE)
-```
-
-As shown in the second plot of @fig-biaspointlevel, this analysis utilises point-level data.
-Overall, the mlvCAR model demonstrates a consistently lower bias across most parameters and scenarios.
-In the $SAR + CAR$ scenario, where spatial complexity is most pronounced, mlvCAR achieves notably low bias, particularly for parameters $\beta_1$ through $\beta_4$, indicating its suitability for handling mixed spatial structures.
-making it a solid choice for a range of spatial modelling tasks.
-
-The GWR model, which provides highly localised estimates, displays slightly higher bias for several parameters compared to mlvCAR, especially for $\beta_1$, $\beta_2$, and $\beta_3$ within the $CAR$ scenario.
-This suggests GWR’s greater sensitivity to spatial patterns without incorporating neighbourhood-based effects.
-In contrast, GLMM, which lacks explicit spatial components, shows larger bias variability, particularly in scenarios with pronounced spatial dependencies, highlighting its limitations for purely spatial analyses.
+In the point-level setting (@fig-biasplot2), the mlvCAR model shows superior performance, achieving minimal bias across all coefficients and data-generating mechanisms.
+This reflects its strength in modeling spatial dependencies while accounting for multilevel data structure (i.e., multiple observations within areas).
+GLMM performs moderately well but tends to show slightly higher bias than mlvCAR.
+In contrast, GWR, despite being designed for point-level spatial variation, yields higher bias, especially under SAR+CAR structures.
+This suggests GWR's limitation when spatial heterogeneity is not smooth or when model assumptions are violated.
 
@@
 
-@tbl-rmse provide insights into model accuracy at different spatial levels.
+Moreover, @tbl-rmse provide insights into model accuracy at different spatial levels.
 For area-level data, the models GLM, SAR, and CAR are compared, with the CAR model consistently achieving the lowest RMSE.
@@
 
-In addition to the point-level dataset (`ppd`), which contains individual property listings, the imputed dataset was also aggregated to the sub-district level by calculating the mean values of key variables.
-This aggregation produced an area-level dataset (`ppd_kc`), allowing the analysis to be conducted at both the individual and administrative levels.
+In addition to the point-level dataset, which contains individual property listings, the imputed dataset was also aggregated to the sub-district level by calculating the mean values of key variables.
+This aggregation produced an area-level dataset, allowing the analysis to be conducted at both the individual and administrative levels.
 These two levels of data granularity provide complementary perspectives for evaluating the performance of spatial models in capturing local property market dynamics.
@@
 Standardization of the predictors ensures that the resulting coefficients are on a comparable scale, facilitating meaningful interpretation and comparison across different modeling approaches.
-The results presented in the following tables provide a comprehensive evaluation of each model’s performance in capturing both the structural and spatial characteristics of the property market in Lombok.
-This comparison allows us to assess the extent to which spatially explicit models (e.g., GWR, CAR, SAR) offer improvements over conventional methods in explaining price variability across regions.
---- Let me know if you’d like to include metrics such as R², AIC, or Moran’s I in the paragraph as well.
+The results presented in @tbl-modelcomp provide a comprehensive evaluation of each model’s performance in capturing both the structural and spatial characteristics of the property market in Lombok.
 
-We conducted model fitting on the area-level data, with detailed results provided in the Appendix (@tbl-modelcomp2).
+We conducted model fitting on the area-level data, with detailed results provided in the Appendix (@tbl-modelcomp1).
 However, this method does not provide optimal parameter estimates because data aggregation can reduce variability and obscure spatial details essential for capturing local effects.
@@
 
+::: {#fig-pricenphi layout-ncol=2}
+
+![phi values](figures/phi_map.png){#fig-pricenphi1}
+
+![Aggregate price data (log scale)](figures/logprice_map.png){#fig-pricenphi2}
+
+Heatmap of aggregated price data and phi values in each Lombok sub-district
+:::
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
+# #| label: fig-pricenphi
+# #| fig-cap: "Heatmap of aggregated price data and phi values in each Lombok sub-district"
+# #| eval: true
+# #| layout-ncol: 2
+# 
+# knitr::include_graphics("figures/phi_map.png", auto_pdf = TRUE)
+# knitr::include_graphics("figures/logprice_map.png", auto_pdf = TRUE)
+
+```
+
+@fig-pricenphi1 and @fig-pricenphi2 respectively present spatial heatmaps illustrating the distribution of estimated spatial random effects ($\phi$) from the CAR model and the average log-transformed house prices per sub-district accross Lombok island.
+
+Here, the $\phi$ values represent the spatially structured residuals not explained by the observed covariates in the CAR model.
+Specifically, it captures latent spatial influences such as neighbourhood desirability, access to infrastructure, or even local policy, that affect house prices beyond what structural features like land and building size can account for.
+
+Because the model uses log-transformed prices, $\phi$ has multiplicative interpretation on the original price scale.
+For instance, if $\phi = 0.57$ for a given sub-district, then after accounting for covariates, properties in that area are estimated to be approximately $exp(0.57) \approx 1.76$ or 76% more expensive than average.
+Similarly, a sub-district with $\phi = -0.2$, this suggests that properties in that area are approximately 18% less expensive than average, since $exp(-0.2) \approx 0.82$.
+
 # Results and Discussion
@@
 This is evidenced by lower RMSE values and reduced bias across various dataset scenarios, highlighting the CAR model’s robustness.
-In datasets with strong spatial structures, such as those generated under $CAR$ and $SAR + CAR$ conditions, the CAR model achieves substantially lower RMSE values compared to GLM and SAR, suggesting its effectiveness in handling localized spatial variations.
+In datasets with strong spatial structures, such as those generated under $CAR$ and $SAR + CAR$ conditions, the CAR model achieves substantially lower RMSE values compared to GLM and SAR, suggesting its effectiveness in handling localised spatial variations.
 
@@
 
+Spatial maps of the estimated random effects $\phi$ from the CAR model offer further insight into the spatial structure of housing prices across Lombok Island.
+These effects represent residual spatial patterns after accounting for structural covariates like land size, building size, and number of rooms.
+Such interpretability supports the use of CAR models not only for improved prediction but also for spatial diagnostics.
+The visual comparison of $\phi$ values and aggregated log-transformed prices (as shown in @fig-pricenphi) highlights spatial patterns that may warrant further investigation—such as clusters of residual over- or under-prediction in West and East Lombok respectively.
+These patterns underscore the added value of CAR-type models in revealing latent geographic effects.
+
+These findings echo the limitations of GWR when precise coordinates are lacking, affirming the relevance of area-level approaches.
+Moreover, while we use a binary adjacency matrix for consistency, we recognise that the choice of spatial weight $\mathbf{W}$ may influences model performance.
+Future studies should explore alternative specifications such as row-standardized or distance-decay matrices to assess their impact on robustness and interpretability.
+
+Additionally, while the spatial dependence parameter $\rho$ is estimated as part of the model fitting, its variability across simulations suggests that further sensitivity analyses—especially under different fixed $\rho$ values—could shed light on model stability in different spatial contexts.
+
 Additionally, the limitations of point-level models such as GWR in the absence of precise location data highlight a practical challenge in spatial modeling.
@@
 Future research could examine the use of CAR and mlvCAR models with different spatial resolutions or apply these models to other datasets with varying levels of spatial dependency to further validate their effectiveness.
+
+In summary, this study highlights the importance of explicitly modeling spatial dependence in house price analysis.
+CAR and mlvCAR models emerge as practical and theoretically sound options, especially when high-resolution spatial data are unavailable.
+Future research may extend these models to finer spatial scales, incorporate additional spatial diagnostics, or apply them in other domains where spatial heterogeneity significantly influences outcomes.
 
@@
 
-::: {.appendices latex=true}
+# Appendix {.unnumbered}
 
-# Simple Feature Data Slice {#secA1}
-
-```{r, echo=FALSE, eval=TRUE, message=FALSE}
-#| label: tbl-geoartfex
-#| tbl-cap: "a slice of a simple feature collection with 6 polygon features and 1 attribute field, area(sqm), representing polygon sizes in square units.
-The dataset, using the WGS 84 CRS, spans longitudes 116.0355–116.5 and latitudes -8.500079–-8.000057.
-The geometry field defines each polygon's spatial boundaries, representing artificial study regions for spatial modeling in this research"
-#| eval: true
-
-artfcountry <- sf::read_sf("data/", quiet = TRUE)
-
-# Create the table with a caption
-artfcountry %>%
-  head(n = 15) %>%
-  kable() %>%
-  kable_styling()
-```
-
-# Bias values for Area-level Data {#secA2}
+## Bias values for Area-level Data {.unnumbered}
 
@@
 ) %>%
-  kable_styling(font_size = 9, latex_options = "hold_position") %>%
+  kable_styling(font_size = 9) %>%
   add_header_above(c(" " = 1, "Mean ± SD" = 3)) %>%
@@
 
-# Model Comparison for area-level data {#secA3}
+## Model Comparison for area-level data{.unnumbered}
 
 ```{r, echo=FALSE, message=FALSE}
-#| label: tbl-modelcomp2
+#| label: tbl-modelcomp1
 #| tbl-cap: "This table compares the performance of GLM, SAR, and CAR models on area-level data.
@@
     booktabs = T, escape = FALSE) %>%
-  kable_styling(font_size = 9, latex_options = "hold_position") %>%
+  kable_styling(font_size = 9) %>%
   pack_rows("Coefficients", 1, 5) %>%
@@
 
-# Example of beta-values in GWR model {#secA4}
+## Example of beta-values in GWR model{.unnumbered}
 

### 2025-05-05 R/02_kcmodel.R

@@ -379,5 +379,66 @@ summary_stats <- apply(posterior_samples, 2, function(x) {
 
 summary_stats
 
-
 resid_car <- model_car$fitted.values - ppd_kc
+
+# Plot--------------------------
+library(tidyr)
+library(ggplot2)
+load("data/ppd_kc.RData")
+load("data/lop_kc.RData")
+lop_kc <- lop_kc %>% rename(SubDistrict = WADMKC,
+                            District    = WADMKK) %>% 
+  mutate(ID = 1:nrow(lop_kc))
+
+# extracting phi values from mlvcar model
+model.mlvcar <- readRDS("data/model_mlvcar.rds")
+phi <- model.mlvcar$samples$phi %>% 
+  colMeans()
+lop_kc$phi <- phi
+
+#price data from ppd_kc dataset
+price_kc <- ppd_kc %>% 
+  select(SubDistrict, prices)
+
+lop_kc <- lop_kc %>% 
+  left_join(price_kc, by = "SubDistrict") %>% 
+  mutate(logprice = log(prices))
+
+lop_kc$phi <- phi
+
+phi_plot <- ggplot(lop_kc) +
+  geom_sf(aes(fill = phi)) +
+  scale_fill_distiller(palette = "RdBu") +
+  theme_minimal() +
+  theme(
+    strip.text = element_text(size = 8),
+    axis.text = element_blank(),
+    legend.title = element_text(size = 8),
+    legend.text = element_text(size = 8),
+    legend.position = "right"
+  )  
+
+phi_plot
+ggsave("figures/phi_map.png", plot = phi_plot, width = 8, 
+       height = 6, dpi = 300)
+
+logprice_plot <- ggplot(lop_kc) +
+  geom_sf(aes(fill = logprice)) +
+  scale_fill_distiller(palette = "RdBu") +
+  theme_minimal() +
+  theme(
+    strip.text = element_text(size = 8),
+    axis.text = element_blank(),
+    legend.title = element_text(size = 8),
+    legend.text = element_text(size = 8),
+    legend.position = "right"
+  )  
+
+logprice_plot
+ggsave("figures/logprice_map.png", plot = logprice_plot, width = 8, 
+       height = 6, dpi = 300) 
+  
+  
+  
+  
+  

### 2025-05-06 docs/response/response.qmd
update manuscript and response

@@
 ---
-title: "Author Response to Reviewer Comments"
-# format:
-#   pdf: default
-#bibliography: references.bib
-engine: knitr
+title: "Response to Journal's Reviewer Comments"
+subtitle: "Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study"
 ---
 
-1.  The choice of W (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?
+**Reviewer 1**
 
-<!-- 2.  While the CAR model shows lower bias and RMSE in the artificial study, does this advantage persist in real-world scenarios, particularly where data sparsity or boundary effects exist? -->
+**This manuscript compares different methods (Conditionally Autoregressive, Geographically Weighted Regression, and Simultaneous Autoregressive) for modeling property prices.
+The problem definition looks reasonable, but the manuscript presents only the application part; nothing is contributed toward the theory.
+Therefore, I will suggest transferring this manuscript to an application-related journal.
+To my understanding, this manuscript does not fall in the scope of this journal.
+Further, the manuscript is not standard format of the journal.**
 
-<!-- 3.
-The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model).
-Have the authors examined how sensitive the model results are to changes in $\rho$?
--->
+We acknowledge that our work is primarily application-driven, focusing on the use of Conditionally Autoregressive (CAR), Geographically Weighted Regression (GWR), and Simultaneous Autoregressive (SAR) models in the context of property price modeling.
+However, we respectfully submit that the manuscript makes a meaningful methodological contribution, particularly in bridging the gap between spatial theory and applied practice—an area of growing relevance, especially in spatial analysis for developing regions.
 
-<!-- 4.
-The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants?
-Would spatial effect maps or additional diagnostic tools enhance this understanding?
--->
+While we do not introduce new theoretical constructs, our work extends the methodological discourse in the following ways:
 
-<!-- 5.
-Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research (page 2).
-Before reaching this conclusion, did you compare GWR to other spatial regression models?
--->
+ - **Shifting the Dominant Modeling Paradigm**: Previous studies on house price modeling have predominantly employed Geographically Weighted Regression (GWR), which requires precise geolocation data for model fitting.
+However, such data are often incomplete, unavailable, or unreliable in many regions.
+Our study explores the potential of the Conditionally Autoregressive (CAR) model—more commonly applied in health and ecological studies—as a viable alternative for house price modeling.
+This cross-domain adaptation is underexplored and opens new methodological pathways for spatial econometrics.
 
-<!-- 6.
-The scope of properties analysed includes residential houses and commercial real estate (page 16).
-How do the authors reconcile this with the abstract, which mentions only house price data?
--->
+ - **Unified Comparative Framework**: We develop and apply a systematic framework to compare CAR, SAR, and GWR models using both simulated data and real-world property data.
+This comparative approach, particularly within the under-researched housing market context of a developing region, fills a critical gap in the literature.
 
-<!-- 7.
-After data cleaning, the final dataset included 598 entries representing multiple observations per sub-district (page 16).
-Why does the distribution of these data not appear on the map (Figure 6)?
--->
+ - **Practical Implementation Insights**: The manuscript offers practical guidance for applied researchers on implementing, diagnosing, and interpreting each spatial model.
+These details are especially valuable for practitioners operating in data-scarce environments or policy-driven contexts.
 
-<!-- 8.
-The authors have applied point-level data using three different models (GLMM, GWR, and multilevel CAR) (page 17).
-How was this approach implemented, given that there is no explanation of how each observation was georeferenced?
--->
+ - **Informing Model Selection**: By revealing the comparative strengths and limitations of each model in capturing spatial heterogeneity, our findings contribute to informed model selection in applied spatial econometrics—balancing methodological soundness with practical feasibility.
 
-<!-- 9.  Why is Table 5 not referred to in the text? -->
+ - **Policy and Planning Relevance**: The results of our study have implications for spatial planning and housing market analysis, offering a framework for integrating spatial models into decision-making processes where traditional approaches may be constrained by data availability.
 
-<!-- 10.
-The authors have argued that point-level models like GWR are limited when precise location data are unavailable (page 18).
+Given these points, we believe the manuscript fits within the journal’s broader mission to advance rigorous spatial analysis methods and encourage cross-disciplinary methodological applications.
+
+**Reviewer 2**
+
+1.  **The choice of $W$ (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?**
+
+    Thank you for this insightful comment regarding the specification of the spatial weight matrix, $W$.
+We fully acknowledge that the choice of $W$—whether binary adjacency, row-standardized, or distance-based—can substantially influence the performance and interpretation of spatial models.
+In the present study, our primary objective was to evaluate and compare the performance of CAR, SAR, and GWR models using both simulated data and real-world data from Lombok.
+To maintain consistency across models and to isolate the structural differences in model behavior, we have, at this stage, limited our analysis to a binary adjacency matrix for $W$.
+We agree that exploring alternative specifications of $W$ (e.g., row-standardized or distance-decay matrices) is a valuable direction for further research.
+We plan to extend our analysis in future work to assess how varying spatial weight structures affect the robustness and predictive accuracy of these spatial models.
+
+2.  **While the CAR model shows lower bias and RMSE in the artificial study, does this advantage persist in real-world scenarios, particularly where data sparsity or boundary effects exist?**
+
+    Thank you for your thoughtful question.
+We acknowledge that in small-area data, bias can indeed become a concern due to limited information at the unit level.
+However, in the case of the multilevel CAR (mlvCAR) model, each area contains multiple observations, allowing us to model area-specific effects—functionally resembling a random effects structure.
+This hierarchical structure helps mitigate the impact of sparsity within individual areas.
+
+    Regarding sample size, prior studies suggest that having approximately 15–30 observations per area is generally sufficient for stable estimation in multilevel frameworks.
+Given our data structure aligns with this recommendation, we believe the mlvCAR model offers a robust approach, especially in real-world scenarios where heterogeneity and partial pooling across areas are important.
+We thus recommend the use of mlvCAR to address both the potential bias and boundary effects effectively.
+
+3.
+**The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model).
+Have the authors examined how sensitive the model results are to changes in $\rho$?**
+
+    In our study, we recognize that the value of $\rho$ (the spatial dependence parameter) plays a crucial role in determining the extent to which spatial autocorrelation is captured by the model.
+While we did not conduct a dedicated sensitivity analysis by fixing $\rho$ at multiple values, our simulation framework implicitly accounts for variability in $\rho$ through repeated data generation and model fitting, where $\rho$ is treated as an estimable parameter.
+The observed variability—such as the estimate of 0.48 in the multilevel CAR model—reflects the differences in underlying spatial structure across datasets rather than instability or inconsistency in the model itself.
+
+    In a Bayesian framework, the prior distribution assigned to $\rho$ also has an important influence, particularly in data-sparse contexts or when spatial signals are weak.
+In our implementation, we used a uniform or weakly informative prior (depending on the model setting), allowing the data to drive the estimation of $\rho$ while still preventing extreme or implausible values.
+In practical applications, however, researchers may wish to incorporate more informative priors based on prior knowledge of spatial processes or domain-specific expectations, which can enhance stability and interpretability in certain contexts.
+
+    That said, we fully agree that a more systematic sensitivity analysis—such as fixing $\rho$ at various plausible levels or experimenting with different prior distributions—could offer deeper insights into model robustness and the influence of spatial structure.
+We consider this an important direction for future research, particularly in settings with heterogeneous spatial patterns or limited sample sizes, where sensitivity to $\rho$ may significantly impact conclusions and policy relevance.
+
+4.
+**The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants?
+Would spatial effect maps or additional diagnostic tools enhance this understanding?**
+
+    We agree that while $\phi$ captures unobserved spatial heterogeneity, its interpretability in the context of house price determinants requires careful consideration.
+Rather than directly reflecting specific covariates, $\phi$ accounts for spatially structured residual variation—such as omitted neighborhood characteristics, local amenities, or spatial spillovers—that are not captured by observed variables.
+
+    For a brief illustration, since the model uses log-transformed prices, $\phi$ has multiplicative interpretation on the original price scale.
+So if $\phi = 0.57$ for a given sub-district, then after accounting for covariates, properties in that area are estimated to be approximately $\exp(0.57) \approx 1.76$ or 76% more expensive than average.
+Similarly, a sub-district with $\phi = -0.2$, this suggests that properties in that area are approximately 18% less expensive than average, since $\exp(-0.2) \approx 0.82$.
+Note that the 95% credible interval of $\phi$ can be interpreted similarly by exponentiating the lower and upper bounds, yielding a multiplicative uncertainty range for the price deviation.
+
+    To facilitate interpretation, we provide a map that visualizes the estimated spatial effects ($\phi$), thereby highlighting the spatial patterns identified by the model.
+
+5.
+**Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research (page 2).
+Before reaching this conclusion, did you compare GWR to other spatial regression models?**
+
+    In our manuscript, we highlight Geographically Weighted Regression (GWR) as a prominent method in the literature, particularly due to its frequent application in house price studies.
+This observation is based on a review of previous works rather than a performance-based comparison, for instance @wang2020geographically, @lu2011geographically, @stewart2018localized, and @cellmer2020spatial.
+While GWR is indeed widely used for its ability to model spatial heterogeneity at the local level, our study goes further by systematically comparing GWR with CAR and SAR models through simulation and real data applications.
+This comparative analysis aims to evaluate their relative strengths and limitations more objectively.
+
+6.  **The scope of properties analysed includes residential houses and commercial real estate (page 16). How do the authors reconcile this with the abstract, which mentions only house price data?**
+
+     We acknowledge that the description in the manuscript may not have been sufficiently clear.
+The original dataset obtained through web-scraping did include a broad range of property types—such as residential houses, villas, boarding houses, and vacant land.
+However, for the purposes of this study, we restricted our analysis to residential houses only.
+We will revise the manuscript accordingly to ensure that this scope is clearly stated in both the abstract and the data collection section.
+
+7.
+**After data cleaning, the final dataset included 598 entries representing multiple observations per sub-district (page 16).
+Why does the distribution of these data not appear on the map (Figure 6)?**
+
+    For clarity, we note that the Lombok house price dataset used in this study does not contain precise geographic coordinates for each observation.
+As a result, it is not possible to accurately display the distribution of the data points on a map.
+While each observation is linked to a sub-district, the absence of point-level location data limits our ability to visualize them spatially within those areas.
+Therefore, for a more informative visualization, we are preparing a heatmap of house prices at the sub-district level, which will reflect the spatial variation in prices across Lombok using aggregated data.
+
+8.
+**The authors have applied point-level data using three different models (GLMM, GWR, and multilevel CAR) (page 17).
+How was this approach implemented, given that there is no explanation of how each observation was georeferenced?**
+
+    Thank you for this important observation.
+We clarify that GLMM and multilevel CAR models do not require explicit geographic coordinates for each individual observation.
+Instead, they rely on an area-level identifier (ID) that links each observation to a corresponding spatial unit (e.g., sub-district).
+This is a similar approach to random effects or multilevel modelling.
+
+    In contrast, GWR requires spatial coordinates for each observation.
+To meet this requirement, we generated random points around the centroid of each sub-district, using `st_sample()`, where the number of points corresponds to the number of observations in that area.
+Each point was placed within a 0.5 km radius from the centroid, ensuring a reasonable spatial spread while maintaining the area-specific context.
+This approach enabled us to simulate realistic spatial positions for property transactions in the absence of precise geolocation data.
+Note that, the same bandwith was used in the fitting of the GWR model.
+By enforcing the area-level grouping, this ensures a fair comparison to the previous method.
+
+9.  **Why is Table 5 not referred to in the text?**
+
+    In the revised version of the manuscript, we have made sure to explicitly refer to Table 5 in the main text to improve clarity and consistency.
+We appreciate your careful reading and helpful suggestion.
+
+10.
+**The authors have argued that point-level models like GWR are limited when precise location data are unavailable (page 18).
 However, for the Lombok House Price Data, the authors have used point-level models.
-How do you reconcile this contradiction?
--->
+How do you reconcile this contradiction?**
 
-<!-- 11.
-In the Lombok house price analysis, no sensitivity analysis was conducted to test how different definitions of adjacency (W matrix) influence the results.
-Would the findings remain consistent if a different spatial weights structure was used?
--->
+    As we noted in our response to Comment 8, the Lombok house price data do not come with precise geographic coordinates.
+In order to apply the `gwr()` function, which requires point-level location data, we generated synthetic coordinates by simulating random points around the centroid of each sub-district.
+While this workaround allowed us to run the GWR model, we acknowledge that it is not an ideal or practical solution for real-world applications.
+This limitation reinforces our argument that point-level models such as GWR can be challenging to apply meaningfully when precise location data are not available.
 
-<!-- 12.
-The study primarily evaluates CAR models within the Lombok context, but how well does it generalize to larger, high-density urban areas (e.g., Jakarta)?
-Would additional testing in a more complex metropolitan setting strengthen the validity of the conclusions?
--->
+11.
+**In the Lombok house price analysis, no sensitivity analysis was conducted to test how different definitions of adjacency (W matrix) influence the results.
+Would the findings remain consistent if a different spatial weights structure was used?**
+
+    As mentioned in our response to Comment 1, in this study we used a binary adjacency matrix to define spatial relationships, primarily to maintain consistency across the CAR and SAR models and to focus on comparing their general performance.
+We acknowledge that the choice of spatial weights matrix ($W$) can significantly influence model outcomes.
+While we have not conducted a sensitivity analysis in this current version, we agree that testing alternative specifications—such as row-standardized or distance-based matrices—would be valuable.
+We consider this an important direction for future work to assess the robustness of our findings under different definitions of spatial connectivity.
+
+12.
+**The study primarily evaluates CAR models within the Lombok context, but how well does it generalize to larger, high-density urban areas (e.g., Jakarta)?
+Would additional testing in a more complex metropolitan setting strengthen the validity of the conclusions?**
+
+    Thank you for this important point.
+Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater.
+In contrast, there is relatively limited research focused on low-density or semi-rural contexts such as Lombok.
+Our study aims to address this gap by evaluating the applicability and performance of spatial models, particularly the CAR model, in a setting that has different data characteristics and spatial structures compared to major metropolitan areas.
+
+    We also acknowledge that in richer datasets—common in urban contexts where more granular, high-frequency, and multidimensional data are available—machine learning approaches may offer competitive or even superior predictive performance.
+However, our focus here is to assess interpretable spatial models in contexts where data limitations remain a challenge.
+
+# References

### 2025-05-06 docs/_tex/manuscript.tex
update manuscript and response

@@
 Recent studies have leveraged GWR to investigate the spatial
-heterogeneity of housing market determinants. For example, Wang and Chen
-(2020) applied GWR to examine the impact of local built-environment
-factors on home prices across different phases of the housing market
-cycle. Their findings highlighted significant spatial variations,
-underscoring the critical role of local context in influencing home
-prices. Similarly, Lu, Charlton, and Fotheringham (2011) employed GWR
-with a non-Euclidean distance metric to analyze London house price data.
-By incorporating this alternative metric, they better captured the
-spatially varying relationships between house prices and their
-determinants, emphasizing the importance of considering different
-distance metrics in spatial analyses. In their 2018 study, Stewart
-Fotheringham and Park (2018) explored spatial and temporal variations in
-the determinants of apartment prices in Seoul, South Korea, over a
-decade. Utilizing a hedonic price model with a spatio-temporal lag,
-calibrated through GWR, they revealed consistent spatial variations and
-strong spatial lag effects on house prices over time. Additionally,
-Cellmer, Cichulska, and Bełej (2020) applied GWR to assess the spatial
-variations in determinants of the housing market in Poland. The study
-identified spatial patterns of autocorrelation in average housing prices
-and market activity, illustrating the value of GWR in examining
-localized effects of socio-demographic and environmental factors on the
-housing market. Collectively, these studies demonstrate the efficacy of
-GWR and its extensions in capturing the nuanced spatial and temporal
-variations in housing market data, providing valuable insights for urban
-planning and policy-making.
+heterogeneity of housing market determinants. For example, C.-H. Wang
+and Chen (2020) applied GWR to examine the impact of local
+built-environment factors on home prices across different phases of the
+housing market cycle. Their findings highlighted significant spatial
+variations, underscoring the critical role of local context in
+influencing home prices. Similarly, Lu, Charlton, and Fotheringham
+(2011) employed GWR with a non-Euclidean distance metric to analyze
+London house price data. By incorporating this alternative metric, they
+better captured the spatially varying relationships between house prices
+and their determinants, emphasizing the importance of considering
+different distance metrics in spatial analyses. In their 2018 study,
+Stewart Fotheringham and Park (2018) explored spatial and temporal
+variations in the determinants of apartment prices in Seoul, South
+Korea, over a decade. Utilizing a hedonic price model with a
+spatio-temporal lag, calibrated through GWR, they revealed consistent
+spatial variations and strong spatial lag effects on house prices over
+time. Additionally, Cellmer, Cichulska, and Bełej (2020) applied GWR to
+assess the spatial variations in determinants of the housing market in
+Poland. The study identified spatial patterns of autocorrelation in
+average housing prices and market activity, illustrating the value of
+GWR in examining localized effects of socio-demographic and
+environmental factors on the housing market. Collectively, these studies
+demonstrate the efficacy of GWR and its extensions in capturing the
+nuanced spatial and temporal variations in housing market data,
+providing valuable insights for urban planning and policy-making.
 
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 Another specific ingredient in spatial modelling is the existence of
@@
 \end{figure}%
-
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
 
@@
 \[
+\mathbf{W} =
 \begin{array}{cccccc}
@@
 E & 1 & 0 & 0 & 1 & 0\\
+\end{array}
+\hspace{2.5cm}
+\mathbf{D} =
+\begin{array}{cccccc}
+   & A & B & C & D & E \\
+A & 4 & 0 & 0 & 0 & 0\\ 
+B & 0 & 2 & 0 & 0 & 0\\ 
+C & 0 & 0 & 3 & 0 & 0\\
+D & 0 & 0 & 0 & 3 & 0\\
+E & 0 & 0 & 0 & 0 & 2\\
 \end{array}
@@
 connectivity of regions (Brunsdon, Fotheringham, and Charlton (1996),
-Stewart Fotheringham and Park (2018)). Here, \(\boldsymbol{D}\) is
-represented as:
-
-\[
-\begin{array}{cccccc}
-   & A & B & C & D & E \\
-A & 4 & 0 & 0 & 0 & 0\\ 
-B & 0 & 2 & 0 & 0 & 0\\ 
-C & 0 & 0 & 3 & 0 & 0\\
-D & 0 & 0 & 0 & 3 & 0\\
-E & 0 & 0 & 0 & 0 & 2\\
-\end{array}
-\]
-
-This matrix format indicates that region \(A\) has 4 neighbour, regions
-\(C\) and \(D\) each have 3 neighbours, and region \(B\) and \(E\) has 2
-neighbour.
+Stewart Fotheringham and Park (2018)). In this case, \(\mathbf{D}\)
+indicates that region \(A\) has 4 neighbours, regions \(C\) and \(D\)
+each have 3 neighbours, and region \(B\) and \(E\) has 2 neighbours.
 
@@
 
-\includegraphics[width=0.7\textwidth,height=\textheight]{figures/nb_artfcountry.png}
+\includegraphics[width=0.6\textwidth,height=\textheight]{figures/nb_artfcountry.png}
 
@@
 
-Here, \(\boldsymbol{\varepsilon} \sim N(0, \nu^2)\) while
-\(\boldsymbol{\phi}\) is a specific component in CAR model that has a
-role as spatial effect. It is also common to mention
+Here, \(\boldsymbol{\varepsilon} \sim N(0, \sigma_{\varepsilon}^2)\)
+while \(\boldsymbol{\phi}\) is a specific component in CAR model that
+has a role as spatial effect. It is also common to mention
 \(\boldsymbol{\phi}\) as a CAR priors. It is a type of Gaussian Markov
@@
 \begin{equation}\phantomsection\label{eq-phidist}{
-\boldsymbol{\phi}\sim N(0,\tau^2\boldsymbol{Q}^{-1}) 
+\boldsymbol{\phi}\sim N(0,\sigma_{\phi}^2\boldsymbol{Q}^{-1}) 
 }\end{equation}
@@
 \begin{equation}\phantomsection\label{eq-phicar}{
-\phi_k|\boldsymbol\phi_{-K}, \boldsymbol{W},\rho, \tau^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\tau^2}{A}\right)
+\phi_k|\boldsymbol{\phi}_{-k}, \mathbf{W},\rho, \sigma_{\phi}^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\sigma_{\phi}^2}{A}\right)
 }\end{equation}
@@
 for all model parameters, such as the regression coefficients \(\beta\),
-variance parameters \(\nu^2\) and \(\tau^2\), and spatial random effects
-\(\phi_k\). Subsequently, Bayesian inference techniques like Markov
-Chain Monte Carlo (MCMC) are employed to derive posterior distributions
-of these parameters. This method facilitates the integration of prior
-knowledge and offers a versatile framework for quantifying uncertainty.
+variance parameters \(\sigma_{\varepsilon}^2\) and \(\sigma_{\phi}^2\),
+and spatial random effects \(\phi_k\). Subsequently, Bayesian inference
+techniques like Markov Chain Monte Carlo (MCMC) are employed to derive
+posterior distributions of these parameters. This method facilitates the
+integration of prior knowledge and offers a versatile framework for
+quantifying uncertainty.
 
@@
 \addlinespace
-Error Term variance ($\sigma^2$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
 Spatial Autocorrelation ($\rho$) & $\times$ & $\checkmark$ & $\checkmark$\\
 Spatial Structure component ($\phi$) & $\times$ & $\times$ & $\checkmark$\\
-Spatial structure variance ($\tau^2$) & $\times$ & $\times$ & $\checkmark$\\
-Additional scale ($\nu^2$) & $\times$ & $\times$ & $\checkmark$\\
-\addlinespace
+Spatial structure variance ($\sigma_{\phi}^2$) & $\times$ & $\times$ & $\checkmark$\\
+Error Term variance ($\sigma_{\varepsilon}^2$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
 Estimation method & MLE & MLE/GMM & MCMC\\
@@
 \endgroup{}
-
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
 
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 Following this, covariates were generated using \texttt{mvrnorm()} for
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 By employing artificially generated covariates and an artificially
@@
 \(\boldsymbol{\beta} = \begin{bmatrix} 1 & 0.9 & 0.7 & 0.5 & 0.3 \end{bmatrix}\),
-\(\nu^2 = 0.2\), \(\tau^2 = 0.6\) and \(\rho = 0.6\). Whereas beta
-values for GWR dataset is generated using a beta function, which is a
-function of coordinate points. The matrix \(\mathbf{W}\) was derived
-from the neighbourhood structure of an artificial study region using the
-functions \texttt{poly2nb()} and \texttt{nb2mat()}. Data generation for
-\(SAR + CAR\) employs a slightly different strategy. In addition to
-being generated per area simultaneously using the joint distribution of
-\(\mathbf{Y}\) as in Equation~\ref{eq-jointysar}, information about the
-spatial effect \(\boldsymbol{\phi}\) is also added at the end of the
-process. Therefore, we refer to it as \(SAR + CAR\) data.
+\(\sigma_{\phi}^2 = 0.2\), \(\sigma_{\phi}^2 = 0.6\) and \(\rho = 0.6\).
+Whereas beta values for GWR dataset is generated using a beta function,
+which is a function of coordinate points. The matrix \(\mathbf{W}\) was
+derived from the neighbourhood structure of an artificial study region
+using the functions \texttt{poly2nb()} and \texttt{nb2mat()}. Data
+generation for \(SAR + CAR\) employs a slightly different strategy. In
+addition to being generated per area simultaneously using the joint
+distribution of \(\mathbf{Y}\) as in Equation~\ref{eq-jointysar},
+information about the spatial effect \(\boldsymbol{\phi}\) is also added
+at the end of the process. Therefore, we refer to it as \(SAR + CAR\)
+data.
 
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 Moreover, Table~\ref{tbl-rmse} provide insights into model accuracy at
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
-Lombok includes five districts (kota/kabupaten), 53 sub-districts
-(kecamatan), and 608 villages (desa/kelurahan), which represent three
-successive levels of administrative division. In this study, the
-sub-district level is used as the spatial unit of analysis, as
-illustrated in Figure~\ref{fig-lombokmap}. The key variables of interest
-include property price, land size, built-up area, number of bedrooms,
-and number of bathrooms.
+Lombok comprises five districts (kota/kabupaten), 53 sub-districts
+(kecamatan), and 608 villages (desa/kelurahan), representing three
+successive levels of administrative division. Mataram District, despite
+its small area compared to other districts in Lombok, is the most
+urbanized and economically active area on the island, characterized by
+dense residential clusters with relatively homogeneous housing
+specifications. In adjacent districts such as West Lombok and Central
+Lombok, housing developments have expanded, particularly to accommodate
+commuters working in Mataram. The coastal areas of West Lombok,
+including Batu Layar and Sekotong, exhibit villa-style housing commonly
+associated with tourism, resulting in generally higher property prices.
+A similar pattern is found in North Lombok, where mountainous terrain
+limits development, but tourism-driven demand has led to villa-style
+housing near popular destinations. East Lombok, by contrast, remains
+more isolated due to its distance from Mataram. Housing in this district
+tends to serve primarily the local population, with limited signs of
+external investment or tourism-driven development. Additionally, the
+southern part of Central Lombok (Pujut) has seen emerging residential
+and real estate activity following the construction of the international
+MotoGP circuit.
+
+In this study, the sub-district level is used as the spatial unit of
+analysis, as illustrated in Figure~\ref{fig-lombokmap}. The key
+variables of interest include property price, land size, built-up area,
+number of bedrooms, and number of bathrooms.
 
@@
 refined to comprise a total of 576 observations. A summary of the key
-variables is presented in Table~\ref{tbl-datasumm}.
+variables is presented in Table~\ref{tbl-datasumm}. Notably, the
+distribution of property prices is skewed and does not exhibit a
+bell-shaped curve, which motivates the inclusion of a log-transformed
+price variable in the analysis.
 
@@
 
-\begin{longtable}[t]{lrrrrrrrr}
-
-\caption{\label{tbl-datasumm}Summary statistics of variables in the
-property price dataset}
+\begin{longtable}[t]{lrrrrrrr}
+
+\caption{\label{tbl-datasumm}Summary statistics of house prices dataset
+(\(N = 576\), \texttt{NA} = 4\%), including variables such as prices and
+its logscale value, land area(sqm), building area(sqm), number of beds,
+and number of baths}
 
@@
 \toprule
-Variable & Mean & Median & $Q_1$ & $Q_3$ & Min & Max & SD & NA\\
+Variable & Mean & SD & Median & $Q_1$ & $Q_3$ & Min & Max\\
 \midrule
-Prices (million IDR) & 1983.00 & 1400 & 800.0 & 2500.0 & 170 & 22500 & 2056.81 & 26\\
-Land-size (sqm) & 299.60 & 255 & 147.2 & 407.2 & 90 & 800 & 179.39 & 24\\
-Built-up area (sqm) & 184.80 & 150 & 100.0 & 210.0 & 70 & 600 & 109.20 & 24\\
-No. of bedroom & 3.24 & 3 & 2.0 & 4.0 & 1 & 6 & 1.19 & 24\\
-No. of bathroom & 2.55 & 2 & 2.0 & 3.0 & 1 & 5 & 1.05 & 24\\
+Prices (million IDR) & 1979.20 & 2054.51 & 1400.00 & 800.00 & 2500.00 & 170.00 & 22500.00\\
+logPrices & 7.26 & 0.80 & 7.24 & 6.68 & 7.82 & 5.14 & 10.02\\
+Land-size (sqm) & 299.08 & 179.33 & 254.50 & 147.00 & 405.75 & 90.00 & 800.00\\
+Built-up area (sqm) & 184.46 & 109.14 & 150.00 & 100.00 & 210.00 & 70.00 & 600.00\\
+No. of bedroom & 3.23 & 1.19 & 3.00 & 2.00 & 4.00 & 1.00 & 6.00\\
+\addlinespace
+No. of bathroom & 2.54 & 1.05 & 2.00 & 2.00 & 3.00 & 1.00 & 5.00\\
 \bottomrule
@@
 
-To address missing values, we employed the \texttt{mice} package (Van
-Buuren and Groothuis-Oudshoorn (2011)) to perform multiple imputation
-with parameters set to \(maxit = 10\), \(m = 10\), and a fixed random
-seed to ensure reproducibility. The imputed values were then
-reintegrated into the main dataset, replacing the original missing
-entries. This process resulted in a final dataset that was complete and
-ready for subsequent analysis using the selected spatial models.
+To address missing values, we used the \texttt{mice} package (Van Buuren
+and Groothuis-Oudshoorn (2011)), which performs multiple imputation by
+chained equations (MICE). This approach iteratively fills in missing
+data by modeling each variable with missing values as a function of
+other variables in the dataset. By generating multiple imputed datasets,
+it accounts for the uncertainty inherent in missing data. In our case,
+we specified 10 iterations and generated 10 imputed datasets. A fixed
+random seed was used to ensure the results were reproducible. The
+imputed datasets were then pooled to obtain final estimates for
+subsequent analyses. The imputed values were then reintegrated into the
+main dataset, replacing the original missing entries. This process
+resulted in a final dataset that was complete and ready for subsequent
+analysis using the selected spatial models.
 
@@
 
+Moreover, as GWR requires spatial coordinates for each observation, we
+addressed this by generating random points around the centroid of each
+sub-district using the \texttt{st\_sample()} function. The number of
+points generated matched the number of observations in each area, and
+each was positioned within a 0.5 km radius of the centroid to maintain a
+realistic spatial distribution while preserving the local context. This
+strategy allowed us to simulate plausible spatial locations for property
+transactions in the absence of precise geolocation data. Importantly, we
+applied the same bandwidth in the GWR model fitting to ensure
+consistency. By preserving the area-level grouping structure, this
+approach supports a fair comparison with the previous method.
+
 \subsection{Model Implementation \&
@@
 
-The cleaned and completed Lombok dataset was then model. The tables
-below present the results of the data fitting, offering a comparative
-view of each model's effectiveness and accuracy in capturing the
-dataset's characteristics.
+The cleaned and completed Lombok dataset was then fitted to the model.
+Equation Equation~\ref{eq-lopreg} presents the proposed model
+specification, which serves as the fixed formula applied consistently
+across all analyses
 
 \begin{equation}\phantomsection\label{eq-lopreg}{
-prices \sim land_{std} + building_{std} + beds_{std} + baths_{std}
+\text{logprices} \sim \text{land}_{\text{std}} + \text{building}_{\text{std}} + \text{beds}_{\text{std}} + \text{baths}_{\text{std}}
 }\end{equation}
@@
 coefficients are on a comparable scale, facilitating meaningful
-interpretation and comparison across different modeling approaches. The
-results presented in Table~\ref{tbl-modelcomp} provide a comprehensive
-evaluation of each model's performance in capturing both the structural
-and spatial characteristics of the property market in Lombok.
-
-We conducted model fitting on the area-level data, with detailed results
+interpretation and comparison across different modeling approaches. We
+conducted model fitting on the area-level data, with detailed results
 provided in the Appendix (Table~\ref{tbl-modelcomp1}). However, this
@@
 to apply three different models: GLMM, GWR, and multilevel CAR.
+
+The results presented in Table~\ref{tbl-modelcomp} provide a
+comprehensive evaluation of each model's performance in capturing both
+the structural and spatial characteristics of the property market in
+Lombok.
 
@@
 \multicolumn{4}{l}{\textbf{Variances}}\\
-\hspace{1em}$\nu^2$ & 0.22 [0.19, 0.25]* & 0.33 [0.29, 0.37]* & 0.22 [0.20, 0.25]*\\
-\hspace{1em}$\tau^2$ & 0.20 [0.11, 0.34]* & - & 0.31 [0.14, 0.58]*\\
+\hspace{1em}$\sigma_{\varepsilon}^2$ & 0.22 [0.19, 0.25]* & 0.33 [0.29, 0.37]* & 0.22 [0.20, 0.25]*\\
+\hspace{1em}$\sigma_{\phi}^2$ & 0.20 [0.11, 0.34]* & - & 0.31 [0.14, 0.58]*\\
 \addlinespace[0.3em]
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 Here, Table~\ref{tbl-modelcomp} compares parameter estimates and fit
@@
 \(\rho\), with a significant estimate of 0.48, showing the model's
-ability to account for spatial dependence. Variance components \(\nu^2\)
-(residual) and \(\tau^2\) (spatial) are also detailed, with both
-variance terms significant. Fit criteria (AIC, WAIC, and Log-likelihood)
-suggest mlvCAR and GWR may better capture local spatial variation, with
-mlvCAR balancing spatial structure and precision.
+ability to account for spatial dependence. Variance components
+\(\sigma_{\varepsilon}^2\) (residual) and \(\sigma_{\phi}^2\) are also
+detailed, with both terms are significant. Fit criteria (AIC, WAIC, and
+Log-likelihood) suggest mlvCAR and GWR may better capture local spatial
+variation, with mlvCAR balancing spatial structure and precision.
 
@@
 
-\subcaption{\label{fig-pricenphi1}phi values}
+\subcaption{\label{fig-pricenphi1}\(\phi\) values}
 
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 Figure~\ref{fig-pricenphi1} and Figure~\ref{fig-pricenphi2} respectively
@@
 spatial random effects (\(\phi\)) from the CAR model and the average
-log-transformed house prices per sub-district accross Lombok island.
-
-Here, the \(\phi\) values represent the spatially structured residuals
-not explained by the observed covariates in the CAR model. Specifically,
-it captures latent spatial influences such as neighbourhood
-desirability, access to infrastructure, or even local policy, that
-affect house prices beyond what structural features like land and
-building size can account for.
+log-transformed house prices per sub-district accross Lombok island. It
+can be seen that both plots in Figure~\ref{fig-pricenphi} exhibit
+similar spatial patterns, with high and low values clustering in
+comparable regions. Note that they do not have to be identical, as
+\(\phi\) represents the spatially structured residual component after
+accounting for covariates, whereas the log-transformed prices reflect
+the observed aggregated market values. Nonetheless, their resemblance
+suggests that some of the spatial variability in house prices is being
+effectively captured by the spatial random effects.
 
@@
 covariates, properties in that area are estimated to be approximately
-\(exp(0.57) \approx 1.76\) or 76\% more expensive than average.
+\(\exp(0.57) \approx 1.76\) or 76\% more expensive than average.
 Similarly, a sub-district with \(\phi = -0.2\), this suggests that
 properties in that area are approximately 18\% less expensive than
-average, since \(exp(-0.2) \approx 0.82\).
+average, since \(\exp(-0.2) \approx 0.82\). Note that the 95\% credible
+interval of \(\phi\) can be interpreted similarly by exponentiating the
+lower and upper bounds, yielding a multiplicative uncertainty range for
+the price deviation.
 
@@
 applications where local spatial variations are critical. The findings
-align with previous studies indicating that CAR models are better suited
-for managing spatial autocorrelation and heterogeneity, especially when
-spatial dependencies are strong.
+align with previous studies (Y. Wang and Kockelman (2013), Soroori,
+Moghaddam, and Salehi (2019), Zeng and Huang (2014), Guelat and Kéry
+(2018)) indicating that CAR models are better suited for managing
+spatial autocorrelation and heterogeneity, especially when spatial
+dependencies are strong.
 
@@
 
-These findings echo the limitations of GWR when precise coordinates are
-lacking, affirming the relevance of area-level approaches. Moreover,
+These findings also echo the limitations of GWR when precise coordinates
+are lacking, affirming the relevance of area-level approaches. Moreover,
 while we use a binary adjacency matrix for consistency, we recognise
@@
 \end{itemize}
+
+\newpage
 
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 \begin{table}
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 \subsection*{Model Comparison for area-level
@@
 \multicolumn{4}{l}{\textbf{Variances}}\\
-\hspace{1em}$\nu^2$ & 0.31 [0.21, 0.45]* & 0.27 [0.19, 0.41]* & 0.28 [0.01, 0.44]*\\
-\hspace{1em}$\tau^2$ & 0.20 [0.11, 0.34]* & - & 0.04 [0.01, 0.91]*\\
+\hspace{1em}$\sigma_{\epsilon}^2$ & 0.31 [0.21, 0.45]* & 0.27 [0.19, 0.41]* & 0.28 [0.01, 0.44]*\\
+\hspace{1em}$\sigma_{\phi}^2$ & 0.20 [0.11, 0.34]* & - & 0.04 [0.01, 0.91]*\\
 \addlinespace[0.3em]
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 \subsection*{Example of beta-values in GWR
@@
 
-\textsubscript{Source:
-\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
-Notebook}}
-
 \section*{References}\label{references}
@@
 
+\bibitem[\citeproctext]{ref-Guelat2018Effects}
+Guelat, J., and M. Kéry. 2018. {``Effects of Spatial Autocorrelation and
+Imperfect Detection on Species Distribution Models.''} \emph{Methods in
+Ecology and Evolution} 9: 1614--25.
+\url{https://doi.org/10.1111/2041-210X.12983}.
+
 \bibitem[\citeproctext]{ref-lee2013carbayes}
@@
 
+\bibitem[\citeproctext]{ref-Soroori2019Application}
+Soroori, Emad, Abolfazl Mohammadzadeh Moghaddam, and M. Salehi. 2019.
+{``Application of Local Conditional Autoregressive Models for
+Development of Zonal Crash Prediction Models and Identification of Crash
+Risk Boundaries.''} \emph{Transportmetrica A: Transport Science} 15:
+1102--23. \url{https://doi.org/10.1080/23249935.2018.1564801}.
+
 \bibitem[\citeproctext]{ref-stern1999inference}
@@
 
+\bibitem[\citeproctext]{ref-Wang2013A}
+Wang, Yiyi, and K. Kockelman. 2013. {``A Poisson-Lognormal
+Conditional-Autoregressive Model for Multivariate Spatial Analysis of
+Pedestrian Crash Counts Across Neighborhoods.''} \emph{Accident;
+Analysis and Prevention} 60: 71--84.
+\url{https://doi.org/10.1016/j.aap.2013.07.030}.
+
 \bibitem[\citeproctext]{ref-yang2019does}
@@
 
+\bibitem[\citeproctext]{ref-Zeng2014Bayesian}
+Zeng, Q., and Helai Huang. 2014. {``Bayesian Spatial Joint Modeling of
+Traffic Crashes on an Urban Road Network.''} \emph{Accident; Analysis
+and Prevention} 67: 105--12.
+\url{https://doi.org/10.1016/j.aap.2014.02.018}.
+
 \end{CSLReferences}

### 2025-05-06 manuscript.qmd
update manuscript and response

--- 2025-05-05 docs/manuscript.qmd
@@
 # Study region
+
 In the spatial modelling, the study region $\mathcal{S}$ is typically divided into $K$ distinct non-overlapping geographic units, denoted as $\mathcal{S}_k$ for $k = 1, 2, \dots, K$.
@@
 $$
+\mathbf{W} =
 \begin{array}{cccccc}
@@
 In GWR, this matrix assists in locally calibrating the model by reflecting the density and connectivity of regions (@brunsdon1996geographically, @stewart2018localized).
-Here, $\boldsymbol{D}$ is represented as:
-
-$$
-\begin{array}{cccccc}
-   & A & B & C & D & E \\
-A & 4 & 0 & 0 & 0 & 0\\ 
-B & 0 & 2 & 0 & 0 & 0\\ 
-C & 0 & 0 & 3 & 0 & 0\\
-D & 0 & 0 & 0 & 3 & 0\\
-E & 0 & 0 & 0 & 0 & 2\\
-\end{array}
-$$
-
-This matrix format indicates that region $A$ has 4 neighbour, regions $C$ and $D$ each have 3 neighbours, and region $B$ and $E$ has 2 neighbour.
+In this case, $\mathbf{D}$ indicates that region $A$ has 4 neighbours, regions $C$ and $D$ each have 3 neighbours, and region $B$ and $E$ has 2 neighbours.
 
@@
 This figure illustrates the spatial connectivity between polygons in the artificial study region, where each line represents a defined neighbour relationship based on spatial adjacency"
-#| out-width: "70%"
+#| out-width: "60%"
 #| fig-align: "center"
@@
 # Spatial regression model
+
 When discussing spatial regression, it's crucial to comprehend the basic notion of linear regression.
@@
 y = \beta_0 + \beta_1x_1 + \dots + \beta_p x_p + \varepsilon
-$${#eq-basicreg}
+$$ {#eq-basicreg}
 
@@
 y_i(s) = \beta_{0}(s) + \sum_{k=1}^p \beta_k(s)x_{ik}(s) + \varepsilon(s), \\ i = 1,\dots, n
-$${#eq-gwr}
+$$ {#eq-gwr}
 
@@
 \beta(s) = (\boldsymbol{x}^T \boldsymbol{W}(s) \boldsymbol{x})^{-1} \boldsymbol{x}^T \boldsymbol{W}(s) \boldsymbol{y}
-$${#eq-betagwr}
+$$ {#eq-betagwr}
 
@@
 ## Simultaneously Autoregressive (SAR) Models
+
 While GWR focuses on capturing localised spatial heterogeneity by allowing coefficients to vary across space, SAR take a different approach by explicitly modelling spatial dependencies through a spatial lag.
@@
 y(s) = \rho \sum_{s'} w(s,s') y(s') + \sum_{k=1}^p \beta_k x_k(s) + \varepsilon(s)
-$${#eq-sar}
+$$ {#eq-sar}
 
@@
 \mathbf{Y} \sim \mathcal{N}\left( \left( I - \rho W \right)^{-1} X\beta, \, \sigma^2 \left( I - \rho W \right)^{-1} \left( I - \rho W^T \right)^{-1} \right)
-$${#eq-jointysar}
+$$ {#eq-jointysar}
 
@@
 y(s) = \sum_{k=1}^p \beta_k x_k(s) + \varepsilon(s) + \phi(s) 
-$${#eq-car}
+$$ {#eq-car}
 
-Here, $\boldsymbol{\varepsilon} \sim N(0, \nu^2)$ while $\boldsymbol{\phi}$ is a specific component in CAR model that has a role as spatial effect.
+Here, $\boldsymbol{\varepsilon} \sim N(0, \sigma_{\varepsilon}^2)$ while $\boldsymbol{\phi}$ is a specific component in CAR model that has a role as spatial effect.
 It is also common to mention $\boldsymbol{\phi}$ as a CAR priors.
@@
 $$
-\boldsymbol{\phi}\sim N(0,\tau^2\boldsymbol{Q}^{-1}) 
-$${#eq-phidist}
+\boldsymbol{\phi}\sim N(0,\sigma_{\phi}^2\boldsymbol{Q}^{-1}) 
+$$ {#eq-phidist}
 
@@
 $$
-\phi_k|\boldsymbol\phi_{-K}, \boldsymbol{W},\rho, \tau^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\tau^2}{A}\right)
-$${#eq-phicar}
+\phi_k|\boldsymbol{\phi}_{-k}, \mathbf{W},\rho, \sigma_{\phi}^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\sigma_{\phi}^2}{A}\right)
+$$ {#eq-phicar}
 
@@
 y_i(s) = \sum_{k=1}^p \beta_k x_{ki}(s) + \varepsilon_i(s) + \phi(s)
-$${#eq-ycar}
+$$ {#eq-ycar}
 
-The Bayesian approach to CAR models entails defining prior distributions for all model parameters, such as the regression coefficients $\beta$, variance parameters $\nu^2$ and $\tau^2$, and spatial random effects $\phi_k$.
+The Bayesian approach to CAR models entails defining prior distributions for all model parameters, such as the regression coefficients $\beta$, variance parameters $\sigma_{\varepsilon}^2$ and $\sigma_{\phi}^2$, and spatial random effects $\phi_k$.
 Subsequently, Bayesian inference techniques like Markov Chain Monte Carlo (MCMC) are employed to derive posterior distributions of these parameters.
@@
       "Covariate Coefficients ($\\beta$)", 
-      "Error Term variance ($\\sigma^2$)", 
       "Spatial Autocorrelation ($\\rho$)", 
       "Spatial Structure component ($\\phi$)", 
-      "Spatial structure variance ($\\tau^2$)", 
-      "Additional scale ($\\nu^2$)",
+      "Spatial structure variance ($\\sigma_{\\phi}^2$)", 
+      "Error Term variance ($\\sigma_{\\varepsilon}^2$)",
       "Estimation method"
     ),
-    GWR = c("$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\times$", "$\\times$", "$\\times$", "MLE"),
-    SAR = c("$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\times$", "$\\times$", "MLE/GMM"),
-    CAR = c("$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "MCMC")
+    GWR = c("$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\times$", "$\\times$", "$\\checkmark$", "MLE"),
+    SAR = c("$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\times$", "$\\times$", "$\\checkmark$", "MLE/GMM"),
+    CAR = c("$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$",  "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "$\\checkmark$", "MCMC")
   ),
@@
 By employing artificially generated covariates and an artificially constructed study region, we developed three distinct property price datasets, each based on a different spatial model: CAR, a combination of CAR and SAR, and GWR.
-The true parameter values for this purpose are $\boldsymbol{\beta} = \begin{bmatrix} 1 & 0.9 & 0.7 & 0.5 & 0.3 \end{bmatrix}$, $\nu^2 = 0.2$, $\tau^2 = 0.6$ and $\rho = 0.6$.
+The true parameter values for this purpose are $\boldsymbol{\beta} = \begin{bmatrix} 1 & 0.9 & 0.7 & 0.5 & 0.3 \end{bmatrix}$, $\sigma_{\phi}^2 = 0.2$, $\sigma_{\phi}^2 = 0.6$ and $\rho = 0.6$.
 Whereas beta values for GWR dataset is generated using a beta function, which is a function of coordinate points.
@@
 
-::: {#fig-biasplot layout-ncol=2}
-
+::: {#fig-biasplot layout-ncol="2"}
 ![Bias values by method and dataset type for area-level data](figures/bias1_plot.png){#fig-biasplot1}
@@
 
-Lombok includes five districts (kota/kabupaten), 53 sub-districts (kecamatan), and 608 villages (desa/kelurahan), which represent three successive levels of administrative division.
+Lombok comprises five districts (kota/kabupaten), 53 sub-districts (kecamatan), and 608 villages (desa/kelurahan), representing three successive levels of administrative division.
+Mataram District, despite its small area compared to other districts in Lombok, is the most urbanized and economically active area on the island, characterized by dense residential clusters with relatively homogeneous housing specifications.
+In adjacent districts such as West Lombok and Central Lombok, housing developments have expanded, particularly to accommodate commuters working in Mataram.
+The coastal areas of West Lombok, including Batu Layar and Sekotong, exhibit villa-style housing commonly associated with tourism, resulting in generally higher property prices.
+A similar pattern is found in North Lombok, where mountainous terrain limits development, but tourism-driven demand has led to villa-style housing near popular destinations.
+East Lombok, by contrast, remains more isolated due to its distance from Mataram.
+Housing in this district tends to serve primarily the local population, with limited signs of external investment or tourism-driven development.
+Additionally, the southern part of Central Lombok (Pujut) has seen emerging residential and real estate activity following the construction of the international MotoGP circuit.
+
 In this study, the sub-district level is used as the spatial unit of analysis, as illustrated in @fig-lombokmap.
@@
 A summary of the key variables is presented in @tbl-datasumm.
+Notably, the distribution of property prices is skewed and does not exhibit a bell-shaped curve, which motivates the inclusion of a log-transformed price variable in the analysis.
 
@@
 #| label: tbl-datasumm
-#| tbl-cap: "Summary statistics of variables in the property price dataset"
+#| tbl-cap: "Summary statistics of house prices dataset ($N = 576$, `NA` = 4%), including variables such as prices and its logscale value, land area(sqm), building area(sqm), number of beds, and number of baths"
 #| eval: true
+
+load("data/propdata.RData")
+load("data/hp_dat_v1.RData")
+
+propdata <- propdata %>%
+  select(-village, -floors, -furnished, -category) %>%
+  rbind(hp_dat_v1) %>%
+  filter(land >= 90 & land <= 800) %>%
+  filter(building >= 70 & building <= 600) %>%
+  filter(beds <= 6) %>%
+  filter(baths > 0 & baths <= 5) %>% 
+  mutate(
+    prices = prices * 1000,
+    logprices = log(prices)
+    )
+
+dat <- propdata %>% 
+  select(prices, logprices, land, building, beds, baths)
 
@@
 data <- data.frame(
-  Variable = c("Prices (million IDR)", "Land-size (sqm)", "Built-up area (sqm)", 
+  Variable = c("Prices (million IDR)", "logPrices", "Land-size (sqm)", "Built-up area (sqm)",
                "No. of bedroom", "No. of bathroom"),
-  Mean   = round(c(1983, 299.6, 184.8, 3.24, 2.55), 2),
-  Median = round(c(1400, 255.0, 150.0, 3.00, 2.00), 2),
-  Q1     = round(c(800, 147.2, 100.0, 2.00, 2.00), 2),
-  Q3     = round(c(2500, 407.2, 210.0, 4.00, 3.00), 2),
-  Min    = round(c(170, 90.0, 70.0, 1.00, 1.00), 2),
-  Max    = round(c(22500, 800.0, 600.0, 6.00, 5.00), 2),
-  SD     = round(c(2056.81, 179.39, 109.20, 1.19, 1.05), 2),
-  `NA`   = c(26, 24, 24, 24, 24)
+  Mean   = round(colMeans(dat, na.rm = TRUE), 2),
+  SD     = round(apply(dat, 2, sd, na.rm = TRUE), 2),
+  Median = round(apply(dat, 2, median, na.rm = TRUE), 2),
+  Q1     = round(apply(dat, 2, quantile, probs = 0.25, na.rm = TRUE), 2),
+  Q3     = round(apply(dat, 2, quantile, probs = 0.75, na.rm = TRUE), 2),
+  Min    = round(apply(dat, 2, min, na.rm = TRUE), 2),
+  Max    = round(apply(dat, 2, max, na.rm = TRUE), 2)
 )
@@
 kbl(data,
-         col.names = c("Variable", "Mean", "Median", 
-                       "$Q_1$", "$Q_3$", "Min", "Max", "SD", "NA"),
+    col.names = c("Variable", "Mean", "SD", "Median", "$Q_1$", "$Q_3$", "Min", "Max"),
+    row.names = FALSE,
     escape = FALSE,
@@
 
-To address missing values, we employed the `mice` package (@van2011mice) to perform multiple imputation with parameters set to $maxit = 10$, $m = 10$, and a fixed random seed to ensure reproducibility.
+To address missing values, we used the `mice` package (@van2011mice), which performs multiple imputation by chained equations (MICE).
+This approach iteratively fills in missing data by modeling each variable with missing values as a function of other variables in the dataset.
+By generating multiple imputed datasets, it accounts for the uncertainty inherent in missing data.
+In our case, we specified 10 iterations and generated 10 imputed datasets.
+A fixed random seed was used to ensure the results were reproducible.
+The imputed datasets were then pooled to obtain final estimates for subsequent analyses.
 The imputed values were then reintegrated into the main dataset, replacing the original missing entries.
@@
 
+Moreover, as GWR requires spatial coordinates for each observation, we addressed this by generating random points around the centroid of each sub-district using the `st_sample()` function.
+The number of points generated matched the number of observations in each area, and each was positioned within a 0.5 km radius of the centroid to maintain a realistic spatial distribution while preserving the local context.
+This strategy allowed us to simulate plausible spatial locations for property transactions in the absence of precise geolocation data.
+Importantly, we applied the same bandwidth in the GWR model fitting to ensure consistency.
+By preserving the area-level grouping structure, this approach supports a fair comparison with the previous method.
+
 ## Model Implementation & Analysis
 
-The cleaned and completed Lombok dataset was then model.
-The tables below present the results of the data fitting, offering a comparative view of each model’s effectiveness and accuracy in capturing the dataset's characteristics.
+The cleaned and completed Lombok dataset was then fitted to the model.
+Equation @eq-lopreg presents the proposed model specification, which serves as the fixed formula applied consistently across all analyses
 
 $$
-prices \sim land_{std} + building_{std} + beds_{std} + baths_{std}
-$${#eq-lopreg}
+\text{logprices} \sim \text{land}_{\text{std}} + \text{building}_{\text{std}} + \text{beds}_{\text{std}} + \text{baths}_{\text{std}}
+$$ {#eq-lopreg}
 
 Standardization of the predictors ensures that the resulting coefficients are on a comparable scale, facilitating meaningful interpretation and comparison across different modeling approaches.
-The results presented in @tbl-modelcomp provide a comprehensive evaluation of each model’s performance in capturing both the structural and spatial characteristics of the property market in Lombok.
-
 We conducted model fitting on the area-level data, with detailed results provided in the Appendix (@tbl-modelcomp1).
@@
 To address this, we used point-level data to apply three different models: GLMM, GWR, and multilevel CAR.
+
+The results presented in @tbl-modelcomp provide a comprehensive evaluation of each model’s performance in capturing both the structural and spatial characteristics of the property market in Lombok.
 
@@
   "$\\rho$",                     "-",                      "-",                      "0.48 [0.07, 0.90]*",
-  "$\\nu^2$",                    "0.22 [0.19, 0.25]*",     "0.33 [0.29, 0.37]*",     "0.22 [0.20, 0.25]*",
-  "$\\tau^2$",                   "0.20 [0.11, 0.34]*",     "-",                      "0.31 [0.14, 0.58]*",
+  "$\\sigma_{\\varepsilon}^2$",  "0.22 [0.19, 0.25]*",     "0.33 [0.29, 0.37]*",     "0.22 [0.20, 0.25]*",
+  "$\\sigma_{\\phi}^2$",          "0.20 [0.11, 0.34]*",     "-",                      "0.31 [0.14, 0.58]*",
   "AIC",                         "803.38",                 "-645.104",               "786.95",
@@
 Here, @tbl-modelcomp compares parameter estimates and fit criteria for three models: GLMM, GWR, and multilevel CAR.
-Significant coefficients are marked with an asterisk (*).
+Significant coefficients are marked with an asterisk (\*).
 Key determinants like land-size and built-up area have positive, significant effects across models, while the number of bedrooms shows only minor, varying effects, indicating model-specific differences in capturing this variable's influence.
 Only the multilevel CAR model includes the spatial parameter $\rho$, with a significant estimate of 0.48, showing the model's ability to account for spatial dependence.
-Variance components $\nu^2$ (residual) and $\tau^2$ (spatial) are also detailed, with both variance terms significant.
+Variance components $\sigma_{\varepsilon}^2$ (residual) and $\sigma_{\phi}^2$ are also detailed, with both terms are significant.
 Fit criteria (AIC, WAIC, and Log-likelihood) suggest mlvCAR and GWR may better capture local spatial variation, with mlvCAR balancing spatial structure and precision.
 
-::: {#fig-pricenphi layout-ncol=2}
-
-![phi values](figures/phi_map.png){#fig-pricenphi1}
+::: {#fig-pricenphi layout-ncol="2"}
+![$\phi$ values](figures/phi_map.png){#fig-pricenphi1}
 
@@
 @fig-pricenphi1 and @fig-pricenphi2 respectively present spatial heatmaps illustrating the distribution of estimated spatial random effects ($\phi$) from the CAR model and the average log-transformed house prices per sub-district accross Lombok island.
-
-Here, the $\phi$ values represent the spatially structured residuals not explained by the observed covariates in the CAR model.
-Specifically, it captures latent spatial influences such as neighbourhood desirability, access to infrastructure, or even local policy, that affect house prices beyond what structural features like land and building size can account for.
+It can be seen that both plots in @fig-pricenphi exhibit similar spatial patterns, with high and low values clustering in comparable regions.
+Note that they do not have to be identical, as $\phi$ represents the spatially structured residual component after accounting for covariates, whereas the log-transformed prices reflect the observed aggregated market values.
+Nonetheless, their resemblance suggests that some of the spatial variability in house prices is being effectively captured by the spatial random effects.
 
 Because the model uses log-transformed prices, $\phi$ has multiplicative interpretation on the original price scale.
-For instance, if $\phi = 0.57$ for a given sub-district, then after accounting for covariates, properties in that area are estimated to be approximately $exp(0.57) \approx 1.76$ or 76% more expensive than average.
-Similarly, a sub-district with $\phi = -0.2$, this suggests that properties in that area are approximately 18% less expensive than average, since $exp(-0.2) \approx 0.82$.
+For instance, if $\phi = 0.57$ for a given sub-district, then after accounting for covariates, properties in that area are estimated to be approximately $\exp(0.57) \approx 1.76$ or 76% more expensive than average.
+Similarly, a sub-district with $\phi = -0.2$, this suggests that properties in that area are approximately 18% less expensive than average, since $\exp(-0.2) \approx 0.82$.
+Note that the 95% credible interval of $\phi$ can be interpreted similarly by exponentiating the lower and upper bounds, yielding a multiplicative uncertainty range for the price deviation.
 
@@
 The superior performance of the CAR model, particularly in datasets with strong spatial structure, suggests that it may be a more reliable choice for applications where local spatial variations are critical.
-The findings align with previous studies indicating that CAR models are better suited for managing spatial autocorrelation and heterogeneity, especially when spatial dependencies are strong.
+The findings align with previous studies (@Wang2013A, @Soroori2019Application, @Zeng2014Bayesian, @Guelat2018Effects) indicating that CAR models are better suited for managing spatial autocorrelation and heterogeneity, especially when spatial dependencies are strong.
 
@@
 
-These findings echo the limitations of GWR when precise coordinates are lacking, affirming the relevance of area-level approaches.
+These findings also echo the limitations of GWR when precise coordinates are lacking, affirming the relevance of area-level approaches.
 Moreover, while we use a binary adjacency matrix for consistency, we recognise that the choice of spatial weight $\mathbf{W}$ may influences model performance.
@@
 
-* Funding
-* Competing interests
+-   Funding
+-   Competing interests
 
@@
 
-* Ethics approval
-* Consent to participate
-* Consent for publication
-* Availability of data and materials
-* Code availability
-* Authors' contributions
+-   Ethics approval
+-   Consent to participate
+-   Consent for publication
+-   Availability of data and materials
+-   Code availability
+-   Authors' contributions
 
+\newpage
 # Appendix {.unnumbered}
@@
 
-## Model Comparison for area-level data{.unnumbered}
+## Model Comparison for area-level data {.unnumbered}
 
@@
   "$\\rho$",                     "-",                      "0.009[-0.005, 0.023]",    "0.39 [0.01, 0.91]*",
-  "$\\nu^2$",                    "0.31 [0.21, 0.45]*",     "0.27 [0.19, 0.41]*",     "0.28 [0.01, 0.44]*",
-  "$\\tau^2$",                   "0.20 [0.11, 0.34]*",     "-",                      "0.04 [0.01, 0.91]*",
+  "$\\sigma_{\\epsilon}^2$",     "0.31 [0.21, 0.45]*",     "0.27 [0.19, 0.41]*",     "0.28 [0.01, 0.44]*",
+  "$\\sigma_{\\phi}^2$",         "0.20 [0.11, 0.34]*",     "-",                      "0.04 [0.01, 0.91]*",
   "AIC",                         "-",                       "-",                          "-",
@@
 
-## Example of beta-values in GWR model{.unnumbered}
+## Example of beta-values in GWR model {.unnumbered}
 

### 2025-05-06 response/response.qmd
Update response

@@
 ---
-title: "Response to Journal's 2nd Reviewer Comments"
+title: "Response to Journal's Reviewer Comments"
 subtitle: "Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study"
-author: 
-  - name: Indira Puteri Kinasih
-    orcid: 0009-0009-6959-6723
-    email: indiraputeri@uinmataram.ac.id
 ---
+
+**Reviewer 1**
+
+**This manuscript compares different methods (Conditionally Autoregressive, Geographically Weighted Regression, and Simultaneous Autoregressive) for modeling property prices.
+The problem definition looks reasonable, but the manuscript presents only the application part; nothing is contributed toward the theory.
+Therefore, I will suggest transferring this manuscript to an application-related journal.
+To my understanding, this manuscript does not fall in the scope of this journal.
+Further, the manuscript is not standard format of the journal.**
+
+We acknowledge that our work is primarily application-driven, focusing on the use of Conditionally Autoregressive (CAR), Geographically Weighted Regression (GWR), and Simultaneous Autoregressive (SAR) models in the context of property price modeling.
+However, we respectfully submit that the manuscript makes a meaningful methodological contribution, particularly in bridging the gap between spatial theory and applied practice—an area of growing relevance, especially in spatial analysis for developing regions.
+
+While we do not introduce new theoretical constructs, our work extends the methodological discourse in the following ways:
+
+ - **Shifting the Dominant Modeling Paradigm**: Previous studies on house price modeling have predominantly employed Geographically Weighted Regression (GWR), which requires precise geolocation data for model fitting.
+However, such data are often incomplete, unavailable, or unreliable in many regions.
+Our study explores the potential of the Conditionally Autoregressive (CAR) model—more commonly applied in health and ecological studies—as a viable alternative for house price modeling.
+This cross-domain adaptation is underexplored and opens new methodological pathways for spatial econometrics.
+
+ - **Unified Comparative Framework**: We develop and apply a systematic framework to compare CAR, SAR, and GWR models using both simulated data and real-world property data.
+This comparative approach, particularly within the under-researched housing market context of a developing region, fills a critical gap in the literature.
+
+ - **Practical Implementation Insights**: The manuscript offers practical guidance for applied researchers on implementing, diagnosing, and interpreting each spatial model.
+These details are especially valuable for practitioners operating in data-scarce environments or policy-driven contexts.
+
+ - **Informing Model Selection**: By revealing the comparative strengths and limitations of each model in capturing spatial heterogeneity, our findings contribute to informed model selection in applied spatial econometrics—balancing methodological soundness with practical feasibility.
+
+ - **Policy and Planning Relevance**: The results of our study have implications for spatial planning and housing market analysis, offering a framework for integrating spatial models into decision-making processes where traditional approaches may be constrained by data availability.
+
+Given these points, we believe the manuscript fits within the journal’s broader mission to advance rigorous spatial analysis methods and encourage cross-disciplinary methodological applications.
+
+**Reviewer 2**
 
@@
     Thank you for this insightful comment regarding the specification of the spatial weight matrix, $W$.
-We fully acknowledge that the choice of $W$—whether binary adjacency, row-standardized, or distance-based—can substantially influence the performance and interpretation of spatial models.
-In the present study, our primary objective was to evaluate and compare the performance of CAR, SAR, and GWR models using both simulated data and real-world data from Lombok.
-To maintain consistency across models and to isolate the structural differences in model behavior, we have, at this stage, limited our analysis to a binary adjacency matrix for $W$.
+The choice of $W$—whether binary adjacency, row-standardized, or distance-based—can indeed influence the performance and interpretation of spatial models.
+In the present study, our primary objective was to evaluate and compare the performance of CAR, SAR, and GWR models using both simulated data and real-world data (Lombok house prices).
+To maintain consistency across models and to isolate the structural differences in model behavior, we have, at this stage, limited our analysis to a binary adjacency matrix for $W$ as a starting point.
 We agree that exploring alternative specifications of $W$ (e.g., row-standardized or distance-decay matrices) is a valuable direction for further research.
@@
 
-    Regarding sample size, prior studies suggest that having approximately 15–30 observations per area is generally sufficient for stable estimation in multilevel frameworks.
+    Regarding sample size, literature suggests that as long as the number of groups (spatial areas) is large enough, the model can still perform well even with a limited number of observations per group.
+See, for example, @maas2005sufficient and @rabe2008multilevel.
 Given our data structure aligns with this recommendation, we believe the mlvCAR model offers a robust approach, especially in real-world scenarios where heterogeneity and partial pooling across areas are important.
@@
 
-    In our study, we recognize that the value of $\rho$ plays a crucial role in determining the extent of spatial autocorrelation captured by the model.
-While we did not conduct a dedicated sensitivity analysis across a range of fixed $\rho$ values, our simulation framework inherently explores variation in $\rho$ through repeated data generation processes, in which $\rho$ is estimated as part of the model fitting.
-The observed variability in $\rho$ reflects the underlying spatial structure in the data rather than instability in the model.
+    Certainly, the $\rho$ parameter plays a crucial role in understanding the extent to which spatial autocorrelation is captured by the model.
+While we did not conduct a dedicated sensitivity analysis by fixing $\rho$ at multiple values, our simulation framework implicitly accounts for variability in $\rho$ through repeated data generation and model fitting, where $\rho$ is treated as an estimable parameter.
+The observed variability—such as the estimate of 0.48 in the multilevel CAR model—reflects the differences in underlying spatial structure across datasets rather than instability or inconsistency in the model itself.
 
-    That said, we agree that a more systematic sensitivity analysis—such as fitting the model under different fixed values of $\rho$—could offer further insights into model robustness.
-We consider this a valuable direction for future work, especially in scenarios with varying levels of spatial autocorrelation or data sparsity.
+    In a Bayesian framework, the prior distribution assigned to $\rho$ also has an important influence, particularly in data-sparse contexts or when spatial signals are weak.
+In our implementation, we used a uniform or weakly informative prior (depending on the model setting), allowing the data to drive the estimation of $\rho$ while still preventing extreme or implausible values.
+In practical applications, however, researchers may wish to incorporate more informative priors based on prior knowledge of spatial processes or domain-specific expectations, which can enhance stability and interpretability in certain contexts.
+
+    That said, we fully agree that a more systematic sensitivity analysis—such as fixing $\rho$ at various plausible levels or experimenting with different prior distributions—could offer deeper insights into model robustness and the influence of spatial structure.
+We consider this an important direction for future research, particularly in settings with heterogeneous spatial patterns or limited sample sizes, where sensitivity to $\rho$ may significantly impact conclusions and policy relevance.
 
@@
 
-    Thank you for highlighting the importance of interpreting the spatial effects $\phi$ in the CAR model.
-We agree that while $\phi$ captures unobserved spatial heterogeneity, its interpretability in the context of house price determinants requires careful consideration.
-Rather than directly reflecting specific covariates, $\phi$ accounts for spatially structured residual variation—such as omitted neighborhood characteristics, local amenities, or spatial spillovers—that are not captured by observed variables.
+    Thank you for raising this point.
+The $\phi$ parameter capture the spatial heterogeneity in the data that is not explained by the model.
+In other words, rather than directly reflecting specific covariates, $\phi$ accounts for spatially structured residual variation—such as omitted neighborhood characteristics, local amenities, or spatial spillovers—that are not captured by observed variables.
+In essence, it represents the "local" price deviations from the overall mean, adjusted for covariates.
 
-    To aid interpretation, we acknowledge that visualizing $\phi$ through spatial effect maps could be a powerful diagnostic tool.
-Such maps can reveal clusters of under- or over-prediction and provide insights into areas where additional covariates or local effects might be relevant.
-In the revised version of this study, we will include a Lombok sub-district-level map displaying the estimated $\phi$ values to better communicate the spatial structure captured by the model and to guide further interpretation and refinement.
+    For a brief illustration, since the model uses log-transformed prices, $\phi$ has multiplicative interpretation on the original price scale.
+So if $\phi = 0.57$ for a given sub-district, then after accounting for covariates, properties in that area are estimated to be approximately $\exp(0.57) \approx 1.76$ or 76% more expensive than average.
+Similarly, a sub-district with $\phi = -0.2$, this suggests that properties in that area are approximately 18% less expensive than average, since $\exp(-0.2) \approx 0.82$.
+Note that the 95% credible interval of $\phi$ can be interpreted similarly by exponentiating the lower and upper bounds, yielding a multiplicative uncertainty range for the price deviation.
+
+    To address this comment, we have amended the manuscript to include the interpretation of $\phi$.
+We also provide a map that visualizes the estimated spatial effects ($\phi$), thereby highlighting the spatial patterns identified by the model.
 
@@
 
-    Thank you for this thoughtful question.
     In our manuscript, we highlight Geographically Weighted Regression (GWR) as a prominent method in the literature, particularly due to its frequent application in house price studies.
-This observation is based on a review of previous works rather than a performance-based comparison.
+This observation is based on a review of previous works rather than a performance-based comparison, for instance @wang2020geographically, @lu2011geographically, @stewart2018localized, and @cellmer2020spatial.
 While GWR is indeed widely used for its ability to model spatial heterogeneity at the local level, our study goes further by systematically comparing GWR with CAR and SAR models through simulation and real data applications.
@@
 
-     We acknowledge that the description in the manuscript may not have been sufficiently clear.
+    Regretably the description in the manuscript is not clear.
+Our response is as follows.
 The original dataset obtained through web-scraping did include a broad range of property types—such as residential houses, villas, boarding houses, and vacant land.
-However, for the purposes of this study, we restricted our analysis to residential houses only.
-We will revise the manuscript accordingly to ensure that this scope is clearly stated in both the abstract and the data collection section.
+However, for the purposes of this study, we restricted our analysis to residential houses only, and thereby omitting all other types include commercial real estate.
+We have revised the manuscript accordingly to ensure that this is clearly stated in both the abstract and the data collection section.
 
@@
 
-    For clarity, we note that the Lombok house price dataset used in this study does not contain precise geographic coordinates for each observation.
-As a result, it is not possible to accurately display the distribution of the data points on a map.
+    To be clear, the Lombok house price dataset used in this study does not contain precise geographic X-Y coordinates for each observation.
 While each observation is linked to a sub-district, the absence of point-level location data limits our ability to visualize them spatially within those areas.
-Therefore, for a more informative visualization, we are preparing a heatmap of house prices at the sub-district level, which will reflect the spatial variation in prices across Lombok using aggregated data.
+One solution here is a choropleth map of house prices at the sub-district level, which will reflect the spatial variation in prices across Lombok using aggregated data.
+This has been added in the manuscript.
 
@@
     Thank you for this important observation.
-We clarify that GLMM and multilevel CAR models do not require explicit geographic coordinates for each individual observation.
-Instead, they rely on an area-level identifier (ID) that links each observation to a corresponding spatial unit (e.g., sub-district).
+To clarify, GLMM and multilevel CAR models do not require explicit geographic coordinates for each individual observation.
+Instead, they rely on an area-level identifiers that links each observation to a corresponding spatial unit (e.g., sub-district).
+This makes it possible for these models to estimate the group-level variability.
 
-    In contrast, GWR requires spatial coordinates for each observation.
-To meet this requirement, we generated random points around the centroid of each sub-district, using `st_sample()`, where the number of points corresponds to the number of observations in that area.
+    In contrast, GWR does indeed require spatial coordinates for each observation.
+To meet this requirement, we generated random points around the centroid of each sub-district where the number of points corresponds to the number of observations in that area.
+This centroid-based method is also mentioned in @fotheringham2002geographically.
 Each point was placed within a 0.5 km radius from the centroid, ensuring a reasonable spatial spread while maintaining the area-specific context.
 This approach enabled us to simulate realistic spatial positions for property transactions in the absence of precise geolocation data.
+Note that, the same bandwith was used in the fitting of the GWR model.
+By enforcing the area-level grouping, this ensures a fair comparison to the previous method.
 
@@
 
-    In the revised version of the manuscript, we have made sure to explicitly refer to Table 5 in the main text to improve clarity and consistency.
-We appreciate your careful reading and helpful suggestion.
+    Thank you for raising this point. In the revised version of the manuscript, we have made sure to explicitly refer to Table 5 in the main text to improve clarity and consistency.
 
@@
 
+    Thank you for this important comment.
 As we noted in our response to Comment 8, the Lombok house price data do not come with precise geographic coordinates.
-In order to apply the `gwr()` function, which requires point-level location data, we generated synthetic coordinates by simulating random points around the centroid of each sub-district.
-While this workaround allowed us to run the GWR model, we acknowledge that it is not an ideal or practical solution for real-world applications.
-This limitation reinforces our argument that point-level models such as GWR can be challenging to apply meaningfully when precise location data are not available.
+In order to apply the GWR model (which requires point-level location data) we generated synthetic coordinates by simulating random points around the centroid of each sub-district.
+It’s important to note that while using centroids is a common practice, there are limitations.
+For instance, in cases where the population or activity within a spatial area is not uniformly distributed, the geometric centroid may not accurately represent the spatial characteristics of the area.
+Limitations such as these reinforces our argument that point-level models such as GWR can be challenging to apply meaningfully when precise location data are not available.
 
@@
     As mentioned in our response to Comment 1, in this study we used a binary adjacency matrix to define spatial relationships, primarily to maintain consistency across the CAR and SAR models and to focus on comparing their general performance.
-We acknowledge that the choice of spatial weights matrix ($W$) can significantly influence model outcomes.
+We acknowledge that the choice of spatial weights matrix ($W$) may influence model outcomes.
 While we have not conducted a sensitivity analysis in this current version, we agree that testing alternative specifications—such as row-standardized or distance-based matrices—would be valuable.
@@
     Thank you for this important point.
-Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater.
-In contrast, there is relatively limited research focused on low-density or semi-rural contexts such as Lombok.
+Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater [see, e.g., @gamal2024urban;@berawi2019land;@han2004spatial].
+In contrast, there is relatively limited research focused on low-density, semi-rural, and small-economy contexts such as Lombok [@jamil2025spatiotemporal].
 Our study aims to address this gap by evaluating the applicability and performance of spatial models, particularly the CAR model, in a setting that has different data characteristics and spatial structures compared to major metropolitan areas.
 
-    We also acknowledge that in richer datasets—common in urban contexts where more granular, high-frequency, and multidimensional data are available—machine learning approaches may offer competitive or even superior predictive performance.
+    We also acknowledge that in richer datasets—common in large ubranised regions where more granular, high-frequency, and multidimensional data are available—machine learning approaches may offer competitive or even superior predictive performance.
 However, our focus here is to assess interpretable spatial models in contexts where data limitations remain a challenge.
+
+# References

### 2025-05-06 response/response.qmd
Update response

@@
 
-1.  **The choice of $W$ (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?**
+1.  **The choice of** $W$ (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?
 
@@
 3.
-**The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model).
-Have the authors examined how sensitive the model results are to changes in $\rho$?**
+**The estimated value of** $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model).
+Have the authors examined how sensitive the model results are to changes in $\rho$?
 
@@
 4.
-**The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants?
-Would spatial effect maps or additional diagnostic tools enhance this understanding?**
+**The study discusses the spatial effects** $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants?
+Would spatial effect maps or additional diagnostic tools enhance this understanding?
 
@@
     Thank you for this important point.
-Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater [see, e.g., @gamal2024urban;@berawi2019land;@han2004spatial].
+Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater [see, e.g., @gamal2024urban; @berawi2019land; @han2004spatial].
 In contrast, there is relatively limited research focused on low-density, semi-rural, and small-economy contexts such as Lombok [@jamil2025spatiotemporal].
